### 2023-03-04 Stimuli List and Data Source/List Generation.Rmd
Added script for stimuli list and data source creation

---
title: "Stimuli List Generation"
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Randomizing Order of Phrases in Counterbalanced Lists

In this block of code, the two counterbalanced phrase lists are first uploaded. Then, the order of the stimulus presentation is randomized separately for each list. Then, 5 practice trial information is added to the beginning of the list (this is done after the randomization because the same 5 practice trials appear before actual data collection begins). Then, new variables in each of the lists are created (this is needed to create the data source csv file)

```{r}

#Loading needed packages

library(rio) #install.packages("rio")
library(tidyverse) #install.packages("tidyverse")

Phrase_List_1 <- rio::import("Phrase List 1.xlsx")
Phrase_List_2 <-rio::import("Phrase List 2.xlsx")

# randomize order of list 1

set.seed(1000)
Phrase_List_1_random <- Phrase_List_1[sample(1:nrow(Phrase_List_1)), ]

# Randomize order of list 2

set.seed(7036)
Phrase_List_2_random <- Phrase_List_2[sample(1:nrow(Phrase_List_2)), ]

# create data frame for 5 practice trials

Practice_trials <- data.frame(
  targetPhrase = c("sticks are best for pencils", "ballot formal circles", "friendly slogan catch it",
                   "she describes a nuisance", "younger rusty viewers"),
  speaker = "Practice") %>%
  mutate(code = paste("PM", row_number(), sep = "_"))

# Adding 5 practice trials to the beginning of new list one

List_1_New <- dplyr::full_join(Practice_trials, Phrase_List_1_random) %>%
  dplyr::relocate(., targetPhrase, .after = speaker) %>%
  dplyr::relocate(., code, .before = speaker) %>%
  #creating new trial number, block number, and counter balance variables
  dplyr::mutate(., trial = row_number(),
                counterBalance = 1,
                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5)/ 10) + 1)) %>%
  #moving trial and block variables in the data frame
  dplyr::relocate(trial, .before = code) %>%
  dplyr::relocate(block, .before = counterBalance)

# Adding 5 practice trials to beginning of new list 2

List_2_New <- dplyr::full_join(Practice_trials, Phrase_List_2_random) %>%
  dplyr::relocate(., targetPhrase, .after = speaker) %>%
  dplyr::relocate(., code, .before = speaker) %>%
  #creating new trial, block number, and counter balance variables
  dplyr::mutate(., trial = row_number(),
                counterBalance = 2,
                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5) /10) + 1)) %>%
  # relocating trial and block variables in data frame
  dplyr::relocate(trial, .before = code) %>%
  dplyr::relocate(block, .before = counterBalance)

#exporting the separate phrase lists in excel (used to eventually create RA stimuli lists in word)

rio::export(List_1_New, "Phrase List 1 New.xlsx")
rio::export(List_2_New, "Phrase List 2 New.xlsx")

## removing objects that are no longer needed from environment

rm(Phrase_List_1_random, Phrase_List_2_random, Phrase_List_1, Phrase_List_2, Practice_trials)

```

## Creating the DataSource file

In this block of code, the DataSource file will be created. This file will be uploaded into experiment builder to inform the program about the order in which to present the stimuli in the experiment. There are two counter balanced conditions that were created in order to control for possible phrase effects.

```{r}

# Merging the two lists together

DataSource <- dplyr::full_join(List_1_New, List_2_New)

# Creating the soundFile and number of target words variables

## soundFile refers to the name of the .wav file that is associated with each phrase/stimulus
## number of target words is a variable created for experiment builder. When listeners repeat the phrase they heard aloud in the pupillometry experiment, the experimenter will record the number of correct words by selecting the appropriate key on the host computer keyboard. Most of the phrases have 3 target words (with the exception of 2 phrases. Those only have 2 target words). Experiment Builder will compare the experimenter's input with this variable to record whether the phrase was repeated correctly (i.e. all target words recognized).

DataSource <- DataSource %>%
  dplyr::mutate(soundFile = case_when(speaker == "Practice" ~ "PM",
                                      speaker == "PD" ~ "PDM10",
                                      TRUE ~ "CM1")) %>%
  dplyr::mutate(soundFile = ifelse(soundFile == "PM", paste(soundFile, trial, sep = "_"), 
                                   paste(soundFile, code, sep = "_"))) %>%
  dplyr::mutate(soundFile = paste(soundFile, "wav", sep = ".")) %>%
  dplyr::mutate(num_targetWords = ifelse(code == "EXP_18" | code == "EXP_75", 2, 3)) %>%
  dplyr::relocate(soundFile, .before = targetPhrase) %>%
  dplyr::relocate(num_targetWords, .before = block)

# Exporting the data source data frame as a csv file. Will be imported into Experiment Builder for programming pupillometry experiment.

rio::export(DataSource, "Dysarthria_LE_DataSource.csv")

#Removing objects from the environment that are not needed anymore

rm(List_1_New, List_2_New)

```

### 2023-03-04 Preliminary Analysis/Data Processing.qmd
Added quarto data preprocessing document for preliminary analysis (still blank)

---
title: "Data Processing"
format: html
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

# About this Document

This is the standing document used for preliminary data analysis for the Listening Effort in Dysarthria project. This document contains all code for processing data from pilot data collection.

```{r}

```

You can add options to executable code like this

```{r}
#| echo: false

```

The `echo: false` option disables the printing of code (only output is displayed).

### 2023-03-05 Stimuli List and Data Source/List Generation.Rmd
added practiceTrials variable to phrase lists and data source

@@
 Practice_trials <- data.frame(
-  targetPhrase = c("sticks are best for pencils", "ballot formal circles", "friendly slogan catch it",
+  targetPhrase = c("sticks are best for pencils", "ballot formal circles", 
+                   "friendly slogan catch it",
                    "she describes a nuisance", "younger rusty viewers"),
@@
                 counterBalance = 1,
-                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5)/ 10) + 1)) %>%
-  #moving trial and block variables in the data frame
+                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5)/ 10) + 1),
+                practiceTrial = ifelse(speaker == "Practice", "Practice", "Experimental")) %>%
+  #moving new variables in the data frame
   dplyr::relocate(trial, .before = code) %>%
-  dplyr::relocate(block, .before = counterBalance)
+  dplyr::relocate(block, .before = counterBalance) %>%
+  dplyr::relocate(practiceTrial, .after = targetPhrase)
 
@@
                 counterBalance = 2,
-                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5) /10) + 1)) %>%
-  # relocating trial and block variables in data frame
+                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5) /10) + 1),
+                practiceTrial = ifelse(speaker == "Practice", "Practice", "Experimental")) %>%
+  # relocating new variables in data frame
   dplyr::relocate(trial, .before = code) %>%
-  dplyr::relocate(block, .before = counterBalance)
+  dplyr::relocate(block, .before = counterBalance) %>%
+  dplyr::relocate(practiceTrial, .after = targetPhrase)
 

### 2023-03-06 Stimuli List and Data Source/Stimuli List Generation.Rmd
Updated stimuli list code to use for loops

--- 2023-03-04 Stimuli List and Data Source/List Generation.Rmd
@@
 
-#Loading needed packages
-
 library(rio) #install.packages("rio")
@@
 
-Phrase_List_1 <- rio::import("Phrase List 1.xlsx")
-Phrase_List_2 <-rio::import("Phrase List 2.xlsx")
-
-# randomize order of list 1
-
-set.seed(1000)
-Phrase_List_1_random <- Phrase_List_1[sample(1:nrow(Phrase_List_1)), ]
-
-# Randomize order of list 2
-
-set.seed(7036)
-Phrase_List_2_random <- Phrase_List_2[sample(1:nrow(Phrase_List_2)), ]
-
-# create data frame for 5 practice trials
+#Creating Practice Trials dataframe
 
 Practice_trials <- data.frame(
-  targetPhrase = c("sticks are best for pencils", "ballot formal circles", "friendly slogan catch it",
+  targetPhrase = c("sticks are best for pencils", "ballot formal circles", 
+                   "friendly slogan catch it",
                    "she describes a nuisance", "younger rusty viewers"),
@@
 
-# Adding 5 practice trials to the beginning of new list one
+# Setting up a loop for randomizing stimuli lists
 
-List_1_New <- dplyr::full_join(Practice_trials, Phrase_List_1_random) %>%
+file_names <- c("Phrase List 1.xlsx", "Phrase List 2.xlsx")
+seed = c(1000, 7036)
+
+for (i in seq_along(file_names))  {
+
+  data <- rio::import(file_names[i])
+  set.seed(seed[i])
+  #Randomize the data
+  data_random <- data[sample(1:nrow(data)), ]
+
+  #Merging the practice trials df with the randomized list df
+  data_random <- dplyr::full_join(Practice_trials, data_random) %>%
     dplyr::relocate(., targetPhrase, .after = speaker) %>%
@@
     dplyr::mutate(., trial = row_number(), 
-                counterBalance = 1,
-                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5)/ 10) + 1)) %>%
-  #moving trial and block variables in the data frame
+                  counterBalance = i, 
+                  block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5)/ 10) + 1), 
+                  practiceTrial = ifelse(speaker == "Practice", "Practice", "Experimental")) %>%
+    #moving new variables in the data frame
     dplyr::relocate(trial, .before = code) %>%
-  dplyr::relocate(block, .before = counterBalance)
+    dplyr::relocate(block, .before = counterBalance) %>%
+    dplyr::relocate(practiceTrial, .after = targetPhrase)
 
-# Adding 5 practice trials to beginning of new list 2
+  # Renaming the new dfs
+  data_name <- paste("newList", i, sep = "")
+  assign(data_name, data_random)
 
-List_2_New <- dplyr::full_join(Practice_trials, Phrase_List_2_random) %>%
-  dplyr::relocate(., targetPhrase, .after = speaker) %>%
-  dplyr::relocate(., code, .before = speaker) %>%
-  #creating new trial, block number, and counter balance variables
-  dplyr::mutate(., trial = row_number(),
-                counterBalance = 2,
-                block = ifelse(speaker == "Practice", 1, ceiling((row_number() - 5) /10) + 1)) %>%
-  # relocating trial and block variables in data frame
-  dplyr::relocate(trial, .before = code) %>%
-  dplyr::relocate(block, .before = counterBalance)
+  # Exporting dfs (to create stimuli lists in word)
+  rio::export(data_random, paste("New_Phrase_List", i, ".csv"))
+}
 
-#exporting the separate phrase lists in excel (used to eventually create RA stimuli lists in word)
+# Removing unneeded items from environment
 
-rio::export(List_1_New, "Phrase List 1 New.xlsx")
-rio::export(List_2_New, "Phrase List 2 New.xlsx")
-
-## removing objects that are no longer needed from environment
-
-rm(Phrase_List_1_random, Phrase_List_2_random, Phrase_List_1, Phrase_List_2, Practice_trials)
+rm(data, data_random, Practice_trials, data_name, file_names, i, seed)
 
@@
 
-## Creating the DataSource file
+# Creating the DataSource file
 
@@
 
-DataSource <- dplyr::full_join(List_1_New, List_2_New)
+DataSource <- dplyr::full_join(newList1, newList2)
 
@@
 
-rm(List_1_New, List_2_New)
+rm(newList1, newList2)
 

### 2023-03-06 Stimuli List and Data Source/Stimuli List Generation.Rmd
updated stimuli list code
small edits

@@
   # Exporting dfs (to create stimuli lists in word)
-  rio::export(data_random, paste("New_Phrase_List", i, ".csv"))
+  rio::export(data_random, paste("New_Phrase_List", i, ".csv", sep = ""))
 }
@@
 
-# Merging the two lists together
+# Merging the two lists together.
 
@@
 
-# Creating the soundFile and number of target words variables
+# Creating the soundFile and number of target words variables.
 

### 2023-03-08 Preliminary Analysis/Data Preprocessing.R
Beginning preliminary analysis preprocessing script

@@ -1,0 +1,30 @@
+# Listening Effort Preliminary Analysis: Data Preprocessing 
+
+# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
+
+# Purpose: To load in edf files from EyeLink and process the raw data 
+# files to prepare for analysis.
+
+# Load Packages
+
+library(gazer) #remotes::install_github("dmirman/gazer")
+library(tidyverse) #install.packages("tidyverse")
+library(zoo) # install.packages("zoo")
+library(knitr) 
+library(devtools)
+library(edfR) #install_github("jashubbard/edfR")
+
+# Set working directories
+
+edf.path <- "iCloud Drive/Documents/Listening Effort Experiment/results"
+csv.path <- "iCloud Drive/Documents/Listening-Effort-in-Dysarthria/Preliminary Analysis/Prepped_Data"
+
+# Loading and Reading the edf files from Pilot Collection
+
+##Loads and parses each edf file
+file_list_edf <- list.files(path = edf.path, pattern = ".edf")
+parse_edf(file_list = file_list_edf, output_dir = csv.path, type = "pupil")
+
+## saving csv file list and merging all sample files 
+file_list_csv <- list.files(path = csv.path, pattern = ".csv")
+sample_gaze_data <- merge_gazer_files(file_list_csv, type = "edf")

### 2023-03-08 Preliminary Analysis/Data Preprocessing.R
edited preprocessing script. Need to change plan for importing edf files

@@ -2,8 +2,10 @@
 
 # Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
 
-# Purpose: To load in edf files from EyeLink and process the raw data 
-# files to prepare for analysis.
+# Purpose: To load in raw data files and process it to prepare for analysis 
+# Raw data files were first exported using the Data Viewer software
+# provided by EyeLink. Deblinking, smoothing, downsampling, and merging pupil
+# data with behavioral data will be done in this script.
 
 # Load Packages
 
@@ -11,12 +13,12 @@ library(gazer) #remotes::install_github("dmirman/gazer")
 library(tidyverse) #install.packages("tidyverse")
 library(zoo) # install.packages("zoo")
 library(knitr) 
-library(devtools)
+library(devtools) #install.packages("devtools")
 library(edfR) #install_github("jashubbard/edfR")
 
 # Set working directories
 
-edf.path <- "iCloud Drive/Documents/Listening Effort Experiment/results"
+edf.path <- "iCloud Drive/Documents/Listening-Effort-in-Dysarthria/Preliminary Analysis/Raw_Data"
 csv.path <- "iCloud Drive/Documents/Listening-Effort-in-Dysarthria/Preliminary Analysis/Prepped_Data"
 
 # Loading and Reading the edf files from Pilot Collection

### 2023-04-24 Stimuli List and Data Source/ALSM1 Intelligibility.Rmd
Calculated mean intelligibility for ALSM1

---
title: "Baseline Intelligibility Calculation"
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# About this Document

In the initial planning of this study, we decided to use a 80-year old male speaker with Parkinson's disease (PDM10) for our speaker with dysarthria. This decision was based on the fact that this speaker had an intelligibility level of 83.6% based on SLP ratings (Hirsch et al., 2022). However, during test runs of the experiment, the PI noticed that listeners had a harder time understanding the speaker with dysarthria. When looking at other recent work in the lab, PDM10 had an initial intelligibility level of a little over 50% (Lansford et al., 2023). This discrepancy is likely due to different speech stimuli used in the studies. This current study uses semantically anomalous phrases, like Lansford et al. (2023). Since we are aiming to choose a speaker with a milder dysarthria in our study, we decided to switch our speaker.

One potential speaker is ALSM1. In Lansford et al. (2023), this speaker had a mean initial intelligibility of about 77%. ALSM1 was also a speaker included in Hirsch et al. (2022). Therefore, in this script, we are combining the data collected from these two studies to calculate an average baseline intelligibility level for ALSM1.

You can find the full data sets for the Hirsch et al. (2022) and Lansford et al. (2023) studies at the links below.

Hirsch et al. (2022): <https://osf.io/sr9aw/>

Lansford et al. (2023): <https://osf.io/y32qc/>

```{r, warning = F, message = F, echo = F}

library(tidyverse)
library(rio)

# Load data from Lansford et al. (2023)

lansford_data <- rio::import("Data from recent lab studies/Lansford_2021_data.csv")

# Load data from Hirsch et al., (2022)

hirsch_slp_data <- rio::import("Data from recent lab studies/SLP_Correlation.csv")

hirsch_naive_data <- rio::import("Data from recent lab studies/NaiveListener_Correlation.csv")

```

## Filtering for data for ALSM1 and just pretest intelligibility (for Lansford data)

The datasets from these studies went through some minor filtering processes. For the Hirsch et al. (2022) data, both the slp and naive listener groups provided two different intelligibility ratings. For the purpose of this averaging, we used the OT scores from the naive listeners and the estimations from the SLP listeners.

```{r, warning = F, message = F}

# Cleaning Lansford Data

lansford_data <- lansford_data %>%
  dplyr::select(c(id, phase, intel, speaker)) %>%
  dplyr::filter(speaker == "ALSM1") %>%
  dplyr::filter(phase == "pretest") %>%
  dplyr::select(!phase)

# Cleaning Hirsch Naive Listener Data

# Naive listeners provided both VAS ratings and OT for two different intelligibility measures. When calculating the average intelligibility across studies, we will use the OT measure to be more consistent with Lansford et al. (2023)

hirsch_naive_data <- hirsch_naive_data %>%
  dplyr::select(!naive_VAS) %>%
  dplyr::filter(SpeakerID == "ALSM1") %>%
  dplyr::rename(speaker = "SpeakerID",
                id = "ListenerID",
                intel = "naive_OT")

# Cleaning Hirsch SLP Data

# SLP listeners provided both estimations and VAS ratings for intelligibility. When calculating the average intelligibility across studies, we will use the estimation measure since Hirsch et al. (2022) found that it was the stronger predictor of intelligibility

hirsch_slp_data <- hirsch_slp_data %>%
  dplyr::select(c(slpID, Speaker, slp_EST)) %>%
  dplyr::filter(Speaker == "ALSM1") %>%
  dplyr::rename(intel = "slp_EST",
                speaker = "Speaker",
                id = slpID)

# Combining the datasets

intel <- rbind(hirsch_naive_data, hirsch_slp_data)

intel <- rbind(intel, lansford_data)

```

## Calculating Mean Intelligibility

```{r, warning = F, message = F}

intel %>%
  dplyr::summarize(mean = mean(intel),
                   median = median(intel),
                   SD = sd(intel))

```

Therefore, across two recent studies, the mean intelligibility level for speaker ALSM1 is 77.21%. We consider this speaker to have a moderate dysarthria (Lansford et al., 2014) with relatively preserved intelligibility (intelligibility over 75%).

### 2023-04-24 Stimuli List and Data Source/Stimuli List Generation.Rmd
Updated Stimuli List

@@
 
-file_names <- c("Phrase List 1.xlsx", "Phrase List 2.xlsx")
+file_names <- c("Phrase lists/Phrase List 1.xlsx", "Phrase lists/Phrase List 2.xlsx")
 seed = c(1000, 7036)
@@
                   practiceTrial = ifelse(speaker == "Practice", "Practice", "Experimental")) %>%
+    # Changing PD speaker to ALS speaker (we decided to use a different speaker for this study)
+    dplyr::mutate(speaker = case_when(speaker == "PD" ~ "ALS",
+                                      TRUE ~ speaker)) %>%
     #moving new variables in the data frame
@@
 
-  # Exporting dfs (to create stimuli lists in word)
-  rio::export(data_random, paste("New_Phrase_List", i, ".csv", sep = ""))
 }
@@
   dplyr::mutate(soundFile = case_when(speaker == "Practice" ~ "PM",
-                                      speaker == "PD" ~ "PDM10",
+                                      speaker == "ALS" ~ "ALSM1",
                                       TRUE ~ "CM1")) %>%
@@
   dplyr::mutate(soundFile = paste(soundFile, "wav", sep = ".")) %>%
-  dplyr::mutate(num_targetWords = ifelse(code == "EXP_18" | code == "EXP_75", 2, 3)) %>%
-  dplyr::relocate(soundFile, .before = targetPhrase) %>%
-  dplyr::relocate(num_targetWords, .before = block)
+  dplyr::relocate(soundFile, .before = targetPhrase)
 
@@
 
-rio::export(DataSource, "Dysarthria_LE_DataSource.csv")
+rio::export(DataSource, "Data Source/Dysarthria_LE_DataSource.csv")
 

### 2023-04-24 Preliminary Analysis/Data Preprocessing.R
Deleted preprocessing folder (will start over)

@@ -1,32 +1,0 @@
-# Listening Effort Preliminary Analysis: Data Preprocessing 
-
-# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
-
-# Purpose: To load in raw data files and process it to prepare for analysis 
-# Raw data files were first exported using the Data Viewer software
-# provided by EyeLink. Deblinking, smoothing, downsampling, and merging pupil
-# data with behavioral data will be done in this script.
-
-# Load Packages
-
-library(gazer) #remotes::install_github("dmirman/gazer")
-library(tidyverse) #install.packages("tidyverse")
-library(zoo) # install.packages("zoo")
-library(knitr) 
-library(devtools) #install.packages("devtools")
-library(edfR) #install_github("jashubbard/edfR")
-
-# Set working directories
-
-edf.path <- "iCloud Drive/Documents/Listening-Effort-in-Dysarthria/Preliminary Analysis/Raw_Data"
-csv.path <- "iCloud Drive/Documents/Listening-Effort-in-Dysarthria/Preliminary Analysis/Prepped_Data"
-
-# Loading and Reading the edf files from Pilot Collection
-
-##Loads and parses each edf file
-file_list_edf <- list.files(path = edf.path, pattern = ".edf")
-parse_edf(file_list = file_list_edf, output_dir = csv.path, type = "pupil")
-
-## saving csv file list and merging all sample files 
-file_list_csv <- list.files(path = csv.path, pattern = ".csv")
-sample_gaze_data <- merge_gazer_files(file_list_csv, type = "edf")

### 2023-04-25 Stimuli List and Data Source/Stimuli List Generation.Rmd
Updated stimuli list

@@
 
+# Recoding the targetPhrase for phrase 69 for the ALS speaker because they said something different
+
+newList2 <- newList2 %>%
+  dplyr::mutate(targetPhrase = ifelse(code == "EXP_69", "spackle enter broken", targetPhrase))
+
 # Merging the two lists together.
@@
 ## soundFile refers to the name of the .wav file that is associated with each phrase/stimulus
-## number of target words is a variable created for experiment builder.
-When listeners repeat the phrase they heard aloud in the pupillometry experiment, the experimenter will record the number of correct words by selecting the appropriate key on the host computer keyboard.
-Most of the phrases have 3 target words (with the exception of 2 phrases.
-Those only have 2 target words).
-Experiment Builder will compare the experimenter's input with this variable to record whether the phrase was repeated correctly (i.e.
-all target words recognized).
 

### 2023-07-25 Data Preprocessing/pupil_data_preprocessing.R
Loading raw pupil data and checking experiment interval timing

@@ -1,0 +1,47 @@
+# Data Preprocessing
+
+# Author: Micah E. Hirsch
+
+# Date: 7/25/2023 (currently wip - Preliminary Processing)
+
+## Purpose: To load in raw pupil dilation data from EyeLink
+
+# Loading required packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+## Need this to install gazer and saccades packages
+library(remotes) # install.packages("remotes")
+library(gazer) # remotes::install_github("dmirman/gazer")
+library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
+library(zoo) # install.packages("zoo")
+library(knitr) # install.packages("knitr")
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Data Preprocessing/Raw Data")
+
+# Loading raw pupil data
+
+file_list <- list.files(path = ".", pattern - ".txt")
+
+pupil_data <- merge_pupil(file_list, blink_colname = "RIGHT_IN_BLINK", pupil_colname = "RIGHT_PUPIL_SIZE")
+
+mode = function(x) {
+  u = unique(x)
+  tab <-tabulate(match(x, u))
+  u[tab == max(tab)]
+}
+
+pupil_data1 <- pupil_data %>%
+  dplyr::filter(sample_message %in% c("TRIAL_START", "PHRASE_START", "PHRASE_END", "RESPONSE_CUE")) %>%
+  dplyr::select(c('subject', 'trial', 'timestamp', 'sample_message')) %>%
+  tidyr::pivot_wider(names_from = 'sample_message', values_from = 'timestamp') %>%
+  dplyr::mutate(baseline_ret = PHRASE_START - TRIAL_START,
+                retention = RESPONSE_CUE - PHRASE_END) %>%
+  dplyr::select(!c(TRIAL_START:RESPONSE_CUE)) %>%
+  tidyr::pivot_longer(cols = c(baseline_ret, retention),
+                      names_to = "silence",
+                      values_to = "interval") %>%
+  dplyr::group_by(silence) %>%
+  dplyr::summarize(min = min(interval), max = max(interval), mean = mean(interval), median = median(interval), mode = mode(interval))
+
+rio::export(pupil_data1, "interval_timing.csv")

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
filtered out rows not included in trial run

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch
 
-# Date: 7/25/2023 (currently wip - Preliminary Processing)
+# Date: 7/26/2023 (currently wip - Preliminary Processing)
 
 ## Purpose: To load in raw pupil dilation data from EyeLink
 
@@ -10,26 +10,34 @@
 
 library(rio) # install.packages("rio")
 library(tidyverse) # install.packages("tidyverse")
-## Need this to install gazer and saccades packages
+## Need remotes to install gazer and saccades packages
 library(remotes) # install.packages("remotes")
 library(gazer) # remotes::install_github("dmirman/gazer")
 library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
 library(zoo) # install.packages("zoo")
 library(knitr) # install.packages("knitr")
 
+# Setting the working directory
+
 setwd("~/Documents/Listening-Effort-in-Dysarthria/Data Preprocessing/Raw Data")
 
 # Loading raw pupil data
 
-file_list <- list.files(path = ".", pattern - ".txt")
+file_list <- list.files(path = ".", pattern = ".txt")
 
 pupil_data <- merge_pupil(file_list, blink_colname = "RIGHT_IN_BLINK", pupil_colname = "RIGHT_PUPIL_SIZE")
+
+# Checking the interval timing for the baseline and post-phrase retention periods
+
+## Creating mode function
 
 mode = function(x) {
   u = unique(x)
   tab <-tabulate(match(x, u))
   u[tab == max(tab)]
 }
+
+## Filtering out landmarks of interest relating to start of trial, phrase onsets and offsets, and presentation of response cue
 
 pupil_data1 <- pupil_data %>%
   dplyr::filter(sample_message %in% c("TRIAL_START", "PHRASE_START", "PHRASE_END", "RESPONSE_CUE")) %>%
@@ -44,4 +52,35 @@ pupil_data1 <- pupil_data %>%
   dplyr::group_by(silence) %>%
   dplyr::summarize(min = min(interval), max = max(interval), mean = mean(interval), median = median(interval), mode = mode(interval))
 
-rio::export(pupil_data1, "interval_timing.csv")
+# Filtering rows before the start of a trial and after the response cue.
+
+## Extract timestamps for start of each trial
+trial_start <- pupil_data %>%
+  dplyr::filter(sample_message == "TRIAL_START") %>%
[…]

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
Removing unneeded variables from the df and filtering out practice trials

@@ -2,9 +2,11 @@
 
 # Author: Micah E. Hirsch
 
-# Date: 7/26/2023 (currently wip - Preliminary Processing)
+# Date: 7/26/2023 (Current Version - Pilot Data, Note: This script is a work-in-progress)
 
-## Purpose: To load in raw pupil dilation data from EyeLink
+## Purpose: To load in raw pupil dilation data from EyeLink, 
+## check timing intervals for baseline and retention periods, 
+## and filter out excess rows.
 
 # Loading required packages
 
@@ -79,8 +81,14 @@ trimmed_pupil_data <- pupil_data2 %>%
 
 rm(trial_start, trial_end, file_list, pupil_data2, pupil_data)
 
-# Filtering out participants with too much data loss.
+# Filtering out trials/participants with too much data loss.
 ## None removed 
 
 trimmed_pupil_data <- gazer::count_missing_pupil(trimmed_pupil_data, missingthresh = 0.5)
+
+# Removing unneeded variables from df and filtering out practice trials.
+
+trimmed_pupil_data <- trimmed_pupil_data %>%
+  dplyr::select(!c(time:averageMissingTrial)) %>%
+  dplyr::filter(practiceTrial != 'Practice')
 

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
Trimmed pupil df created

@@ -2,11 +2,13 @@
 
 # Author: Micah E. Hirsch
 
-# Date: 7/26/2023 (Current Version - Pilot Data, Note: This script is a work-in-progress)
+# Date: 7/26/2023 
+# Current Version - Pilot Data Preprocessing
 
-## Purpose: To load in raw pupil dilation data from EyeLink, 
-## check timing intervals for baseline and retention periods, 
-## and filter out excess rows.
+## Purpose: To load in raw pupil dilation data from EyeLink and 
+## prepare data for analysis and visualization.
+
+## Note: This script is a work-in-progress.
 
 # Loading required packages
 
@@ -29,7 +31,8 @@ file_list <- list.files(path = ".", pattern = ".txt")
 
 pupil_data <- merge_pupil(file_list, blink_colname = "RIGHT_IN_BLINK", pupil_colname = "RIGHT_PUPIL_SIZE")
 
-# Checking the interval timing for the baseline and post-phrase retention periods
+# Checking the interval timing for the baseline and post-phrase retention periods.
+## We are doing this because there were some concerns about the timing of some experimental events.
 
 ## Creating mode function
 
@@ -50,7 +53,9 @@ pupil_data1 <- pupil_data %>%
   dplyr::select(!c(TRIAL_START:RESPONSE_CUE)) %>%
   tidyr::pivot_longer(cols = c(baseline_ret, retention),
                       names_to = "silence",
-                      values_to = "interval") %>%
+                      values_to = "interval") 
+
+interval_summary <- pupil_data1 %>%
   dplyr::group_by(silence) %>%
   dplyr::summarize(min = min(interval), max = max(interval), mean = mean(interval), median = median(interval), mode = mode(interval))
 
@@ -79,7 +84,7 @@ trimmed_pupil_data <- pupil_data2 %>%
 
 ## removing unneeded objects from the environment
 
-rm(trial_start, trial_end, file_list, pupil_data2, pupil_data)
+rm(trial_start, trial_end, file_list, pupil_data2, pupil_data1, pupil_data, interval_summary)
 
 # Filtering out trials/participants with too much data loss.
 ## None removed 
@@ -90,5 +95,6 @@ trimmed_pupil_data <- gazer::count_missing_pupil(trimmed_pupil_data, missingthre
 
 trimmed_pupil_data <- trimmed_pupil_data %>%
   dplyr::select(!c(time:averageMissingTrial)) %>%
[…]

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
deblinking, interpolation, and smoothing pupil tracking

@@ -98,3 +98,21 @@ trimmed_pupil_data <- trimmed_pupil_data %>%
   dplyr::filter(practicetrial != 'Practice') %>%
   dplyr::select(!practicetrial)
 
+# Deblinking
+
+pupil_extend <- trimmed_pupil_data %>%
+  dplyr::group_by(subject, trial) %>%
+  dplyr::mutate(extendpupil = extend_blinks(pupil, fillback = 50, fillforward = 160, hz = 1000))
+
+# Interpolation and Smoothing
+
+## linear interpolation
+interp <- interpolate_pupil(pupil_extend,
+                            extendblinks = T, 
+                            type = "linear", 
+                            hz = 1000)
+
+## 10 Hz moving average filter
+smoothed <- interp %>%
+  dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5))
+

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
selecting needed variables

@@ -114,5 +114,9 @@ interp <- interpolate_pupil(pupil_extend,
 
 ## 10 Hz moving average filter
 smoothed <- interp %>%
-  dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5))
+  dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) %>%
+  ## Selecting relevant variables
+  dplyr::select(c(subject, trial, timestamp, code, speaker, targetphrase, counterbalance, smoothed_pupil)) %>%
+  dplyr::relocate(smoothed_pupil, .after = timestamp)
+
 

### 2023-07-26 Data Preprocessing/pupil_data_preprocessing.R
Baseline correction, wip

@@ -116,7 +116,10 @@ interp <- interpolate_pupil(pupil_extend,
 smoothed <- interp %>%
   dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) %>%
   ## Selecting relevant variables
-  dplyr::select(c(subject, trial, timestamp, code, speaker, targetphrase, counterbalance, smoothed_pupil)) %>%
-  dplyr::relocate(smoothed_pupil, .after = timestamp)
+  dplyr::select(c(subject, trial, sample_message, timestamp, code, speaker, targetphrase, counterbalance, smoothed_pupil)) %>%
+  dplyr::relocate(smoothed_pupil, .after = timestamp) 
 
+# Baseline Correction
 
+baseline_pupil <- baseline_correction_pupil_msg(smoothed, pupil_colname = "smoothed_pupil", 
+                                                baseline_dur = 500, event = "PHRASE_START", baseline_method = "sub")

### 2023-07-27 Data Preprocessing/pupil_data_preprocessing.R
Aligned data to onset of phrase

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch
 
-# Date: 7/26/2023 
+# Date: 7/27/2023 
 # Current Version - Pilot Data Preprocessing
 
 ## Purpose: To load in raw pupil dilation data from EyeLink and 
@@ -82,21 +82,37 @@ pupil_data2 <- pupil_data2 %>%
 trimmed_pupil_data <- pupil_data2 %>%
   dplyr::filter(timestamp >= start_time & timestamp <= end_time)
 
-## removing unneeded objects from the environment
-
-rm(trial_start, trial_end, file_list, pupil_data2, pupil_data1, pupil_data, interval_summary)
-
 # Filtering out trials/participants with too much data loss.
 ## None removed 
 
 trimmed_pupil_data <- gazer::count_missing_pupil(trimmed_pupil_data, missingthresh = 0.5)
 
-# Removing unneeded variables from df and filtering out practice trials.
+# Prepping the df for pupil processing
+
+## Extracting phrase onset times
+phrase_start <- pupil_data %>%
+  dplyr::filter(sample_message == "PHRASE_START") %>%
+  dplyr::select(subject, trial, phrase_start_time = timestamp)
+
+## Merging phrase onset times with pupil df
 
 trimmed_pupil_data <- trimmed_pupil_data %>%
-  dplyr::select(!c(time:averageMissingTrial)) %>%
+  dplyr::left_join(phrase_start, by = c("subject", "trial"))
+
+trimmed_pupil_data <- trimmed_pupil_data %>%
+  ## Removing practice trials from df
   dplyr::filter(practicetrial != 'Practice') %>%
-  dplyr::select(!practicetrial)
+  dplyr::select(!practicetrial) %>%
+  ## Aligning data to onset of phrase presentation
+  dplyr::mutate(time_c = timestamp - phrase_start_time) %>%
+  ## Removing unneeded variables
+  dplyr::select(!c(timestamp, time:phrase_start_time)) %>%
+  dplyr::relocate(time_c, .after = pupil)
+  
+
+## removing unneeded objects from the environment
+
+rm(trial_start, trial_end, phrase_start, file_list, pupil_data2, pupil_data1, pupil_data, interval_summary)
 
 # Deblinking
 

### 2023-07-27 Data Preprocessing/pupil_data_preprocessing.R
Baseline pupil correction and rapid pupil change check

@@ -95,7 +95,6 @@ phrase_start <- pupil_data %>%
   dplyr::select(subject, trial, phrase_start_time = timestamp)
 
 ## Merging phrase onset times with pupil df
-
 trimmed_pupil_data <- trimmed_pupil_data %>%
   dplyr::left_join(phrase_start, by = c("subject", "trial"))
 
@@ -128,14 +127,28 @@ interp <- interpolate_pupil(pupil_extend,
                             type = "linear", 
                             hz = 1000)
 
-## 10 Hz moving average filter
+## 10 Hz 5-point moving average filter
 smoothed <- interp %>%
   dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) %>%
   ## Selecting relevant variables
-  dplyr::select(c(subject, trial, sample_message, timestamp, code, speaker, targetphrase, counterbalance, smoothed_pupil)) %>%
-  dplyr::relocate(smoothed_pupil, .after = timestamp) 
+  dplyr::select(c(subject, trial, sample_message, time_c, code, speaker, targetphrase, counterbalance, smoothed_pupil)) %>%
+  dplyr::relocate(smoothed_pupil, .after = time_c) %>%
+  dplyr::rename(time = time_c)
 
 # Baseline Correction
 
-baseline_pupil <- baseline_correction_pupil_msg(smoothed, pupil_colname = "smoothed_pupil", 
-                                                baseline_dur = 500, event = "PHRASE_START", baseline_method = "sub")
+baseline_pupil <- baseline_correction_pupil(smoothed, pupil_colname = "smoothed_pupil",
+                                            baseline_window = c(-500, 0))
+
+# Artifact Rejection
+
+## Looking for rapid changes in pupil dilation using median absolute deviation
+
+mad_removal <- baseline_pupil %>%
+  dplyr::group_by(subject, trial) %>%
+  dplyr::mutate(speed = speed_pupil(baselinecorrectedp, time)) %>%
+  dplyr::mutate(MAD = calc_mad(speed, n=16)) %>%
+  dplyr::filter(speed < MAD)
+
+## Proportion of rows removed (7/27/23: 0.104%)
+((nrow(baseline_pupil) - nrow(mad_removal)) / nrow(baseline_pupil)) * 100

### 2023-07-27 Data Preprocessing/pupil_data_preprocessing.R
downsampled pupil data

@@ -152,3 +152,19 @@ mad_removal <- baseline_pupil %>%
 
 ## Proportion of rows removed (7/27/23: 0.104%)
 ((nrow(baseline_pupil) - nrow(mad_removal)) / nrow(baseline_pupil)) * 100
+
+# Removing unneeded items from environment
+
+rm(baseline_pupil, interp, pupil_extend, smoothed, trimmed_pupil_data)
+
+# Downsampling
+
+bin.length <- 20
+
+data.binned <- mad_removal %>%
+  mutate(timebins = round(time/bin.length)*bin.length) %>%
+  dplyr::group_by(subject, trial, speaker, timebins, code, targetphrase, counterbalance) %>%
+  dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
+  dplyr::ungroup()
+
+

### 2023-07-27 Data Preprocessing/pupil_data_preprocessing.R
Exported preliminary cleaned pupil data csv

@@ -167,4 +167,10 @@ data.binned <- mad_removal %>%
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
   dplyr::ungroup()
 
+# Exporting Data (7/27/23: Pilot Data)
 
+## set wd
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Cleaned Data")
+
+rio::export(data.binned, "cleaned_pupil_data.csv")

### 2023-07-27 Data Preprocessing/transcription_preprocessing.R
calculated repetition accuracy variable

@@ -1,0 +1,95 @@
+# Intelligibility Data Preprocessing
+
+# Author: Micah E. Hirsch
+
+# Date: 7/27/23 
+# Current Version - Pilot Data
+
+## Purpose: To load in transcriptions of listener responses, 
+## determine phrase recognition accuracy, and to merge with the pupil data.
+
+## Note: This script is a work-in-progress
+
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Remotes is needed to install autoscore
+library(remotes) # install.packages("autoscore")
+library(autoscore) # remotes::install_github("autoscore/autoscore")
+library(SnowballC) # install.packages("SnowballC)
+
+# Load in transcription data
+
+## Set working directory to raw transcription csv file
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Data Preprocessing/Raw Data")
+
+## Load in transcription data
+transcriptions <- rio::import("transcriptions.csv", header = T)
+
+# Preparing the script for scoring
+
+transcriptions <- transcriptions %>%
+  # Removing unneeded columns generated by Qualtrics
+  dplyr::select(!c(StartDate:RA, Time, Counter_List)) %>%
+  # Removing first two rows
+  dplyr::filter(grepl("LE", ID)) %>%
+  dplyr::filter(!grepl("Participant", ID)) %>%
+  # LE01 has multiple submissions. Only keeping the completed submission.
+  dplyr::filter(`85` != "") %>%
+  # converting df to long format
+  tidyr::pivot_longer(cols = "1":"85",
+                      names_to = "trial",
+                      values_to = "response") %>%
+  dplyr::rename(id = ID) %>%
+  dplyr::mutate(trial = as.numeric(trial))
+
+# Load in cleaned pupil data
+
+pupil <- rio::import("cleaned_pupil_data.csv")
+
+targets <- pupil %>%
+  dplyr::rename(id = subject,
+                target = targetphrase) %>%
+  dplyr::select(id, trial, target) %>%
+  distinct()
+
+# Merging listener responses with pupil data
+
+data <- targets %>%
+  dplyr::left_join(transcriptions, by = c("id", "trial"))
+
+# Count number of words in target phrase
+
+data <- data %>%
+  dplyr::mutate(target_number = str_count(target, "\\S+"))
+
+# Counting correct number of words using Autoscore
+
[…]

### 2023-07-27 Data Preprocessing/transcription_preprocessing.R
calculated rep accuracy variable. saved script

@@ -93,3 +93,10 @@ cleaned_data <- pupil %>%
 
 rm(data, pupil, targets, transcriptions)
 
+# Export csv (7/27/23: Pilot Data)
+
+## Setting working directory
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Cleaned Data")
+
+rio::export(cleaned_data, "cleaned_data.csv")

### 2023-07-27 Peliminary Data Analysis and Vizualization.Rmd
Beginning preliminary data visualizations

---
title: "Preliminary Data Analysis and Visualization"
author: "Micah E. Hirsch"
date: "2023-07-27"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Document Purpose

The purpose of this document is to conduct preliminary analyses on pilot data for the author's dissertation project. This preliminary Analysis is based on data collected from 3 participants: LE01, LE02, and LE04. Something went wrong with LE05's transcriptions, so their data is not included in these visualizations.

```{r, include = F, warning = F, message = F}

# Load packages

library(rio)
library(tidyverse)

# import dataset

cleaned_df <- rio::import("Cleaned Data/cleaned_data.csv")

```

# Visualizations

Below are a few plots showing pupil dilation over time.

## Overall Pupil Dilation Tracking

```{r, warning = F, message = F}

cleaned_df %>%
  # For this graph, we are only looking at when the listener was accurate
  dplyr::filter(rep_acc == "accurate") %>%
  dplyr::group_by(timebins, speaker) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = speaker,
      fill = speaker) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
  theme_bw() + 
  theme(legend.position = "bottom")

```

## Comparing Accurate and Inaccurate Responses

```{r, warning = F, message = F}

cleaned_df %>%
  dplyr::filter(subject != "LE05") %>% #There is an issue with LE05 at the moment
  dplyr::group_by(timebins, speaker, rep_acc) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = rep_acc,
      fill = rep_acc) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Repetition Accuracy", color = "Repetition Accuracy") +
  theme_bw() + 
  facet_wrap("speaker") +
  theme(legend.position = "bottom")

```
## By Participant

```{r}

cleaned_df %>%
  dplyr::filter(rep_acc == "accurate") %>%
  dplyr::group_by(timebins, speaker, subject) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      fill = speaker,
      color = speaker) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,3000)) +
  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
  theme_bw() + 
  facet_wrap("subject") +
  theme(legend.position = "bottom")

```

### 2023-07-27 Peliminary Data Analysis and Vizualization.Rmd
prelim visualizations

@@
 
-```{r}
+```{r, message = F, warning = F}
 
@@
   theme_bw() + 
-  facet_wrap("subject") +
+  facet_wrap("subject", ncol = 2) +
   theme(legend.position = "bottom")

### 2023-10-23 Data Preprocessing/pupil_data_preprocessing.R
Testing Github Desktop App
Made a small test change for using Github desktop app

@@ -182,3 +182,6 @@ rm(mad_removal, bin.length)
 # Exporting Data (8/28/23: Pilot Data)
 
 rio::export(data.binned, "cleaned_pupil_data.csv")
+
+
+## Test Change

### 2023-10-23 Data Preprocessing/pupil_data_preprocessing.R
test change #2

@@ -182,6 +182,3 @@ rm(mad_removal, bin.length)
 # Exporting Data (8/28/23: Pilot Data)
 
 rio::export(data.binned, "cleaned_pupil_data.csv")
-
-
-## Test Change

### 2023-10-23 ASHA 2023 Analysis/pupil_data_preprocessing.R
Reorganized analysis files and updated gitignore

@@ -1,0 +1,11 @@
+# Pupil Data Preprocessing
+
+# Author: Micah E. Hirsch
+
+# Date: 10/23/2023
+
+## Purpose: To load in raw data from  Eyelink and prepare pupil data for analysis and visualization.
+## This data processing and analysis is being completed for a poster presentation 
+## at the Annual American Speech Language Hearing Association Convention being
+## held in Boston, MA in November 2023.
+

### 2023-10-23 ASHA 2023 Analysis/pupil_data_preprocessing.R
Started pupil data processing script
data preprocessing for ASHA 2023 presentation

@@ -1,11 +1,62 @@
 # Pupil Data Preprocessing
 
-# Author: Micah E. Hirsch
+# Author: Micah E. Hirsch, mhirsch@fsu.edu 
 
 # Date: 10/23/2023
 
-## Purpose: To load in raw data from  Eyelink and prepare pupil data for analysis and visualization.
+## Purpose: To load in raw data from Eyelink and prepare pupil data viz and analysis.
 ## This data processing and analysis is being completed for a poster presentation 
 ## at the Annual American Speech Language Hearing Association Convention being
 ## held in Boston, MA in November 2023.
 
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+## Need remotes to install gazer and saccades packages
+library(remotes) # install.packages("remotes")
+library(gazer) # remotes::install_github("dmirman/gazer")
+library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
+library(zoo) # install.packages("zoo")
+library(knitr) # install.packages("knitr")
+
+# Set the working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Load raw pupil files
+
+## Get file names from raw data folder
+file_list <- list.files(path = ".", pattern = ".txt")
+
+## Create empty list to temporarily store imported participant data
+data_list <- list()
+
+## Initiating loop to load in data for each participant
+for (file in file_list) {
+  
+  # Import file
+  data <- rio::import(file) 
+  
+  data <- data %>%
+    # Creating pupil and blink variables. 
+    # Most participants had their right eye tracked, but some had their left eye tracked.
+    dplyr::mutate(pupil = ifelse(EYE_TRACKED == "Right", RIGHT_PUPIL_SIZE, LEFT_PUPIL_SIZE),
+                  blink = ifelse(EYE_TRACKED == "Right", RIGHT_IN_BLINK, LEFT_IN_BLINK)) %>%
+    # Bring variable names to lower case
+    dplyr::rename_all(., .funs = tolower) %>%
+    # Renaming variables so it is consistent with gazer's requirements
+    dplyr::rename(subject = recording_session_label,
+                  trial = trial_index) %>%
+    # Selecting needed variables for processing
+    dplyr::select(subject, trial, eye_tracked, blink, timestamp, pupil, ip_start_time, 
+                  sample_message, effort_rating, code, speaker, practicetrial, 
[…]

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Finished for loop for loading pupil data

@@ -5,9 +5,9 @@
 # Date: 10/23/2023
 
 ## Purpose: To load in raw data from Eyelink and prepare pupil data viz and analysis.
-## This data processing and analysis is being completed for a poster presentation 
-## at the Annual American Speech Language Hearing Association Convention being
-## held in Boston, MA in November 2023.
+## This particular data processing and analysis is being completed for a poster 
+## presentation at the Annual American Speech Language Hearing Association Convention 
+## being held in Boston, MA in November 2023.
 
 # Loading Required Packages
 
@@ -35,28 +35,59 @@ data_list <- list()
 ## Initiating loop to load in data for each participant
 for (file in file_list) {
   
-  # Import file
+  ## Import file
   data <- rio::import(file) 
   
-  data <- data %>%
-    # Creating pupil and blink variables. 
-    # Most participants had their right eye tracked, but some had their left eye tracked.
-    dplyr::mutate(pupil = ifelse(EYE_TRACKED == "Right", RIGHT_PUPIL_SIZE, LEFT_PUPIL_SIZE),
-                  blink = ifelse(EYE_TRACKED == "Right", RIGHT_IN_BLINK, LEFT_IN_BLINK)) %>%
-    # Bring variable names to lower case
-    dplyr::rename_all(., .funs = tolower) %>%
-    # Renaming variables so it is consistent with gazer's requirements
-    dplyr::rename(subject = recording_session_label,
-                  trial = trial_index) %>%
-    # Selecting needed variables for processing
-    dplyr::select(subject, trial, eye_tracked, blink, timestamp, pupil, ip_start_time, 
-                  sample_message, effort_rating, code, speaker, practicetrial, 
-                  targetphrase, counterbalance)
+  ## For subject id's ending in _2, we have to discard their first 15 trials
+  ## because the experiment was restarted for them after the 15th trial
+  ## (due to technical errors). The first 15 trials were recorded in 
+  ## the initial output file for the participant.
+  if (any(grepl("_2", data$RECORDING_SESSION_LABEL))) {
+    
+    data <- data %>%
+      # Creating pupil and blink variables. 
+      # Most participants had their right eye tracked, but some had their left eye tracked.
+      dplyr::mutate(pupil = ifelse(EYE_TRACKED == "Right", RIGHT_PUPIL_SIZE, LEFT_PUPIL_SIZE),
[…]

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Initial pupil deblinking, interpolation, deblinking, and baseline correction

@@ -91,3 +91,94 @@ pupil_data <- do.call(rbind, data_list)
 ## Removing unneeded items from the environment
 
 rm(data, data_list, file, file_list)
+
+# Filtering out rows before the trial start and after the response cue
+
+## Extracting trial start times
+
+trial_start <- pupil_data %>%
+  dplyr::filter(sample_message == "TRIAL_START") %>%
+  dplyr::select(subject, trial, start_time = timestamp)
+
+## Extracting phrase start times
+
+phrase_start <- pupil_data %>%
+  dplyr::filter(sample_message == "PHRASE_START") %>%
+  dplyr::select(subject, trial, phrase_start = timestamp)
+
+## Extracting phrase end times
+
+phrase_end <- pupil_data %>%
+  dplyr::filter(sample_message == "PHRASE_END") %>%
+  dplyr::select(subject, trial, phrase_end = timestamp)
+
+## Extracting trial end times (e.g. time response cue was presented)
+
+trial_end <- pupil_data %>%
+  dplyr::filter(sample_message == "RESPONSE_CUE") %>%
+  dplyr::select(subject, trial, end_time = timestamp)
+
+## Merging the dfs together
+
+pupil_data2 <- pupil_data %>%
+  dplyr::left_join(trial_start, by = c("subject", "trial")) %>%
+  dplyr::left_join(phrase_start, by =c("subject", "trial")) %>%
+  dplyr::left_join(phrase_end, by = c("subject", "trial")) %>%
+  dplyr::left_join(trial_end, by = c("subject", "trial"))
+
+## Filter out unneeded rows and trials
+
+trimmed_pupil_data <- pupil_data2 %>%
+  ## Filtering out rows before trial start and after trial end
+  dplyr::filter(timestamp >= start_time & timestamp <= end_time) %>%
+  ## Removing practice trials from df
+  dplyr::filter(practicetrial != 'Practice') %>%
+  dplyr::select(!practicetrial) %>%
+  ## Aligning data to onset of phrase presentation
+  dplyr::mutate(time_c = timestamp - phrase_start) %>%
+  ## Marking the end of a phrase presentation for each trial
+  dplyr::mutate(end_phrase = ifelse(phrase_end == timestamp, 1, 0)) %>%
+  ## Removing unneeded variables
+  dplyr::select(!c(timestamp, start_time:end_time)) %>%
+  dplyr::relocate(time_c, .after = pupil)
+
[…]

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Artifact removal

@@ -181,4 +181,20 @@ smoothed <- interp %>%
 
 baseline_pupil <- baseline_correction_pupil(smoothed, pupil_colname = "smoothed_pupil",
                                             baseline_window = c(-500, 0))
+# Artifact Rejection
 
+## Looking for rapid changes in pupil dilation using median absolute deviation
+
+mad_removal <- baseline_pupil %>%
+  dplyr::group_by(subject, trial) %>%
+  dplyr::mutate(speed = speed_pupil(baselinecorrectedp, time)) %>%
+  dplyr::mutate(MAD = calc_mad(speed, n=16)) %>%
+  dplyr::filter(speed < MAD)
+
+## Proportion of rows removed (as of 10/24/2023: 1.28%)
+
+((nrow(baseline_pupil) - nrow(mad_removal)) / nrow(baseline_pupil)) * 100
+
+## Removing unneeded items from the environment
+
+rm(baseline_pupil, interp, pupil_extend, smoothed, trimmed_pupil_data)

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Downsampling
There still seems to be something off by this output. Will need to investigate further

@@ -198,3 +198,15 @@ mad_removal <- baseline_pupil %>%
 ## Removing unneeded items from the environment
 
 rm(baseline_pupil, interp, pupil_extend, smoothed, trimmed_pupil_data)
+
+# Downsampling
+
+bin.length <- 20
+
+data.binned <- mad_removal %>%
+  mutate(timebins = round(time/bin.length)*bin.length) %>%
+  dplyr::group_by(subject, trial, speaker, timebins, effort_rating,
+                  code, targetphrase, counterbalance, end_phrase) %>%
+  dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
+  dplyr::ungroup()
+

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Loading and cleaning transcription data from Qualtrics

@@ -209,4 +209,3 @@ data.binned <- mad_removal %>%
                   code, targetphrase, counterbalance, end_phrase) %>%
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
   dplyr::ungroup()
-

### 2023-10-24 ASHA 2023 Analysis/transcription_data_preprocessing.R

@@ -1,0 +1,45 @@
+# Intelligibility Data Processing
+
+# Author: Micah E. Hirsch, mhirsch@fsu.edu
+
+# Date: 10/24/2023
+
+## Purpose: To load in transcriptions of the listener responses,
+## determine phrase recognition accuracy, and merge with the pupil data.
+
+# Loading required packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Remotes is needed to download autoscore
+library(remotes) # install.packages("remotes")
+library(autoscore) # remotes::install_github("autoscore/autoscore")
+library(SnowballC) # install.packages("SnowballC")
+
+# Set working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Load in data
+
+transcriptions <- rio::import("transcriptions.csv", header = T)
+
+# Preparing the script for scoring
+
+transcriptions <- transcriptions %>%
+  # Removing unneeded columns generated by Qualtrics
+  dplyr::select(!c(StartDate:RA, Time, Counter_List)) %>%
+  # Removing first two rows
+  dplyr::filter(grepl("LE", ID)) %>%
+  dplyr::filter(!grepl("Participant", ID)) %>%
+  # LE01 has multiple submissions. Only keeping the completed submission.
+  dplyr::filter(`85` != "") %>%
+  # converting df to long format
+  tidyr::pivot_longer(cols = "1":"85",
+                      names_to = "trial",
+                      values_to = "response") %>%
+  dplyr::rename(id = ID) %>%
+  dplyr::mutate(trial = as.numeric(trial))
+
+
+

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Small edits to both pupil and transcription preprocessing scripts

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, mhirsch@fsu.edu 
 
-# Date: 10/23/2023
+# Date: 10/24/2023
 
 ## Purpose: To load in raw data from Eyelink and prepare pupil data viz and analysis.
 ## This particular data processing and analysis is being completed for a poster 
@@ -209,3 +209,9 @@ data.binned <- mad_removal %>%
                   code, targetphrase, counterbalance, end_phrase) %>%
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
   dplyr::ungroup()
+
+# Set working directory to export data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Cleaned Data")
+
+rio::export(data.binned, "cleaned_pupil_data.csv")

### 2023-10-24 ASHA 2023 Analysis/transcription_data_preprocessing.R

@@ -6,6 +6,9 @@
 
 ## Purpose: To load in transcriptions of the listener responses,
 ## determine phrase recognition accuracy, and merge with the pupil data.
+## This particular data processing and analysis is being completed for a
+## poster presentation at the Annual American Speech Language Hearing 
+## Association Convention being held in Boston, MA in November 2023.
 
 # Loading required packages
 

### 2023-10-24 ASHA 2023 Analysis/pupil_data_preprocessing.R
Exported cleaned pupil data
Fixed minor errors in pupil preprocessing script and created cleaned data folder in ASHA 2023 Analysis subfolder.

@@ -136,8 +136,6 @@ trimmed_pupil_data <- pupil_data2 %>%
   dplyr::select(!practicetrial) %>%
   ## Aligning data to onset of phrase presentation
   dplyr::mutate(time_c = timestamp - phrase_start) %>%
-  ## Marking the end of a phrase presentation for each trial
-  dplyr::mutate(end_phrase = ifelse(phrase_end == timestamp, 1, 0)) %>%
   ## Removing unneeded variables
   dplyr::select(!c(timestamp, start_time:end_time)) %>%
   dplyr::relocate(time_c, .after = pupil)
@@ -172,8 +170,7 @@ smoothed <- interp %>%
   dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) %>%
   ## Selecting relevant variables
   dplyr::select(c(subject, trial, sample_message, time_c, effort_rating, 
-                  code, speaker, targetphrase, counterbalance, 
-                  end_phrase, smoothed_pupil)) %>%
+                  code, speaker, targetphrase, counterbalance,smoothed_pupil)) %>%
   dplyr::relocate(smoothed_pupil, .after = time_c) %>%
   dplyr::rename(time = time_c)
 
@@ -206,12 +203,12 @@ bin.length <- 20
 data.binned <- mad_removal %>%
   mutate(timebins = round(time/bin.length)*bin.length) %>%
   dplyr::group_by(subject, trial, speaker, timebins, effort_rating,
-                  code, targetphrase, counterbalance, end_phrase) %>%
+                  code, targetphrase, counterbalance) %>%
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
   dplyr::ungroup()
 
 # Set working directory to export data
 
-setwd("~/Documents/Listening-Effort-in-Dysarthria/Cleaned Data")
+setwd("~/Documents/Listening-Effort-in-Dysarthria/ASHA 2023 Analysis/Cleaned Data")
 
 rio::export(data.binned, "cleaned_pupil_data.csv")

### 2023-10-24 ASHA 2023 Analysis/transcription_data_preprocessing.R
Load cleaned pupil data into transcriptions preprocessing script

@@ -44,5 +44,11 @@ transcriptions <- transcriptions %>%
   dplyr::rename(id = ID) %>%
   dplyr::mutate(trial = as.numeric(trial))
 
+# Loading in pupil data 
 
+## Set working directory to location of cleaned pupil data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/ASHA 2023 Analysis/Cleaned Data")
+
+pupil <- rio::import("cleaned_pupil_data.csv")
 

### 2023-10-25 ASHA 2023 Analysis/transcription_data_preprocessing.R
Cleaned transcription data and merged with cleaned pupil data

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, mhirsch@fsu.edu
 
-# Date: 10/24/2023
+# Date: 10/25/2023
 
 ## Purpose: To load in transcriptions of the listener responses,
 ## determine phrase recognition accuracy, and merge with the pupil data.
@@ -44,7 +44,7 @@ transcriptions <- transcriptions %>%
   dplyr::rename(id = ID) %>%
   dplyr::mutate(trial = as.numeric(trial))
 
-# Loading in pupil data 
+# Calculating phrase accuracy 
 
 ## Set working directory to location of cleaned pupil data
 
@@ -52,3 +52,50 @@ setwd("~/Documents/Listening-Effort-in-Dysarthria/ASHA 2023 Analysis/Cleaned Dat
 
 pupil <- rio::import("cleaned_pupil_data.csv")
 
+## Extract phrase targets from pupil df
+
+targets <- pupil %>%
+  dplyr::rename(id = subject,
+                target = targetphrase) %>%
+  dplyr::select(id, trial, target) %>%
+  distinct()
+  
+## Merging transcription and target dfs
+
+phrase_acc <- targets %>%
+  dplyr::left_join(transcriptions, by = c("id", "trial")) %>%
+  ## Count number of words in target phrase
+  dplyr::mutate(target_number = str_count(target, "\\S+")) 
+
+## Counting number of correctly recognized words
+
+phrase_acc <- autoscore::autoscore(
+  phrase_acc,
+  acceptable_df = autoscore::acceptable_spellings,
+  plural_rule = T,
+  plural_add_rule = T,
+  tense_rule = T,
+  tense_add_rule = T,
+  a_the_rule = T,
+  double_letter_rule = T) %>%
+  dplyr::rename(., correct_words = autoscore)
+
+# Creating dichotomous phrase repetition accuracy variable
+
+phrase_acc <- phrase_acc %>%
+  dplyr::mutate(rep_acc = ifelse(target_number == correct_words, "accurate", "inaccurate")) %>%
+  dplyr::rename(subject = id,
+                targetphrase = target)
+
+# Merge phrase accuracy back with the rest of the pupil data
+
+cleaned_data <- pupil %>%
+  dplyr::left_join(phrase_acc, by = c("subject", "trial", "targetphrase"))
+
+# Removing unneeded items from the environment
+
+rm(phrase_acc, pupil, targets, transcriptions)
+
[…]

### 2023-10-25 ASHA 2023 Analysis/listener_demo_preprocessing.R
Cleaned listener demo data

@@ -1,0 +1,69 @@
+# Participant Data Processing 
+
+# Author: Micah E. Hirsch, mhirsch@fsu.edu
+
+# Date: 10/25/2023
+
+## Purpose: To load and clean listener demographic information 
+## and export a cleaned version of the data. This preprocessing was completed
+## for a poster presentation at the Annual American Speech Language Hearing 
+## Association Convention held in Boston, MA in November 2023.
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+
+# Load in listener demographic data from RedCap
+
+## Set working directory
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+demo <- rio::import("participant_demo_raw.csv")
+
+# Cleaning demographic data
+
+demo1 <- demo %>%
+  # Removing identifiable and unneeded variables from df
+  dplyr::select(!c(date_participated, dob, sex, fluent_eng, demographic_information_complete)) %>%
+  # Recoding variables so they are easily interpretable
+  dplyr::mutate(gender = case_when(gender == 1 ~ "man",
+                                   gender == 2 ~ "woman",
+                                   gender == 3 ~ "nonbinary", 
+                                   gender == 4 ~ "questioning",
+                                   gender == 5 ~ "prefer not to answer",
+                                   TRUE ~ "gender not listed"),
+                ethnicity = case_when(ethnicity == 1 ~ "Hispanic/Latino(a/e)",
+                                      ethnicity == 2 ~ "Not Hispanic/Latino(a/e)",
+                                      TRUE ~ "prefer not to answer"),
+                race = case_when(race == 1 ~ "white/Caucasian",
+                                 race == 2 ~ "Black/African American",
+                                 race == 3 ~ "Asian/Asian American",
+                                 race == 4 ~ "Native Hawaiian or Other Pacific Islander",
+                                 race == 5 ~ "Native American or Alaska Native",
+                                 race == 6 ~ "Biracial or Multiracial",
+                                 race == 7 ~ "prefer not to answer",
+                                 TRUE ~ "race not listed"),
+                native_lang = ifelse(native_lang == 1, "American English", "Not American English"),
+                commdis_hx = ifelse(commdis_hx == 1, "no", "yes"),
+                fam_commdis = ifelse(fam_commdis == 1, "no", "yes"),
+                fam_comdis_2 = case_when(fam_comdis_2 == 1 ~ "no experience",
[…]

### 2023-10-25 Pilot Data Analysis/Peliminary Data Analysis and Vizualization.Rmd

---
title: "Preliminary Data Analysis and Visualization"
author: "Micah E. Hirsch"
date: "2023-07-27"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Document Purpose

The purpose of this document is to conduct preliminary analyses on pilot data for the author's dissertation project.

```{r, include = F, warning = F, message = F}

# Load packages

library(rio)
library(tidyverse)
library(ggpubr)

# import dataset

cleaned_df <- rio::import("Cleaned Data/cleaned_data.csv")

```

# Descriptives

First, we are going to determine on average, how long the phrases were for each trial. This is done so we can highlight our correct region of interest (ROI) for future analyses in the figures showing pupil dilation. Our ROI is from 500 ms after the phrase onset and 1000 ms after the phrase ends.

As you can see, on average, the ALS speaker's phrases were much longer than the control speaker's phrases. We will need to keep this in mind when creating the visuals.

```{r}

# Extracting the end times into a new df. 
#The timebins are aligned to the start of the phrase, so knowing the timebins where the phrase ends will tell us how long the phrase presentation was for each listener's trial.

# Saving each unique phrase end time in a df (might need for a future analysis)

end_times <- cleaned_df %>%
  dplyr::select(c(subject, trial, speaker, timebins, code, end_phrase)) %>%
  dplyr::filter(end_phrase == 1) %>%
  dplyr::mutate(phrase_file = paste(speaker, code, sep = "_")) %>%
  dplyr::select(phrase_file, speaker, code, timebins) %>%
  dplyr::distinct(phrase_file, .keep_all = T)

# Getting the average end time for each speaker's phrase.

end_times %>%
  dplyr::group_by(speaker) %>%
  dplyr::summarize(phrase_length = mean(timebins), sd = sd(timebins))

```
We can also see the overall average phrase length across the two speakers

```{r}

end_times %>%
  dplyr::summarize(phrase_length = mean(timebins))

```

# Visualizations

```{r}

normed <- cleaned_df %>%
  dplyr::filter(timebins >= 0) %>%
  dplyr::mutate(normed_time = scales::rescale(timebins, scale = c(0, 100)))

```

## Pupil Dilation

Below are a few plots showing pupil dilation over time.

### Comparing Accurate and Inaccurate Responses for the ALS speaker

```{r, warning = F, message = F}

ALS <- cleaned_df %>%
  dplyr::filter(speaker == "ALS") %>%
  dplyr::group_by(timebins, rep_acc) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = rep_acc,
      fill = rep_acc) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,5000), ylim = c(-100, 200)) +
  labs(title = "ALS", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Repetition Accuracy", color = "Repetition Accuracy") +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = 1)

ALS

```
## By Participant

```{r, message = F, warning = F}

control <- cleaned_df %>%
  dplyr::filter(speaker == "Control") %>%
  dplyr::group_by(timebins, rep_acc) %>%
  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
  ggplot() +
  aes(x = timebins,
      y = pupil_dilation,
      color = rep_acc,
      fill = rep_acc) +
  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
  geom_line() +
  coord_cartesian(xlim = c(0,2500), ylim = c(-100, 200)) +
  labs(title = "Control", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Repetition Accuracy", color = "Repetition Accuracy") +
  theme_bw() + 
  theme(legend.position = "bottom", aspect.ratio = 1)

control

```

```{r}

MSC_plot <- ggarrange(ALS, control, ncol = 2, common.legend = T, legend = "bottom")

MSC_plot

ggsave("MSC_plot.png", plot = MSC_plot, width = 10, height = 5, units = "in")

```

## Perceived Listening Effort

```{r}

perceived_effort <- cleaned_df %>%
  dplyr::select(subject, trial, speaker, effort_rating, rep_acc) %>%
  dplyr::distinct() %>%
  dplyr::group_by(speaker, rep_acc) %>%
  dplyr::summarize(perceived_effort = mean(effort_rating), 
                   se = sd(effort_rating, na.rm = T)/sqrt(n()))

perceived_effort %>%
  ggplot() +
  aes(x = rep_acc,
      y = perceived_effort,
      color = rep_acc,
      fill = rep_acc) +
  geom_bar(stat = "identity", alpha = 0.6) +
  geom_errorbar(aes(x = rep_acc, ymin = perceived_effort - se, ymax = perceived_effort + se), width = 0.4) +
  facet_wrap("speaker") +
  labs(x = "Repetition Accuracy", y = "Perceived Listening Effort Rating") +
  theme_bw() +
  theme(legend.position = "none", aspect.ratio = 1)

ggsave("Perceived_effort.png", plot = last_plot())

```

### 2023-10-26 ASHA 2023 Analysis/data_analysis.Rmd
Initiated data analysis rmd file

---
title: "Data Analysis"
author: "Micah E. Hirsch"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
summary(cars)
```

# Descriptive Statistics

## Participant Demographics

```{r}
plot(pressure)
```

## Phrase Accuracy

## Perceived Listening Effort Ratings

### 2023-10-26 ASHA 2023 Analysis/data_analysis.Rmd
Created space for describing listener demographics
This section is wip. Just loaded in the listener data for now. Will work on creating demographic tables later.

@@
 ---
-title: "Data Analysis"
+title: "ASHA 2023 Listening Effort Data Analysis"
 author: "Micah E. Hirsch"
@@
 
-```{r}
-summary(cars)
+# Purpose
+
+```{r, include = F, warning = F, message = F}
+
+# Load Needed Packages
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Need remotes package to install gazer and saccades
+library(remotes) # install.packages("remotes")
+library(gazer) # remotes::install_github("dmirman/gazer")
+library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
+library(zoo) # install.packages("zoo")
+library(ggpubr)
+library(gt) # install.packages("gt")
+
 ```
@@
 
+First, we will present the demographics of our listeners To date, we recruited 20 listeners to participate in the study.
+Three listeners were removed from the analysis.
+One listener decided to withdraw from the study, so their data was removed from the database.
+For the other two participants, pupillometry data was either not recorded or lost due to instrumental issues.
+Therefore, for this analysis, we have 17 participants.
+The descriptive information for these participants are presented below.
+
 ```{r}
-plot(pressure)
+
+# Load in dataset for listener demographics
+
+listener_demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv")
+
 ```

### 2023-10-26 ASHA 2023 Analysis/data_analysis.Rmd
Started counts for accurate vs inaccurate responses
need to investigate why number of trials does not equal number of expected trials

@@
 
+```{r}
+
+# Load in cleaned data
+
+data <- rio::import("Cleaned Data/cleaned_data.csv")
+
+# Selecting relevant variables to explore phrase accuracy and perceived listening effort
+data1 <- data %>%
+  dplyr::select(subject, trial, speaker, effort_rating, rep_acc) %>%
+  dplyr::distinct()
+
+# Calculating repetition accuracy descriptive data
+
+rep_acc <- data1 %>%
+  dplyr::group_by(speaker, rep_acc) %>%
+  dplyr::summarize(n = n())
+
+```
+
 ## Perceived Listening Effort Ratings

### 2023-10-26 ASHA 2023 Analysis/transcription_data_preprocessing.R
Updated transcription script
identified trials with missing data.

@@ -42,7 +42,13 @@ transcriptions <- transcriptions %>%
                       names_to = "trial",
                       values_to = "response") %>%
   dplyr::rename(id = ID) %>%
-  dplyr::mutate(trial = as.numeric(trial))
+  # formatting trial and response variables
+  dplyr::mutate(trial = as.numeric(trial),
+                response = tolower(response),
+                response = trimws(response, "both")) %>%
+  # Removing first 5 trials since these are practice
+  dplyr::filter(!trial %in% (1:5))
+
 
 # Calculating phrase accuracy 
 
@@ -58,7 +64,10 @@ targets <- pupil %>%
   dplyr::rename(id = subject,
                 target = targetphrase) %>%
   dplyr::select(id, trial, target) %>%
-  distinct()
+  distinct() %>%
+  # Added this code to ensure targets variable is formatted correctly
+  dplyr::mutate(target = tolower(target),
+                target = trimws(target, "both"))
   
 ## Merging transcription and target dfs
 
@@ -80,10 +89,19 @@ phrase_acc <- autoscore::autoscore(
   double_letter_rule = T) %>%
   dplyr::rename(., correct_words = autoscore)
 
+
 # Creating dichotomous phrase repetition accuracy variable
 
+## There are some trials with missing transcription data. The lab RAs coded these 
+## transcriptions with missing data. Therefore, this section of code is going to check to see
+## if the response variable contains "missing data". If it does, then repetition
+## accuracy will be set to NA. Otherwise, repetition accuracy will be determined 
+## as "accurate" or "inaccurate" for the remaining trials.
+
+
 phrase_acc <- phrase_acc %>%
-  dplyr::mutate(rep_acc = ifelse(target_number == correct_words, "accurate", "inaccurate")) %>%
+  dplyr::mutate(rep_acc = ifelse(response == "missing data", NA, 
+                                 ifelse(target_number == correct_words, "accurate", "inaccurate"))) %>%
   dplyr::rename(subject = id,
                 targetphrase = target)
 

### 2023-10-26 ASHA 2023 Analysis/data_analysis.Rmd
Counted number of accurate responses

@@
   dplyr::select(subject, trial, speaker, effort_rating, rep_acc) %>%
-  dplyr::distinct()
+  dplyr::distinct() %>%
+  # removing trials with missing rep accuracy: 3 trials removed
+  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
+  dplyr::filter(!is.na(rep_acc))
 

### 2023-10-27 ASHA 2023 Analysis/data_analysis.Rmd
small change to analysis script
calculating mean speaker intel and repetition accuracy

--- 2023-10-26 ASHA 2023 Analysis/data_analysis.Rmd
@@
 
-## Phrase Accuracy
+## Speaker Intelligibility and Phrase Accuracy
+
+### Mean Speaker Intelligibility
+
+Across all trials in the study, the intelligibility level for each of the speaker is reported below.
+Across listeners, the mean intelligibility for the ALS speaker was 81.08% with a standard deviation of 3.86.
+For the control speaker, the mean intelligibility level was 93.62% with a standard deviation of 2.90.
+
+```{r}
+
+# Load in cleaned data
+
+data <- rio::import("Cleaned Data/cleaned_data.csv")
+
+intel_df <- data %>%
+  dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) %>%
+  dplyr::distinct() %>%
+  # removing trials with missing rep accuracy: 3 trials removed
+  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
+  dplyr::filter(!is.na(rep_acc))
+
+intel_byListener <- intel_df %>%
+  dplyr::group_by(speaker, subject) %>%
+  dplyr::summarize(Intelligibility = sum(correct_words)/sum(target_number)*100)
+
+intel_byListener %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize("Mean Intelligibility" = mean(Intelligibility), SD = sd(Intelligibility))
+
+```
+
+### Phrase Accuracy
+
+```{r}
+
+# Calculating repetition accuracy descriptive data
+
+rep_acc <- intel_df %>%
+  dplyr::group_by(speaker, rep_acc) %>%
+  dplyr::summarize(n = n()) %>%
+  tidyr::pivot_wider(names_from = rep_acc,
+                     values_from = n) %>%
+  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate))
+
+```
 

### 2023-10-27 ASHA 2023 Analysis/data_analysis.Rmd
streamline code for data import

@@
 
-# Load in cleaned data
+# Load in cleaned data for pupil and perceived listening effort data
 
-data <- rio::import("Cleaned Data/cleaned_data.csv")
+data <- rio::import("Cleaned Data/cleaned_data.csv") %>%
+  # removing trials with missing rep accuracy: 3 trials removed
+  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
+  dplyr::filter(!is.na(rep_acc))
 
@@
   dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) %>%
-  dplyr::distinct() %>%
-  # removing trials with missing rep accuracy: 3 trials removed
-  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
-  dplyr::filter(!is.na(rep_acc))
+  dplyr::distinct()
 
@@
                      values_from = n) %>%
-  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate))
+  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate)*100)
+
+rep_acc
 
@@
 
-## Perceived Listening Effort Ratings
+# Perceived Listening Effort Ratings
+
+## Descriptives
+
+```{r}
+
+perceived_effort_df <-
+
+```

### 2023-10-27 ASHA 2023 Analysis/data_analysis.Rmd
Preliminary analysis on perceived effort data

@@
 
+The purpose of this document is to run preliminary analyses and visualizations for this project.
+These results were prepared specifically for a poster presentation at the the 2023 Annual American Speech Language Hearing Association Convention being held in Boston, MA in November 2023.
+
 ```{r, include = F, warning = F, message = F}
@@
 library(gt) # install.packages("gt")
+library(glmmTMB) # install.packages("glmmTMB")
+library(performance) # install.packages("performance")
 
@@
 
-perceived_effort_df <-
+perceived_effort_df <- data %>%
+  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code) %>%
+  dplyr::distinct() %>%
+  # Creating a trial variable that codes trial 6 (i.e. the first experimental trial) as 0.
+  dplyr::mutate(trial_c = trial - 6) %>%
+  # Releveling speaker variable
+  dplyr::mutate(speaker = as.factor(speaker),
+                speaker = fct_relevel(speaker, "Control", "ALS")) %>%
+  # For this particular analysis, we are just looking at results for accurate repetition
+  # However, the final project will look at rep_acc too
+  dplyr::filter(rep_acc == "accurate")
+
+per_effort_des <- perceived_effort_df %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize(per_effort = mean(effort_rating), 
+                   sd = sd(effort_rating, na.rm = T),
+                   se = sd/sqrt(n()))
+
+per_effort_des
 
 ```
+
+### Visualization
+
+```{r}
+
+per_effort_des %>%
+  ggplot() +
+   aes(x = speaker,
+      y = per_effort,
+      color = speaker,
+      fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.6) +
+  geom_errorbar(aes(x = speaker, ymin = per_effort - se, ymax = per_effort + se), width = 0.4) +
+  labs(x = "Speaker", y = "Perceived Listening Effort Rating") +
+  theme_bw() +
+  theme(legend.position = "none", aspect.ratio = 1)
+
+```
+
+## Perceived Listening Effort Model
+
+### Fully Unconditional Model
+
+```{r}
+
+m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = perceived_effort_df)
+
+summary(m0)
+
+performance::icc(m0)
+
+```
+
+## Model 1
+
+Adding the centered trial random slope
+
+```{r}
+
+m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+
+summary(m1)
+
+anova(m0, m1)
+
+```
+## Model 2
+
+Adding trial as a fixed effect
+
+```{r}
+
+m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+
+summary(m2)
+
+anova(m1, m2)
+
+```
+
+## Model 3
+
+```{r}
+
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+
+summary(m3)
+
+anova(m2, m3)
+
+```
+
+## Final Model
+
+```{r}
+
+per_effort_model_final <- glmmTMB(effort_rating ~ speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+
+summary(per_effort_model_final)
+
+# Removing unneeded items from the environment
+
+rm(m0, m1, m2, m3)
+
+```
+
+# Pupil Dilation
+
+## Descriptives and Visualization
+
+```{r}
+
+```

### 2023-10-27 ASHA 2023 Analysis/data_analysis.Rmd
Streamlining code

@@
   dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
-  dplyr::filter(!is.na(rep_acc))
+  dplyr::filter(!is.na(rep_acc)) %>%
+  # Releveling speaker variable
+  dplyr::mutate(speaker = as.factor(speaker),
+                speaker = fct_relevel(speaker, "Control", "ALS")) %>%
+  # Creating a trial variable that codes trial 6 (i.e. the first experimental trial) as 0.
+  dplyr::mutate(trial_c = trial - 6)
 
@@
   dplyr::distinct() %>%
-  # Creating a trial variable that codes trial 6 (i.e. the first experimental trial) as 0.
-  dplyr::mutate(trial_c = trial - 6) %>%
-  # Releveling speaker variable
-  dplyr::mutate(speaker = as.factor(speaker),
-                speaker = fct_relevel(speaker, "Control", "ALS")) %>%
   # For this particular analysis, we are just looking at results for accurate repetition

### 2023-10-27 ASHA 2023 Analysis/data_analysis.Rmd
troubleshooting pupil dilation visualization and analysis

@@
 
+data %>%
+  # For this analysis, we are only looking at accurate repetition trials
+  dplyr::filter(rep_acc == "accurate") %>%
+  dplyr::group_by(timebins, speaker) %>%
+  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
+                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
+  ggplot() +
+  aes(x = timebins,
+      y = pupil_dilation,
+      color = speaker,
+      fill = speaker) +
+  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
+  geom_line() +
+  coord_cartesian(xlim = c(0,5000), ylim = c(-100, 200)) +
+  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
+  theme_bw() + 
+  theme(legend.position = "bottom", aspect.ratio = 1)
+
 ```
+
+# Fully Unconditional Model
+
+```{r}
+
+# Specifying the region of interest for analysis (500 ms to 1000 ms after relative offset of the trial)
+
+pupil_roi <- data %>%
+  # For this analysis, we are only analyzing pupil dilation for accurate responses
+  dplyr::filter(rep_acc == "accurate") %>%
+  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
+  dplyr::filter(timebins >= 500) %>%
+  dplyr::group_by(subject, trial) %>%
+  dplyr::filter(timebins <= max(timebins) - 2000) %>%
+  ungroup() %>%
+  # For the purpose of modeling, we are centering time at 500 ms (i.e. the beginning ROI time)
+  dplyr::mutate(time_c = timebins - 500)
+
+# Creating polynomial variables for model
+
+pupil_roi <- gazer::code_poly(pupil_roi, predictor = "time_c", 
+                              poly.order = 3, orthogonal = T, draw.poly = F)
+
+# specifying unconditional growth model for pupil data
+
+m0 <- glmmTMB(pupil.binned ~ poly1 + poly2 + poly3 + 
+                (poly1 + poly2 + poly3|subject) + 
+                (poly1 + poly2 +poly3|code), 
+              data = pupil_roi)
+
+summary(m0)
+
+performance::icc(m0)
+
+```

### 2023-10-28 ASHA 2023 Analysis/data_analysis.Rmd
exporting roi df
Ran initial analyses for pupil dilation data

@@
 
+First, we ran an unconditional growth model.
+I first specified a model with the orthogonal polynomials added as fixed effects and subject (i.e.
+listener) and code (i.e.
+phrase) as random effects.
+I also added the orthogonal polynomials as random slopes.
+I also used the AR(1) covariance structure to specifiy the random effect structure.
+I was not able to run the model.
+Therefore, I attempted to respecify the random effect structure by removing the AR(1) specification.
+This led to a model with singular fit.
+In response to this, I decided to remove code (i.e.) from the random effect structure in order to simplify it.
+
 ```{r}
@@
   # For the purpose of modeling, we are centering time at 500 ms (i.e. the beginning ROI time)
-  dplyr::mutate(time_c = timebins - 500)
+  dplyr::mutate(time_c = timebins - 500) %>%
+  # Rescaling pupil size 
+  dplyr::group_by(subject, trial) %>%
+  dplyr::mutate(norm_pupil = as.numeric(scale(pupil.binned))) %>%
+  dplyr::ungroup()
 
-# Creating polynomial variables for model
+# Polynomials
 
-pupil_roi <- gazer::code_poly(pupil_roi, predictor = "time_c", 
-                              poly.order = 3, orthogonal = T, draw.poly = F)
+pupil_roi <- gazer::code_poly(pupil_roi, predictor = "timebins", poly.order = 3, orthogonal = T, draw.poly = F)
+
+# Exporting region of interest df
+
+rio::export(pupil_roi, "Cleaned Data/pupil_analysis_roi.csv")
+
+## I kept running into issues using glmmTMB, so I switched to lme4
+
+library(lme4)
+library(lmerTest)
 
@@
 
-m0 <- glmmTMB(pupil.binned ~ poly1 + poly2 + poly3 + 
-                (poly1 + poly2 + poly3|subject) + 
-                (poly1 + poly2 +poly3|code), 
+m0 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 +
+                (poly1 + poly2 + poly3|subject), 
               data = pupil_roi)
@@
 ```
+
+# Model 1
+
+```{r}
+
+m1 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
+                (poly1 + poly2 + poly3|subject), 
+              data = pupil_roi)
+
+summary(m1)
+
+anova(m0, m1)
+
+```
+
+## Model 2
+
+```{r}
+
+# Model 2 initially failed to converge, so adding a control statement for bobyqa optimizer
+
+control_bobyqa <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e9))
+
+m2 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
+             poly1*speaker +
+             (poly1 + poly2 + poly3|subject), 
+              data = pupil_roi, control = control_bobyqa)
+
+summary(m2)
+
+anova(m1, m2)
+
+```
+## Model 3
+
+```{r}
+
+m3 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
+             poly1*speaker + poly2*speaker +
+             (poly1 + poly2 + poly3|subject), 
+              data = pupil_roi, control = control_bobyqa)
+
+summary(m3)
+
+anova(m2, m3)
+
+```
+## Model 4
+
+```{r}
+
+m4 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
+             poly1*speaker + poly2*speaker + poly3*speaker +
+             (poly1 + poly2 + poly3|subject), data = pupil_roi, 
+           control = control_bobyqa)
+
+summary(m4)
+
+anova(m3, m4)
+
+```

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
Calculated participant demographic information

@@
 author: "Micah E. Hirsch"
-date: "2023-10-26"
+date: "2023-10-29"
 output: html_document
@@
 
+Any questions or comments regarding this project or analysis may be directed to Micah Hirsch, mhirsch@fsu.edu.
+
 ```{r, include = F, warning = F, message = F}
@@
 library(zoo) # install.packages("zoo")
-library(ggpubr)
+library(ggpubr) # installl.packages("ggpubr")
 library(gt) # install.packages("gt")
@@
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 listener_demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv")
+
+# Extracting listener gender data
+
+gender <- listener_demo %>%
+  dplyr::select(gender) %>%
+  dplyr::group_by(gender) %>%
+  dplyr::summarize(Frequency = n()) %>%
+  dplyr::rename(Category = "gender") %>%
+  dplyr::mutate(Group = "Gender",
+                Category = fct_relevel(Category, c("woman", "man", "nonbinary"))) %>%
+  dplyr::arrange(Category)
+
+## Extracting listener race data
+
+race <- listener_demo %>%
+  dplyr::select(race) %>%
+  dplyr::group_by(race) %>%
+  dplyr::summarize(Frequency = n()) %>%
+  dplyr::rename(Category = "race") %>%
+  dplyr::mutate(Group = "Race",
+                Category = fct_relevel(Category, c("white/Caucasian", "Biracial or Multiracial", 
+                                                   "race not listed"))) %>%
+  dplyr::arrange(Category)
+
+## Extracting listener ethnicity data
+
+ethnicity <- listener_demo %>%
+  dplyr::select(ethnicity) %>%
+  dplyr::group_by(ethnicity) %>%
+  dplyr::summarize(Frequency = n()) %>%
+  dplyr::rename(Category = "ethnicity") %>%
+  dplyr::mutate(Group = "Ethnicity",
+                Category = fct_relevel(Category, c("Hispanic/Latino(a/e)", 
+                                                   "Not Hispanic/Latino(a/e)"))) %>%
+  dplyr::arrange(Category)
+
+# Creating Demographic Table 1
+
+dplyr::full_join(gender, race) %>%
+  dplyr::full_join(., ethnicity) %>%
+  # Initiating a gt table
+  gt::gt(rowname_col = "Category",
+         groupname_col = "Group")
+
+```
+
+As indicated in the table above, our participants were primarily white/Caucasian, women, and not Hispanic or Latino(a/e).
+All participants reported no history of a speech, language, or hearing disorder.
+All participants were fluent English speakers.
+However, two indicated their native language was different than American English.
+These languages were Spanish (n=1) and Turkish (n=1).
+
+```{r, warning = F, message = F}
+
+listener_demo %>%
+  dplyr::select(c(age)) %>%
+  dplyr::summarize(Mean = mean(age), SD = sd(age),
+                   Min = min(age), Max = max(age)) %>%
+  dplyr::mutate(Category = "Age") %>%
+  dplyr::relocate(Category, .before = Mean) %>%
+  gt::gt() %>%
+  gt::fmt_number(columns = c(Mean, SD),
+                decimals = 2) %>%
+  gt::cols_label(Category = "")
+
+```
+
+The mean age of the listeners is 26.53 years old. Listener age ranged from 18 to 37 years old.
+
+```{r, include = F, warning = F, message = F}
+
+# Removing unneeded items from the environment
+
+rm(ethnicity, gender, listener_demo, race)
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
streamlined code for phrase accuracy descriptives

@@
 
+# Selecting variables needed to calculate mean speaker intelligibility
 intel_df <- data %>%
@@
 
+# Calculating intelligibility for the control and ALS speaker for each listener
 intel_byListener <- intel_df %>%
@@
   dplyr::group_by(speaker) %>%
-  dplyr::summarize("Mean Intelligibility" = mean(Intelligibility), SD = sd(Intelligibility))
+  dplyr::summarize("Mean Intelligibility" = mean(Intelligibility), SD = sd(Intelligibility)) %>%
+  gt::gt() %>%
+  gt::fmt_number(columns = c("Mean Intelligibility", SD),
+                 decimals = 2) %>%
+  gt::cols_label(speaker = "Speaker")
 
@@
 
-```{r}
+For this analysis, we are only looking at the trials that listeners were able to accurately recognize (e.g.
+100% intelligible to the listener).
+Therefore, we are looking at the proportion of trials for each speaker that listeners were able to accurately recognize compared to the total number of trials.
+Although our ALS speaker's intelligibility was fairly high, only about 49% of all the ALS speaker's trials were accurately recognized across listeners.
+
+```{r, warning = F, message = F}
 
@@
 
-rep_acc <- intel_df %>%
+intel_df %>%
   dplyr::group_by(speaker, rep_acc) %>%
@@
                      values_from = n) %>%
-  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate)*100)
+  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate)*100) %>%
+  dplyr::ungroup() %>%
+  gt::gt(rowname_col = "speaker") %>%
+  gt::fmt_number(columns = percent_accurate,
+                 decimals = 2) %>%
+  gt::cols_label(accurate = "Accurate",
+                 inaccurate = "Inaccurate",
+                 percent_accurate = "Percent Accurate")
 
-rep_acc
+```
+```{r, include = F, warning = F, message = F}
+
+# removing unneeded items from the environment
+
+rm(intel_byListener, intel_df)
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
descriptives for perceived effort ratings
streamlined code

@@
 
-per_effort_des
+per_effort_des %>%
+  gt::gt() %>%
+  gt::fmt_number(columns = per_effort:se,
+                 decimals = 2) %>%
+  gt::cols_label(speaker = "Speaker",
+                  per_effort = "Perceived Effort",
+                  sd = "SD",
+                  se = "SE")
 
@@
 
-### Visualization
+As reported in the table above and shown in the plot below, the mean perceived effort ratings for the ALS speaker is about 2.13 points higher than the control speaker.
+This is reflecting the difference in mean perceived effort ratings between the ALS and control speaker, even when the phrase was intelligible to the listener.
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
updated and exported perceived listening effort figure

@@
 
-```{r}
+```{r, warning = F, message = F}
 
+# Creating a color palette vector for plots
+my_pal <- c("#CD9FCC", "#7E2E84")
+
+# Plotting the mean perceived effort ratings for the ALS and control speaker
 per_effort_des %>%
@@
       y = per_effort,
+      group = speaker,
       color = speaker,
@@
   geom_errorbar(aes(x = speaker, ymin = per_effort - se, ymax = per_effort + se), width = 0.4) +
-  labs(x = "Speaker", y = "Perceived Listening Effort Rating") +
+  labs(x = "Speaker", y = "Mean Perceived Listening Effort Rating") +
   theme_bw() +
-  theme(legend.position = "none", aspect.ratio = 1)
+  theme(legend.position = "none", aspect.ratio = 1) +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal)
+
+# Saving the ggplot
+
+ggsave("Plots/Perceived_effort.png", plot = last_plot(), height = 5, width = 5, units = "in")
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
updated perceived listening effort analysis
streamlined code and exported final model results

@@
 library(performance) # install.packages("performance")
+library(sjPlot) # install.packages("sjPlot")
 
@@
 perceived_effort_df <- data %>%
-  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code) %>%
+  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code, trial_c) %>%
   dplyr::distinct() %>%
@@
 
-```{r}
+To analyze the perceived effort ratings, we are going to use linear mixed-effects models.
+To start, we are going to specify our fully unconditional model.
+In this model, we included random intercepts for subject (i.e.
+our listeners) and code (i.e.
+the different phrases).
+No fixed effects have been added to the model.
+We also calculated ICC from this model.
+
+Both the conditional and unconditional ICC is 0.99.
+
+```{r, warning = F, message = F}
 
@@
 
-## Model 1
+### Model 1
 
-Adding the centered trial random slope
+Now we will add trial number (i.e.
+trial_c) as a random slope for both subject and code.
+A likelihood ratio test was also ran to test whether model 1 provided a better fit to the data than the unconditional model.
 
@@
 ```
-## Model 2
 
-Adding trial as a fixed effect
+The likelihood ratio test indicated that model 1 provided a better fit to the data than the unconditional model (chi-sq(4) = 227.79, p <0.001). Therefore, we will retain model 1.
+
+### Model 2
+
+To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to mdoel 1
 
@@
 
-## Model 3
+As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1)m p = 0.52).
+This indicates that trial order is not an important predictor for perceived listening effort ratings and that ratings for the speakers were fairly consistent across trials.
 
-```{r}
+### Model 3
+
+In model 3, we added the speaker fixed effect to test whether the mean difference in perceived listening effort rating is significantly difference for the ALS speaker compared to the control speaker.
+A likelihood ratio test was also conducted to test whether model 3 provides a better fit to the data compared to model 2.
+
+```{r, warning = F, message = F}
 
@@
 
+Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1), p <0.001).
+The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.
+
+### Model 4
+
+To test whether there are any potential differences in perceived effort ratings across trials, we are testing whether adding an interaction between trial_c and speaker improves model fit.
+Again, a likelihood ratio test was conducted to see if model 4 provides a better fit to the data compared to model 3.
+
+```{r}
+
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+
+summary(m4)
+
+anova(m3, m4)
+
+```
+
+Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = 0.70).
+The fixed effect estimate for the interaction term is also not significant.
+This indicates that effort ratings across the trials for the ALS speaker were not significantly different from each other.
+
 ## Final Model
+
+Based on our model building, we will now specify our final model.
+Since the trial fixed effect as well as the interaction between trial and speaker were not significant, nor did they improve model fit, they were removed from the final model.
+All other output remained the same.
 
@@
 
-# Removing unneeded items from the environment
+sjPlot::tab_model(per_effort_model_final, file = "Final Model Outputs/perceived_effort_model.html")
+```
 
-rm(m0, m1, m2, m3)
-
-```
+The final model indicates that the perceived effort ratings for the ALS speaker is about 2.35 points higher than they control speaker.
+This is consistent with the results shown in the descriptive statistics.
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
Updated pupil dilation visualization

@@
 
-## Final Model
+### Final Model
 
@@
 
+```{r, include = F, warning = F, message = F}
+
+# Remove unneeded items from the environment
+
+rm(m0, m1, m2, m3, m4, per_effort_des, perceived_effort_df, per_effort_model_final, model_0, model0)
+
+```
+
 # Pupil Dilation
 
-## Descriptives and Visualization
+## Descriptives
+
+The ALS speaker has a slower speech rate compared to the control speaker.
+Therefore, we are going to calculate the average trial legnth to see on average, how different the lengths were for ALS trials and control speaker trials.
+We will need this information to be able to mark the average phrase offset time and average end point for the analysis region of interest for visualization purposes.
 
@@
 data %>%
-  # For this analysis, we are only looking at accurate repetition trials
-  dplyr::filter(rep_acc == "accurate") %>%
-  dplyr::group_by(timebins, speaker) %>%
-  dplyr::summarize(pupil_dilation = mean(pupil.binned, na.rm = T),
-                   pupil_se = sd(pupil.binned, na.rm = T)/sqrt(n())) %>%
-  ggplot() +
-  aes(x = timebins,
-      y = pupil_dilation,
-      color = speaker,
-      fill = speaker) +
-  geom_ribbon(aes(ymin = pupil_dilation - pupil_se, ymax = pupil_dilation + pupil_se), alpha = .3) +
-  geom_line() +
-  coord_cartesian(xlim = c(0,5000), ylim = c(-100, 200)) +
-  labs(x = "Time (ms)", y = "Pupil Dilation (arbitrary units)", fill = "Speaker", color = "Speaker") +
-  theme_bw() + 
-  theme(legend.position = "bottom", aspect.ratio = 1)
+  # Selecting relevant parameters
+  dplyr::select(subject, trial, speaker, timebins, code) %>%
+  # Filtering out time before start of phrase
+  dplyr::filter(timebins >= 0) %>%
+  dplyr::group_by(speaker, code) %>%
+  dplyr::summarize(length = max(timebins)) %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize(av_length = mean(length),
+                   av_offset = av_length - 3000,
+                   av_end_roi = av_length - 2000) %>%
+  gt::gt() %>%
+  gt::cols_label(speaker = "Speaker",
+                 av_length = "Average Trial Length (ms)",
+                 av_offset = "Average Phrase Length (ms)",
+                 av_end_roi = "Average Endpoint for ROI (ms)")
 
@@
 
-# Fully Unconditional Model
+Average trial length includes the phrase presentation plus the 3000 ms retention period.
+Average phrase legnth is the average length of the recording.
+Average Endpoint of the Analysis ROI is the 1000 ms after the average phrase endpoint (calculated this for data visualization purposes).
+On average, the phrase length (phrase presentation), was about 2461 ms longer for the ALS trials than the control speaker trials.
+
+### Visualization
+
+This is the smoothed pupil dilation tracking for the Control Speaker.
+The solid line indicates the start of the phrase.
+The dotted line indicates the average phrase offset.
+The grey box represents the region of interest for analysis.
+
+```{r}
+
+control <- data %>%
+  dplyr::filter(rep_acc == "accurate") %>%
+  dplyr::filter(speaker == "Control") %>%
+  ggplot() +
+  aes(x = timebins,
+      y = pupil.binned,
+      color = speaker) +
+  geom_smooth() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1908, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2908, ymin = -50, ymax = 150, alpha = 0.2) +
+  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
+  labs(title = "Control Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
+  theme_bw() +
+  scale_color_manual(values = my_pal[1]) +
+  theme(legend.position = "none", aspect.ratio = 1)
+
+control
+
+```
+
+This is the smoothed pupil dilation tracking for the ALS Speaker.
+The solid line indicates the start of the phrase.
+The dotted line indicates the average phrase offset.
+The grey box represents the region of interest for analysis.
+
+```{r}
+
+als <- data %>%
+  dplyr::filter(rep_acc == "accurate") %>%
+  dplyr::filter(speaker == "ALS") %>%
+  ggplot() +
+  aes(x = timebins,
+      y = pupil.binned,
+      color = speaker) +
+  geom_smooth() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 4369, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 5369, ymin = -50, ymax = 150, alpha = 0.2) +
+  coord_cartesian(xlim = c(0,7369), ylim = c(-100, 200)) +
+  labs(title = "ALS Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
+  theme_bw() +
+  scale_color_manual(values = my_pal[2]) +
+  theme(legend.position = "none", aspect.ratio = 1)
+
+als
+
+```
+
+```{r, include = F, warning = F, message = F}
+
+pupil_plot <- ggarrange(control, als, ncol = 2)
+
+ggsave("Plots/Pupil_dilation.png", plot = pupil_plot, width = 10, height = 5, units = "in")
+
+```
+
+## Pupil Dilation Analysis
+
+### Fully Unconditional Model
 
 First, we ran an unconditional growth model.
-I first specified a model with the orthogonal polynomials added as fixed effects and subject (i.e.
+In the intial model, I first specified a model with the orthogonal polynomials added as fixed effects and subject (i.e.
 listener) and code (i.e.
-phrase) as random effects.
+phrase) as random intercepts.
 I also added the orthogonal polynomials as random slopes.
 I also used the AR(1) covariance structure to specifiy the random effect structure.
-I was not able to run the model.
-Therefore, I attempted to respecify the random effect structure by removing the AR(1) specification.
+However, I was not able to run the model and ran into convergence issues.
+
+Therefore, I decided to respecify the random effect structure by removing the AR(1) specification.
 This led to a model with singular fit.
-In response to this, I decided to remove code (i.e.) from the random effect structure in order to simplify it.
+In response to this, I decided to remove code (i.e.phrase) from the random effect structure in order to simplify it.
+
+More details about how I respecified the model can be seen in the code block.
 
@@
   dplyr::filter(timebins <= max(timebins) - 2000) %>%
-  ungroup() %>%
-  # For the purpose of modeling, we are centering time at 500 ms (i.e. the beginning ROI time)
-  dplyr::mutate(time_c = timebins - 500) %>%
-  # Rescaling pupil size 
-  dplyr::group_by(subject, trial) %>%
-  dplyr::mutate(norm_pupil = as.numeric(scale(pupil.binned))) %>%
-  dplyr::ungroup()
+  ungroup()
 
-# Polynomials
+# Adding linear, quadratic, and cubic orthogonal polynomials to the df
 
@@
 
-# Exporting region of interest df
+# Exporting region of interest df (this can be used to )
 

### 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
updated pupil analysis code

@@
 
+rm(als, control)
+
 ```
@@
 
-```{r}
+The adjusted ICC for the unconditional growth model is .109. The unadjusted ICC for the conditional growth model is .107.
+
+```{r, warning = F, message = F}
 
@@
 
-# Model 1
+### Model 1
 
-```{r}
+In the next model, we are adding a fixed effect for speaker to test whether on average, pupil dilation is greater for the ALS speaker compared to the control speaker.
+A likelihood ratio test was also conducted to test whether model 1 provides a better fit to the data than the unconditional growth model.
+
+```{r, warning = F, message = F}
 
@@
 
-## Model 2
+As indicated in the model output, when listening to the ALS speaker, on average, listeners had greater pupil dilation compared to when they were listening to the control speaker.
+The likelihood ratio test also indicates model 1 is a better fit to the data compared to the unconditional model (chi-sq(1) = 353.67, p <0.001).
 
-```{r}
+### Model 2
+
+In model two, the interaction between poly1 and speaker was added to the model as a fixed effect to test whether linear growth was greater for the ALS speaker compared to the control speaker.
+A likelihood ratio test was also conducted to test if model 2 provides a better fit to the data compared to model 1.
+
+Note: The initial run of this model did not converge, so an optimizer control was added to the model. This optimizer will remain in each subsequent model.
+
+```{r, warning = F, message = F}
 
@@
 ```
-## Model 3
 
-```{r}
+Based on the model output, the interaction between poly1 and speaker is significant, indicating that, across listeners, linear growth in pupil dilation while listening to the ALS speaker was significantly greater than when listeners were listening to the control speaker.
+The likelihood ratio test also indicated that model 2 provided a better fit to the data than model 1 (chi-sq(1) = 649.22, p <0.001).
+
+### Model 3
+
+In model three, the interaction between the quadratic growth term and speaker was added to the model.
+This tests whether the quadratic shape of pupil dilation is significantly different between the two speakers.
+A likelihood ratio test was also conducted to test whether model 3 provides a better fit to the data compared to model 2.
+
+```{r, warning = F, message = F}
 
@@
 ```
-## Model 4
+Based on the model output, the interaction between the quadratic growth term and speaker improved the model fit (ch-sq(1) = 147.72, p<0.001).
+This indicates, on average, the shape of the listeners' pupil dilation is different while listening to the ALS speaker vs the control speaker.
 
-```{r}
+### Model 4
+
+In model for, we added the interaction between the cubic growth term and speaker.
+As with the other models, a likelihood ratio test was conducted to test whether model 4 provides a better fit to the data compared to model 3.
+
+```{r, message = F, warning = F}
 
@@
 ```
+Based on the model output, model 4 provides a better fit to the data compared to model 3 (chi-sq(1) = 9.25, P = 0.002).
+What is interesting is that the only fixed effects that are significant in this model are the main effect of the cubic growth term and the interaction between the cubic growth term and speaker, indicating the cubic pupil dilation growth is significantly different between these two speakers.
+
+```{r, include = F, message = F, warning = F}
+
+# Saving model 4 output
+
+sjPlot::tab_model(m1, m2, m3, m4, file = "Final Model Outputs/pupil_dilation_model.html")
+
+# removing unneeded items from the environment
+
+rm(m0, m1, m2, m3, m4, pupil_plot, pupil_roi, data, control_bobyqa, my_pal)
+
+```

### 2023-10-30 ASHA 2023 Analysis/data_analysis.Rmd
started alternative analysis-code not working

@@
 date: "2023-10-29"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
 ---
@@
 ```
+
+# Alternative Analysis
+
+Since the trials for the ALS speaker are much longer than the control speaker's trials, we are going to conduct an alternative analysis.
+First, we are going to resample the ALS speaker's trials so the length of the trials are comparable to the control speaker's trials.
+Then, we will conduct the mixed-effects models on the resampled data.
+
+```{r}
+
+# Rescaling the ALS speaker's trials to be more consistent with the Control speaker.
+
+## Bin.width
+bin.width <- 1.3
+
+ALS.binned <- data %>%
+  dplyr::filter(speaker == "ALS") %>%
+  mutate(bins = floor(timebins/bin.width)*bin.width) %>%
+  dplyr::group_by(subject, trial_c, speaker, bins, effort_rating,
+                  code, counterbalance) %>%
+  dplyr::summarize(pupil_dilation = mean(pupil.binned)) %>%
+  dplyr::ungroup()
+
+Control <- data %>%
+  dplyr::filter(speaker == "Control")
+
+  # Selecting relevant parameters
+  dplyr::select(subject, trial, speaker, timebins, code) %>%
+  dplyr::distinct() %>%
+  # Filtering out time before start of phrase
+  dplyr::filter(timebins >= 0) %>%
+  dplyr::group_by(speaker, code) %>%
+  dplyr::summarize(length = max(timebins)) %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize(av_length = mean(length),
+                   av_offset = av_length - 3000,
+                   av_end_roi = av_length - 2000) %>%
+  gt::gt() %>%
+  gt::cols_label(speaker = "Speaker",
+                 av_length = "Average Trial Length (ms)",
+                 av_offset = "Average Phrase Length (ms)",
+                 av_end_roi = "Average Endpoint for ROI (ms)")
+
+```
+
+```{r}
+data %>%
+  # For this analysis, we are only analyzing pupil dilation for accurate responses
+  dplyr::filter(rep_acc == "accurate") %>%
+  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
+  dplyr::filter(timebins >= 500) %>%
+  dplyr::group_by(subject, trial) %>%
+  dplyr::filter(timebins <= max(timebins) - 2000) %>%
+  ungroup()
+```

### 2023-10-30 ASHA 2023 Analysis/data_analysis.Rmd
Completed dilation reanalysis

--- 2023-10-29 ASHA 2023 Analysis/data_analysis.Rmd
@@
 library(sjPlot) # install.packages("sjPlot")
+library(ggridges) # install.packages("ggridges")
+library(lme4) # install.packages("lme4")
+library(lmerTest) # install.packages("lmerTest")
 
@@
 
+```{r}
+
+peak_pupil_df %>%
+  ggplot() +
+  aes(y = speaker,
+      x = peak_pupil,
+      color = speaker,
+      fill = speaker) +
+  geom_density_ridges(scale = 0.7, alpha = 0.5) +
+  theme_bw() +
+  scale_fill_manual(values = my_pal) +
+  scale_color_manual(values = my_pal)
+
+```
+
 ## Perceived Listening Effort Model
@@
 
-Both the conditional and unconditional ICC is 0.99.
+Both the conditional and unconditional ICC is 0.099.
 
@@
 
-m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
 
@@
 
-To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to mdoel 1
+To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to model 1
 
@@
 
-m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), 
+              data = perceived_effort_df)
 
@@
 
-m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
 
@@
 
-Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1), p <0.001).
+Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1) = 469.74, p <0.001).
 The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.
@@
 
-m4 <- glmmTMB(effort_rating ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
 
@@
 ```
+
+# Alternative Pupil Dilation Analysis
+
+Since the trials for the ALS speaker are much longer than the control speaker's trials, we are going to conduct an exploratory analysis for the pupil dilation measure.
+For each of the listener's trial, we are going to extract the peak pupil dilation during the analysis region and model it as a function of speaker.
+
+## Peak Pupil Dilation Descriptives
+
+```{r}
+
+peak_pupil_df <- pupil_roi %>%
+  dplyr::group_by(subject, trial_c, speaker, code) %>%
+  dplyr::summarize(peak_pupil = max(pupil.binned)) %>%
+  ungroup()
+
+peak_pupil_des <- peak_pupil_df %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize(pupil = mean(peak_pupil), 
+                   sd = sd(peak_pupil, na.rm = T),
+                   se = sd/sqrt(n()))
+
+peak_pupil_des %>%
+  gt::gt() %>%
+  gt::fmt_number(columns = pupil:se,
+                 decimals = 2) %>%
+  gt::cols_label(speaker = "Speaker",
+                  pupil = "Peak Pupil Dilation",
+                  sd = "SD",
+                  se = "SE")
+
+```
+
+### Peak Pupil Dilation Distribution Plot
+
+```{r}
+
+peak_pupil_df %>%
+  ggplot() +
+  aes(y = speaker,
+      x = peak_pupil,
+      color = speaker,
+      fill = speaker) +
+  geom_density_ridges(scale = 0.7, alpha = 0.5) +
+  labs(x = "Mean Peak Pupil Dilation (arbitrary units)", y = "Speaker") +
+  theme_bw() +
+  theme(legend.position = "none", aspect.ratio = 1) +
+  scale_fill_manual(values = my_pal) +
+  scale_color_manual(values = my_pal)
+
+ggsave("Plots/peak_pupil_dis.png", plot = last_plot(), height = 5, width = 5, units = "in")
+
+```
+### Peak Pupil Dilation Bar Plot
+
+```{r}
+
+peak_pupil_des %>%
+  ggplot() +
+   aes(x = speaker,
+      y = pupil,
+      group = speaker,
+      color = speaker,
+      fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.6) +
+  geom_errorbar(aes(x = speaker, ymin = pupil - se, ymax = pupil + se), width = 0.4) +
+  labs(x = "Speaker", y = "Mean Peak Pupil Dilation (arbitrary units)") +
+  theme_bw() +
+  theme(legend.position = "none", aspect.ratio = 1) +
+  scale_color_manual(values = my_pal) +
+  scale_fill_manual(values = my_pal)
+
+ggsave("Plots/peak_pupil_bar.png", plot = last_plot(), height = 5, width = 5, units = "in")
+
+```
+## Analysis
+
+### Fully Unconditional Model
+
+```{r}
+
+m0 <- lmer(peak_pupil ~ 1 + (1|subject) + (1|code), data = peak_pupil_df)
+
+summary(m0)
+
+performance::icc(m0)
+
+```
+
+### Model 1
+
+```{r}
+
+# Ran into convergence issues, so added the nlminb optimizer to the model.
+# Model had simgular fit issues, so removed the (trial_c|code) random effect.
+
+library(optimx)
+
+m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject), 
+           data = peak_pupil_df, control = lmerControl(
+                           optimizer ='optimx', optCtrl=list(method='nlminb')))
+
+summary(m1)
+
+anova(m0, m1)
+
+```
+
+### Model 2
+
+```{r}
+
+m2 <- lmer(peak_pupil ~ trial_c + (trial_c|subject), 
+           data = peak_pupil_df, control = lmerControl(
+                           optimizer ='optimx', optCtrl=list(method='nlminb')))
+
+summary(m2)
+
+anova(m1, m2)
+
+```
+
+### Model 3
+
+```{r}
+
+m3 <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
+              data = peak_pupil_df, control = lmerControl(
+                           optimizer ='optimx', optCtrl=list(method='nlminb')))
+
+summary(m3)
+
+anova(m2, m3)
+
+```
+
+### Model 4
+
+```{r}
+
+m4 <- lmer(peak_pupil ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) , 
+              data = peak_pupil_df, control = lmerControl(
+                           optimizer ='optimx', optCtrl=list(method='nlminb')))
+
+summary(m4)
+
+anova(m3, m4)
+
+```
+
+#### Final Model
+
+```{r}
+
+final_model <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
+              data = peak_pupil_df, control = lmerControl(
+                           optimizer ='optimx', optCtrl=list(method='nlminb')))
+
+sjPlot::tab_model(final_model, file = "Final Model Outputs/peak_pupil_model.html")
+
+```

### 2023-10-30 ASHA 2023 Analysis/data_analysis.Rmd
streamlined code

@@
 library(lmerTest) # install.packages("lmerTest")
+library(optimx) # install.packages("optimx")
 
@@
 
-library(optimx)
-
 m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject), 

### 2023-10-30 ASHA 2023 Analysis/data_analysis.Rmd
Streamlined code in rmd file

@@
 
-```{r}
+```{r, warning = F, message = F}
 
-peak_pupil_df %>%
+perceived_effort_df %>%
   ggplot() +
   aes(y = speaker,
-      x = peak_pupil,
+      x = effort_rating,
       color = speaker,
@@
   geom_density_ridges(scale = 0.7, alpha = 0.5) +
+  labs(x = "Perceived Listening Effort Rating", y = "Speaker") +
   theme_bw() +
+   theme(legend.position = "none", aspect.ratio = 1) +
   scale_fill_manual(values = my_pal) +
   scale_color_manual(values = my_pal)
+
+# Saving the ggplot
+
+ggsave("Plots/Perceived_effort_dis.png", plot = last_plot(), height = 5, width = 5, units = "in")
 
@@
 
-As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1)m p = 0.52).
+As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1) = 0.412, p = 0.52).
 This indicates that trial order is not an important predictor for perceived listening effort ratings and that ratings for the speakers were fairly consistent across trials.
@@
 
-Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = 0.70).
+Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = 0.14, p= 0.70).
 The fixed effect estimate for the interaction term is also not significant.
@@
 
-per_effort_model_final <- glmmTMB(effort_rating ~ speaker + (trial_c|subject) + (trial_c|code), data = perceived_effort_df)
-
-summary(per_effort_model_final)
+per_effort_model_final <- glmmTMB(effort_rating ~ speaker + (trial_c|subject) + (trial_c|code), 
+                               data = perceived_effort_df)
 
@@
 
-rm(m0, m1, m2, m3, m4, per_effort_des, perceived_effort_df, per_effort_model_final, model_0, model0)
+rm(m0, m1, m2, m3, m4, per_effort_des, perceived_effort_df, per_effort_model_final)
 
@@
 
+```{r, message = F, warning = F}
+
+sjPlot::tab_model(m4, file = "Final Model Outputs/final_pupil_dilation_model.html")
+
+```
+
 ```{r, include = F, message = F, warning = F}
@@
 
-rm(m0, m1, m2, m3, m4, pupil_plot, pupil_roi, data, control_bobyqa, my_pal)
+rm(m0, m1, m2, m3, m4, pupil_plot, control_bobyqa)
 
@@
 
-```{r}
+We first fit a fully unconditional model to the data to calculate ICC. The adjusted and unadjusted ICC is 0.364.
+
+```{r, warning = F, message = F}
+
+# Note: I ran into issues using glmmTMB, so used lme4 instead
 
@@
 
-```{r}
+We added trial as random slopes for both the subject and code random effects.
+The model initially failed to converge, so an optimizer was added to the model.
+Then a singular fit warning was displayed, so the random slope of trial_c was removed from the code random effect.
 
-# Ran into convergence issues, so added the nlminb optimizer to the model.
-# Model had simgular fit issues, so removed the (trial_c|code) random effect.
+```{r, warning = F, message = F}
 
-m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject), 
+# Model failed to converge, so added an optimizer
+# Singular fit so removed random slope for code.
+
+m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject) + (1|code), 
            data = peak_pupil_df, control = lmerControl(
@@
 
+Model 1 provided a better fit to the data (chi-sq(2) = 43.44, p < 0.001).
+
 ### Model 2
 
-```{r}
+Next, in order to control for trial order, trial was added to the model as a fixed effect.
 
-m2 <- lmer(peak_pupil ~ trial_c + (trial_c|subject), 
+```{r, warning = F, message = F}
+
+m2 <- lmer(peak_pupil ~ trial_c + (trial_c|subject) + (1|code), 
            data = peak_pupil_df, control = lmerControl(
@@
 
+Adding trial to the model as a fixed effect significantly improved model fit (chi-sq(1) = p<0.001).
+Based on the fixed effect estimate for trial order, peak pupil dilation decreased approximately -1.23 units on average with each trial.
+
 ### Model 3
 
-```{r}
+Next, we added the fixed effect of speaker to the model to test whether peak pupil dilation was, on average, higher for when listening to the ALS speaker compared to the control speaker.
 
-m3 <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
+```{r, warning = F, message = F}
+
+m3 <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) + (1|code), 
            data = peak_pupil_df, control = lmerControl(
@@
 
+Adding speaker to the model improved the model fit to the data (ch-sq(1), p <0.001).
+Controlling for trial order, listeners' peak pupil dilation, on average, was about 85.22 units higher compared to when they were listening to the control speaker.
+This indicates listeners were exerting more effort while listening to the ALS speaker compared to the control speaker.
+
 ### Model 4
 
-```{r}
+Finally, the interaction between trial and speaker was added to the model to test whether peak pupil dilation for the ALS speaker change at a different rate across trials.
 
-m4 <- lmer(peak_pupil ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) , 
+```{r, warning = F, message = F}
+
+# Ran into convergence problem, so removed the trial_c random slope for code
+
+m4 <- lmer(peak_pupil ~ trial_c + speaker + speaker*trial_c + (trial_c|subject) + (1|code), 
               data = peak_pupil_df, control = lmerControl(
@@
 
+Based on the model output, adding the speaker*trial interaction did not significantly improve model fit to the data (chi-sq(1) = 0.06, p = .80).
+This indicates that when listening to the speaker with ALS, the listeners' peak pupil dilation declined at a similar rate as when they were listening to the control speaker.
+
 #### Final Model
 
-```{r}
+Based on the model comparisons, model 3 provided the best fit to the data. Therefore, the final model output for peak pupil dilation is provided below.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
+```{r, include = F, warning = F, message = F}
+
+# Removing unneeded items from the environment
+
+rm(data, m0, m1, m2, m3, m4, peak_pupil_des, peak_pupil_df, pupil_roi, my_pal)
+
+```

### 2023-10-31 ASHA 2023 Analysis/data_analysis.Rmd
Rendered rmd file
Started alternate figure
Code not working as intended yet. WIP

@@
 author: "Micah E. Hirsch"
-date: "2023-10-29"
+date: "2023-10-30"
 output: 
@@
   annotate("rect", xmin = 500, xmax = 2908, ymin = -50, ymax = 150, alpha = 0.2) +
-  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
+  coord_cartesian(xlim = c(0,4908), ylim = c(-100, 200)) +
   labs(title = "Control Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
@@
 als
+
+```
+
+```{r, include = F, eval = F}
+
+bin.length = 26
+
+data1 <- data %>%
+  filter(speaker == "ALS") %>%
+  group_by(code) %>%
+  mutate(bins = round(timebins/bin.length)*bin.length) %>%
+  group_by(subject, trial, bins, speaker, code) %>%
+  dplyr::summarize(pupil = mean(pupil.binned)) %>%
+  ungroup()
+
+control <- data %>%
+  filter(speaker == "Control") %>% 
+  select(subject, trial, timebins, speaker, code, pupil.binned) %>%
+  rename(pupil = pupil.binned,
+         bins = timebins)
+
+normed_pupil <- rbind(data1, control)
+
+normed_pupil %>%
+  count(speaker, code) %>%
+  group_by(speaker) %>%
+  summarize(row_m = mean(n), sd = sd(n))
+
+normed_pupil %>%
+  # Selecting relevant parameters
+  dplyr::select(subject, trial, speaker, bins, code) %>%
+  # Filtering out time before start of phrase
+  dplyr::filter(bins >= 0) %>%
+  dplyr::group_by(speaker, code) %>%
+  dplyr::summarize(length = max(bins)) %>%
+  dplyr::group_by(speaker) %>%
+  dplyr::summarize(av_length = mean(length),
+                   av_offset = av_length - 3000,
+                   av_end_roi = av_length - 2000) %>%
+  gt::gt() %>%
+  gt::cols_label(speaker = "Speaker",
+                 av_length = "Average Trial Length (ms)",
+                 av_offset = "Average Phrase Length (ms)",
+                 av_end_roi = "Average Endpoint for ROI (ms)")
 

### 2023-11-03 ASHA 2023 Analysis/data_analysis.Rmd
put code chunk on hold
normalizing the ALs trials is currently a wip. This commit is just saving the progress made so far. However, the code will not be evaluated when creating the rendered rmd document

@@
 
-```{r, include = F, eval = F}
+```{r, echo = F, eval = F}
 
-bin.length = 26
+# This code block is a Work in Progress.
+Since the phrase legnths between the ALS speaker and the Control speaker are very different, this block is attempting to downsample the ALS trials to be a similar length of the control trials.
+The current code is able to downsample to a point where the number of rows for ALS trials and control trials are similar (in the new normed_pupil df).
+However, the timestamp labels were not updated yet.
+This analysis may be moved to an alternative script to further flesh this out.
+
+bin.length = 26.27
 
@@
 
-normed_pupil <- rbind(data1, control)
-
 normed_pupil %>%
@@
 
-normed_pupil %>%
-  # Selecting relevant parameters
-  dplyr::select(subject, trial, speaker, bins, code) %>%
-  # Filtering out time before start of phrase
-  dplyr::filter(bins >= 0) %>%
-  dplyr::group_by(speaker, code) %>%
-  dplyr::summarize(length = max(bins)) %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize(av_length = mean(length),
-                   av_offset = av_length - 3000,
-                   av_end_roi = av_length - 2000) %>%
-  gt::gt() %>%
-  gt::cols_label(speaker = "Speaker",
-                 av_length = "Average Trial Length (ms)",
-                 av_offset = "Average Phrase Length (ms)",
-                 av_end_roi = "Average Endpoint for ROI (ms)")
+data %>%
+  count(speaker, code) %>%
+  group_by(speaker) %>%
+  summarize(row_m = mean(n), sd = sd(n))
 

### 2023-11-03 ASHA 2023 Analysis/data_analysis.Rmd

@@
 # This code block is a Work in Progress.
-Since the phrase legnths between the ALS speaker and the Control speaker are very different, this block is attempting to downsample the ALS trials to be a similar length of the control trials.
+Since the phrase legnth between the ALS speaker and the Control speaker are very different, this block is attempting to downsample the ALS trials to be a similar length of the control trials.
 The current code is able to downsample to a point where the number of rows for ALS trials and control trials are similar (in the new normed_pupil df).
@@
 This analysis may be moved to an alternative script to further flesh this out.
+
+# This code block is not included in the ASHA 2023 output file.
 

### 2023-11-03 ASHA 2023 Analysis/data_analysis.Rmd
Formatted rmd script
to prepare for rendering

--- 2023-10-31 ASHA 2023 Analysis/data_analysis.Rmd
@@
 Any questions or comments regarding this project or analysis may be directed to Micah Hirsch, mhirsch@fsu.edu.
+
+This analysis was conducted using R version 4.3.1
 
@@
 For the control speaker, the mean intelligibility level was 93.62% with a standard deviation of 2.90.
+Based on their intelligibility level, the speaker with ALS is determined to have a moderate severity of dysarthria (Stipancic et al., 2021).
 
@@
 ```
+
 ```{r, include = F, warning = F, message = F}
@@
 
-Both the conditional and unconditional ICC is 0.099.
+Both the conditional and unconditional ICC is .099 (rounded to .1 in model output).
 
@@
 
-summary(m0)
+sjPlot::tab_model(m0, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
 
@@
 
-summary(m1)
+sjPlot::tab_model(m1, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
 
@@
 
-To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to model 1
+To control for trial order, the trial order variable was added as a fixed effect to model 2.
+A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to model 1
 
@@
 
-summary(m2)
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Perceived Listening Effort")
 
@@
 
-As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1) = 0.412, p = 0.52).
+As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1) = .412, p = .52).
 This indicates that trial order is not an important predictor for perceived listening effort ratings and that ratings for the speakers were fairly consistent across trials.
@@
 
-summary(m3)
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
 
@@
 
-Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1) = 469.74, p <0.001).
-The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.
+Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1) = 469.74, p <.001).
+The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for the ALS speaker.
 
@@
 
-To test whether there are any potential differences in perceived effort ratings across trials, we are testing whether adding an interaction between trial_c and speaker improves model fit.
+To test whether there are any potential differences in perceived effort ratings across trials, we are testing whether adding an interaction between trial order and speaker improves model fit.
 Again, a likelihood ratio test was conducted to see if model 4 provides a better fit to the data compared to model 3.
@@
 
-summary(m4)
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Trial Order * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
 
@@
 
-Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = 0.14, p= 0.70).
+Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = .14, p= .70).
 The fixed effect estimate for the interaction term is also not significant.
-This indicates that effort ratings across the trials for the ALS speaker were not significantly different from each other.
+This indicates that effort ratings across the trials for both the ALS speaker and the control speaker were not significantly different from each other.
+In other words, the ratings for both the ALS speaker and control speaker stayed consistent across the experimental trials.
 
@@
 
-sjPlot::tab_model(per_effort_model_final, file = "Final Model Outputs/perceived_effort_model.html")
+sjPlot::tab_model(per_effort_model_final, pred.labels = c("Intercept", "Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Final Model Outputs/perceived_effort_model.html")
 ```
 
-The final model indicates that the perceived effort ratings for the ALS speaker is about 2.35 points higher than they control speaker.
+The final model indicates that the perceived effort ratings for the ALS speaker is about 2.35 points higher than the control speaker.
 This is consistent with the results shown in the descriptive statistics.
@@
 The ALS speaker has a slower speech rate compared to the control speaker.
-Therefore, we are going to calculate the average trial legnth to see on average, how different the lengths were for ALS trials and control speaker trials.
+Therefore, we are going to calculate the average phrase length to see on average, how different the phrase lengths were for ALS trials and control speaker trials.
+To calculate this, the first 3000 ms were filtered out of the df to remove the trial time before the phrase onset.
+Then, the average trial length was calculated (phrase presentation + 3000 ms retention period).
+The average phrase offset was calculated by subtracting 2000 ms from the average trial length, since our analysis region of interests stops 1000 ms after the phrase offset.
+
 We will need this information to be able to mark the average phrase offset time and average end point for the analysis region of interest for visualization purposes.
@@
   dplyr::select(subject, trial, speaker, timebins, code) %>%
-  # Filtering out time before start of phrase
+  # Filtering out times before phase onset
   dplyr::filter(timebins >= 0) %>%
   dplyr::group_by(speaker, code) %>%
-  dplyr::summarize(length = max(timebins)) %>%
+  dplyr::summarize(length = max(timebins) - min(timebins)) %>%
   dplyr::group_by(speaker) %>%
   dplyr::summarize(av_length = mean(length),
-                   av_offset = av_length - 3000,
+                   av_phrase = av_length - 3000,
                    av_end_roi = av_length - 2000) %>%
@@
                  av_length = "Average Trial Length (ms)",
-                 av_offset = "Average Phrase Length (ms)",
+                 av_phrase = "Average Phrase Length (ms)",
                  av_end_roi = "Average Endpoint for ROI (ms)")
@@
 
-Average trial length includes the phrase presentation plus the 3000 ms retention period.
-Average phrase legnth is the average length of the recording.
-Average Endpoint of the Analysis ROI is the 1000 ms after the average phrase endpoint (calculated this for data visualization purposes).
 On average, the phrase length (phrase presentation), was about 2461 ms longer for the ALS trials than the control speaker trials.
@@
 ### Visualization
+
+#### Average Pupil Response to the Control Speaker
 
@@
   annotate("rect", xmin = 500, xmax = 2908, ymin = -50, ymax = 150, alpha = 0.2) +
-  coord_cartesian(xlim = c(0,4908), ylim = c(-100, 200)) +
+  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
   labs(title = "Control Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
@@
 
+#### Individual Listener Pupil Response to the Control Speaker
+
+We can also look at this tracking grouped by the listeners in our study. From this figure, we can see there are variable responses to the the control speaker across listeners.
+
+```{r, warning = F, message = F}
+
+control_byListener <- data %>%
+  dplyr::filter(rep_acc == "accurate") %>%
+  dplyr::filter(speaker == "Control") %>%
+  ggplot() +
+  aes(x = timebins,
+      y = pupil.binned,
+      group = subject,
+      color = speaker) +
+  geom_smooth() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1908, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2908, ymin = -50, ymax = 150, alpha = 0.2) +
+  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
+  labs(title = "Pupil Response to Control Speaker (Grouped by Listener)", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
+  theme_bw() +
+  scale_color_manual(values = my_pal[1]) +
+  theme(legend.position = "none", aspect.ratio = 1)
+
+control_byListener
+
+```
+
+#### Individual Listener Pupil Response to the ALS Speaker
+
 This is the smoothed pupil dilation tracking for the ALS Speaker.
@@
 
-```{r, include = F, eval = F}
+#### Average Pupil Response to the ALS Speaker
 
-bin.length = 26
+We can also look at this tracking grouped by the listeners in our study.
+From this figure, we can see there are variable responses to the the  speaker with ALS across listeners.
+Some of listener trackings for this speaker are following an odd pattern.
+Future analyses will investigate these individual listener and trial trackings to check for any potential tracking errors.
 
-data1 <- data %>%
-  filter(speaker == "ALS") %>%
-  group_by(code) %>%
-  mutate(bins = round(timebins/bin.length)*bin.length) %>%
-  group_by(subject, trial, bins, speaker, code) %>%
-  dplyr::summarize(pupil = mean(pupil.binned)) %>%
-  ungroup()
+```{r, warning = F, message = F}
 
-control <- data %>%
-  filter(speaker == "Control") %>% 
-  select(subject, trial, timebins, speaker, code, pupil.binned) %>%
-  rename(pupil = pupil.binned,
-         bins = timebins)
+als_byListener <- data %>%
+  dplyr::filter(rep_acc == "accurate") %>%
+  dplyr::filter(speaker == "ALS") %>%
+  ggplot() +
+  aes(x = timebins,
+      y = pupil.binned,
+      group = subject,
+      color = speaker) +
+  geom_smooth() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 4369, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 5369, ymin = -50, ymax = 150, alpha = 0.2) +
+  coord_cartesian(xlim = c(0,7369), ylim = c(-100, 200)) +
+  labs(title = "ALS Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
+  theme_bw() +
+  scale_color_manual(values = my_pal[2]) +
+  theme(legend.position = "none", aspect.ratio = 1)
 
-normed_pupil <- rbind(data1, control)
-
-normed_pupil %>%
-  count(speaker, code) %>%
-  group_by(speaker) %>%
-  summarize(row_m = mean(n), sd = sd(n))
-
-normed_pupil %>%
-  # Selecting relevant parameters
-  dplyr::select(subject, trial, speaker, bins, code) %>%
-  # Filtering out time before start of phrase
-  dplyr::filter(bins >= 0) %>%
-  dplyr::group_by(speaker, code) %>%
-  dplyr::summarize(length = max(bins)) %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize(av_length = mean(length),
-                   av_offset = av_length - 3000,
-                   av_end_roi = av_length - 2000) %>%
-  gt::gt() %>%
-  gt::cols_label(speaker = "Speaker",
-                 av_length = "Average Trial Length (ms)",
-                 av_offset = "Average Phrase Length (ms)",
-                 av_end_roi = "Average Endpoint for ROI (ms)")
+als_byListener
 
@@
 
-rm(als, control)
+rm(als, control, als_byListener, control_byListener)
 
@@
 
-### Fully Unconditional Model
+### Unconditional Growth Model
 
@@
 phrase) as random intercepts.
-I also added the orthogonal polynomials as random slopes.
-I also used the AR(1) covariance structure to specifiy the random effect structure.
+I also added the orthogonal polynomials as random slopes (poly1, poly2, and poly3).
+I also attempted to use the AR(1) covariance structure to specifiy the random effect structure.
 However, I was not able to run the model and ran into convergence issues.
@@
 
-More details about how I respecified the model can be seen in the code block.
+For more details on how the unconditional growth model was specified, please click on "show" to see the code.
 
-The adjusted ICC for the unconditional growth model is .109. The unadjusted ICC for the conditional growth model is .107.
+The adjusted ICC for the unconditional growth model is .109 (rounded to .11 in model output). The unadjusted ICC for the conditional growth model is .107.
 
@@
 
-## I kept running into issues using glmmTMB, so I switched to lme4
-
-library(lme4)
-library(lmerTest)
+# I kept running into issues using glmmTMB. Therefore, I switched to using lme4 for these analyses
 
@@
 
-summary(m0)
+sjPlot::tab_model(m0, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term"),
+                  dv.labels = "Pupil Dilation")
 
@@
 
-summary(m1)
+sjPlot::tab_model(m1, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]"),
+                  dv.labels = "Pupil Dilation")
 
@@
 
-In model two, the interaction between poly1 and speaker was added to the model as a fixed effect to test whether linear growth was greater for the ALS speaker compared to the control speaker.
+In model two, the interaction between the linear growth term and speaker was added to the model as a fixed effect.
 A likelihood ratio test was also conducted to test if model 2 provides a better fit to the data compared to model 1.
@@
 
-summary(m2)
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]", 
+                                      "Linear Term * Speaker [ALS]"),
+                  dv.labels = "Pupil Dilation")
 
@@
 
-Based on the model output, the interaction between poly1 and speaker is significant, indicating that, across listeners, linear growth in pupil dilation while listening to the ALS speaker was significantly greater than when listeners were listening to the control speaker.
-The likelihood ratio test also indicated that model 2 provided a better fit to the data than model 1 (chi-sq(1) = 649.22, p <0.001).
+Based on the model output, the interaction between the linear growth term and speaker is significant, indicating that, across listeners, pupil dilation grew at a faster rate for the ALS speaker compared to the control speaker.
+The likelihood ratio test also indicated that model 2 provided a better fit to the data than model 1 (chi-sq(1) = 649.22, p <.001).
 
@@
 
-summary(m3)
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]", 
+                                      "Linear Term * Speaker [ALS]", "Quadratic Term * Speaker [ALS]"),
+                  dv.labels = "Pupil Dilation")
 
@@
 ```
-Based on the model output, the interaction between the quadratic growth term and speaker improved the model fit (ch-sq(1) = 147.72, p<0.001).
-This indicates, on average, the shape of the listeners' pupil dilation is different while listening to the ALS speaker vs the control speaker.
+
+Based on the model output, the interaction between the quadratic growth term and speaker improved the model fit (ch-sq(1) = 147.72, p<.001).
+This indicates, on average, the shape of the listeners' pupil dilation is significantly different while listening to the ALS speaker vs the control speaker.
 
@@
 
-summary(m4)
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]", 
+                                      "Linear Term * Speaker [ALS]", "Quadratic Term * Speaker [ALS]", "Cubic Term * Speaker [ALS]"),
+                  dv.labels = "Pupil Dilation")
 
@@
 ```
-Based on the model output, model 4 provides a better fit to the data compared to model 3 (chi-sq(1) = 9.25, P = 0.002).
-What is interesting is that the only fixed effects that are significant in this model are the main effect of the cubic growth term and the interaction between the cubic growth term and speaker, indicating the cubic pupil dilation growth is significantly different between these two speakers.
 
-```{r, message = F, warning = F}
-
-sjPlot::tab_model(m4, file = "Final Model Outputs/final_pupil_dilation_model.html")
-
-```
+Based on the model output, model 4 provides a better fit to the data compared to model 3 (chi-sq(1) = 9.25, P = .002).
+Interestingly, the only fixed effects that are significant in this model are the main effect of the cubic growth term and the interaction between the cubic growth term and speaker.
+However, the model output still indicates that there are significant differences in the shape of pupil dilation for the ALS speaker compared to the control speaker.
 
@@
 
-sjPlot::tab_model(m1, m2, m3, m4, file = "Final Model Outputs/pupil_dilation_model.html")
+sjPlot::tab_model(m1, m2, m3, m4, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]", 
+                                      "Linear Term * Speaker [ALS]", "Quadratic Term * Speaker [ALS]", "Cubic Term * Speaker [ALS]"),
+                  dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
+                  file = "Final Model Outputs/pupil_dilation_model.html")
 
@@
 ## Peak Pupil Dilation Descriptives
+
+The mean peak pupil dilation, standard deviation, and standard error was calculated for each speaker and are reported below.
+In general, the mean peak pupil dilation for the ALS speaker is about 69 units higher than for the control speaker.
 
@@
 
-### Peak Pupil Dilation Distribution Plot
+### Visualization
+
+Below are two figures to visualize the distribution of peak pupil dilation across listeners and trials for the ALS and Control Speakers.
+
+#### Peak Pupil Dilation Distribution Plot
 
@@
 ```
-### Peak Pupil Dilation Bar Plot
+
+#### Peak Pupil Dilation Bar Plot
 
@@
 
-summary(m0)
+sjPlot::tab_model(m0, pred.labels = "Intercept",
+                  dv.labels = "Peak Pupil Dilation")
 
@@
 
-summary(m1)
+sjPlot::tab_model(m1, pred.labels = "Intercept",
+                  dv.labels = "Peak Pupil Dilation")
 
@@
 
-Model 1 provided a better fit to the data (chi-sq(2) = 43.44, p < 0.001).
+Model 1 provided a better fit to the data (chi-sq(2) = 43.44, p < .001).
 
@@
 
-summary(m2)
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Peak Pupil Dilation")
 
@@
 
-Adding trial to the model as a fixed effect significantly improved model fit (chi-sq(1) = p<0.001).
+Adding trial to the model as a fixed effect significantly improved model fit (chi-sq(1) = 12.73, p<.001).
 Based on the fixed effect estimate for trial order, peak pupil dilation decreased approximately -1.23 units on average with each trial.
@@
 
-summary(m3)
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 
@@
 
-Adding speaker to the model improved the model fit to the data (ch-sq(1), p <0.001).
+Adding speaker to the model improved the model fit to the data (ch-sq(1) = 102.96, p <.001).
 Controlling for trial order, listeners' peak pupil dilation, on average, was about 85.22 units higher compared to when they were listening to the control speaker.
@@
 
-summary(m4)
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Trial Order * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 
@@
 
-final_model <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
+final_model <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) + (1|code), 
               data = peak_pupil_df, control = lmerControl(
@@
 
-sjPlot::tab_model(final_model, file = "Final Model Outputs/peak_pupil_model.html")
+sjPlot::tab_model(final_model, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation",
+                  file = "Final Model Outputs/peak_pupil_model.html")
 
@@
 
-rm(data, m0, m1, m2, m3, m4, peak_pupil_des, peak_pupil_df, pupil_roi, my_pal)
+rm(data, m0, m1, m2, m3, m4, final_model, peak_pupil_des, peak_pupil_df, pupil_roi, my_pal)
 

### 2023-12-01 ASHA 2023 Analysis/standardized_pupil_analysis.Rmd
Started normalization-WIP

--- 2023-11-03 ASHA 2023 Analysis/data_analysis.Rmd
@@
 ---
-title: "ASHA 2023 Listening Effort Data Analysis"
+title: "Standardized Pupil Analysis"
 author: "Micah E. Hirsch"
-date: "2023-10-30"
-output: 
-  html_document:
-    toc: true
-    toc_float: true
-    code_folding: hide
+date: "2023-12-01"
+output: html_document
 ---
@@
 
-The purpose of this document is to run preliminary analyses and visualizations for this project.
-These results were prepared specifically for a poster presentation at the the 2023 Annual American Speech Language Hearing Association Convention being held in Boston, MA in November 2023.
+The preliminary pupillometry results (found in the data_analysis.Rmd file) indicicated that the shape of the pupil dilation tracking is different while listening to the speaker with ALS verses listening to the control speaker.
+However, it is not clear whether this finidng is influenced by the difference in the length of the audio files for the control speaker and the ALS speaker.
+Due to the speaker's slowed speech rate, the ALS speaker's audio files were longer than the control speaker's files.
 
-Any questions or comments regarding this project or analysis may be directed to Micah Hirsch, mhirsch@fsu.edu.
+Therefore, in order to control for the effect of the difference in speech rate, we will normalize the trial-level data for the ALS speaker so that the trial lengths are similar to the control speaker's trials.
+
+This reanalysis is conducted using R version 4.3.1
 
@@
 
-# Descriptive Statistics
+# Pupil Dilation Normalization
 
-## Participant Demographics
+Work in Progress
 
-First, we will present the demographics of our listeners To date, we recruited 20 listeners to participate in the study.
-Three listeners were removed from the analysis.
-One listener decided to withdraw from the study, so their data was removed from the database.
-For the other two participants, pupillometry data was either not recorded or lost due to instrumental issues.
-Therefore, for this analysis, we have 17 participants.
-The descriptive information for these participants are presented below.
+```{r}
 
-```{r, warning = F, message = F}
-
-# Load in dataset for listener demographics
-
-listener_demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv")
-
-# Extracting listener gender data
-
-gender <- listener_demo %>%
-  dplyr::select(gender) %>%
-  dplyr::group_by(gender) %>%
-  dplyr::summarize(Frequency = n()) %>%
-  dplyr::rename(Category = "gender") %>%
-  dplyr::mutate(Group = "Gender",
-                Category = fct_relevel(Category, c("woman", "man", "nonbinary"))) %>%
-  dplyr::arrange(Category)
-
-## Extracting listener race data
-
-race <- listener_demo %>%
-  dplyr::select(race) %>%
-  dplyr::group_by(race) %>%
-  dplyr::summarize(Frequency = n()) %>%
-  dplyr::rename(Category = "race") %>%
-  dplyr::mutate(Group = "Race",
-                Category = fct_relevel(Category, c("white/Caucasian", "Biracial or Multiracial", 
-                                                   "race not listed"))) %>%
-  dplyr::arrange(Category)
-
-## Extracting listener ethnicity data
-
-ethnicity <- listener_demo %>%
-  dplyr::select(ethnicity) %>%
-  dplyr::group_by(ethnicity) %>%
-  dplyr::summarize(Frequency = n()) %>%
-  dplyr::rename(Category = "ethnicity") %>%
-  dplyr::mutate(Group = "Ethnicity",
-                Category = fct_relevel(Category, c("Hispanic/Latino(a/e)", 
-                                                   "Not Hispanic/Latino(a/e)"))) %>%
-  dplyr::arrange(Category)
-
-# Creating Demographic Table 1
-
-dplyr::full_join(gender, race) %>%
-  dplyr::full_join(., ethnicity) %>%
-  # Initiating a gt table
-  gt::gt(rowname_col = "Category",
-         groupname_col = "Group")
-
-```
-
-As indicated in the table above, our participants were primarily white/Caucasian, women, and not Hispanic or Latino(a/e).
-All participants reported no history of a speech, language, or hearing disorder.
-All participants were fluent English speakers.
-However, two indicated their native language was different than American English.
-These languages were Spanish (n=1) and Turkish (n=1).
-
-```{r, warning = F, message = F}
-
-listener_demo %>%
-  dplyr::select(c(age)) %>%
-  dplyr::summarize(Mean = mean(age), SD = sd(age),
-                   Min = min(age), Max = max(age)) %>%
-  dplyr::mutate(Category = "Age") %>%
-  dplyr::relocate(Category, .before = Mean) %>%
-  gt::gt() %>%
-  gt::fmt_number(columns = c(Mean, SD),
-                decimals = 2) %>%
-  gt::cols_label(Category = "")
-
-```
-
-The mean age of the listeners is 26.53 years old. Listener age ranged from 18 to 37 years old.
-
-```{r, include = F, warning = F, message = F}
-
-# Removing unneeded items from the environment
-
-rm(ethnicity, gender, listener_demo, race)
-
-```
-
-## Speaker Intelligibility and Phrase Accuracy
-
-### Mean Speaker Intelligibility
-
-Across all trials in the study, the intelligibility level for each of the speaker is reported below.
-Across listeners, the mean intelligibility for the ALS speaker was 81.08% with a standard deviation of 3.86.
-For the control speaker, the mean intelligibility level was 93.62% with a standard deviation of 2.90.
-
-```{r, warning = F, message = F}
-
-# Load in cleaned data for pupil and perceived listening effort data
-
+# Import pupil dilation data
 data <- rio::import("Cleaned Data/cleaned_data.csv") %>%
@@
 
-# Selecting variables needed to calculate mean speaker intelligibility
-intel_df <- data %>%
-  dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) %>%
-  dplyr::distinct()
+# Pupil Normalization
 
-# Calculating intelligibility for the control and ALS speaker for each listener
-intel_byListener <- intel_df %>%
-  dplyr::group_by(speaker, subject) %>%
-  dplyr::summarize(Intelligibility = sum(correct_words)/sum(target_number)*100)
+bin.length = 2.29
 
-intel_byListener %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize("Mean Intelligibility" = mean(Intelligibility), SD = sd(Intelligibility)) %>%
-  gt::gt() %>%
-  gt::fmt_number(columns = c("Mean Intelligibility", SD),
-                 decimals = 2) %>%
-  gt::cols_label(speaker = "Speaker")
+ALS_trials <- data %>%
+  dplyr::filter(speaker == "ALS") %>%
+  dplyr::mutate(time_n = round(timebins/bin.length)*bin.length) %>%
+  dplyr::group_by(subject, trial_c, speaker, time_n, code) %>%
+  dplyr::summarize(normed_pupil = mean(pupil.binned)) %>%
+  dplyr::ungroup()
 
@@
 
-### Phrase Accuracy
-
-For this analysis, we are only looking at the trials that listeners were able to accurately recognize (e.g.
-100% intelligible to the listener).
-Therefore, we are looking at the proportion of trials for each speaker that listeners were able to accurately recognize compared to the total number of trials.
-Although our ALS speaker's intelligibility was fairly high, only about 49% of all the ALS speaker's trials were accurately recognized across listeners.
-
-```{r, warning = F, message = F}
-
-# Calculating repetition accuracy descriptive data
-
-intel_df %>%
-  dplyr::group_by(speaker, rep_acc) %>%
-  dplyr::summarize(n = n()) %>%
-  tidyr::pivot_wider(names_from = rep_acc,
-                     values_from = n) %>%
-  dplyr::mutate(percent_accurate = accurate/(accurate + inaccurate)*100) %>%
-  dplyr::ungroup() %>%
-  gt::gt(rowname_col = "speaker") %>%
-  gt::fmt_number(columns = percent_accurate,
-                 decimals = 2) %>%
-  gt::cols_label(accurate = "Accurate",
-                 inaccurate = "Inaccurate",
-                 percent_accurate = "Percent Accurate")
-
-```
-```{r, include = F, warning = F, message = F}
-
-# removing unneeded items from the environment
-
-rm(intel_byListener, intel_df)
-
-```
-
-# Perceived Listening Effort Ratings
-
-## Descriptives
-
-```{r, warning = F, message = F}
-
-perceived_effort_df <- data %>%
-  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code, trial_c) %>%
-  dplyr::distinct() %>%
-  # For this particular analysis, we are just looking at results for accurate repetition
-  # However, the final project will look at rep_acc too
-  dplyr::filter(rep_acc == "accurate")
-
-per_effort_des <- perceived_effort_df %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize(per_effort = mean(effort_rating), 
-                   sd = sd(effort_rating, na.rm = T),
-                   se = sd/sqrt(n()))
-
-per_effort_des %>%
-  gt::gt() %>%
-  gt::fmt_number(columns = per_effort:se,
-                 decimals = 2) %>%
-  gt::cols_label(speaker = "Speaker",
-                  per_effort = "Perceived Effort",
-                  sd = "SD",
-                  se = "SE")
-
-```
-
-As reported in the table above and shown in the plot below, the mean perceived effort ratings for the ALS speaker is about 2.13 points higher than the control speaker.
-This is reflecting the difference in mean perceived effort ratings between the ALS and control speaker, even when the phrase was intelligible to the listener.
-
-```{r, warning = F, message = F}
-
-# Creating a color palette vector for plots
-my_pal <- c("#CD9FCC", "#7E2E84")
-
-# Plotting the mean perceived effort ratings for the ALS and control speaker
-per_effort_des %>%
-  ggplot() +
-   aes(x = speaker,
-      y = per_effort,
-      group = speaker,
-      color = speaker,
-      fill = speaker) +
-  geom_bar(stat = "identity", alpha = 0.6) +
-  geom_errorbar(aes(x = speaker, ymin = per_effort - se, ymax = per_effort + se), width = 0.4) +
-  labs(x = "Speaker", y = "Mean Perceived Listening Effort Rating") +
-  theme_bw() +
-  theme(legend.position = "none", aspect.ratio = 1) +
-  scale_color_manual(values = my_pal) +
-  scale_fill_manual(values = my_pal)
-
-# Saving the ggplot
-
-ggsave("Plots/Perceived_effort.png", plot = last_plot(), height = 5, width = 5, units = "in")
-
-```
-
-```{r, warning = F, message = F}
-
-perceived_effort_df %>%
-  ggplot() +
-  aes(y = speaker,
-      x = effort_rating,
-      color = speaker,
-      fill = speaker) +
-  geom_density_ridges(scale = 0.7, alpha = 0.5) +
-  labs(x = "Perceived Listening Effort Rating", y = "Speaker") +
-  theme_bw() +
-   theme(legend.position = "none", aspect.ratio = 1) +
-  scale_fill_manual(values = my_pal) +
-  scale_color_manual(values = my_pal)
-
-# Saving the ggplot
-
-ggsave("Plots/Perceived_effort_dis.png", plot = last_plot(), height = 5, width = 5, units = "in")
-
-```
-
-## Perceived Listening Effort Model
-
-### Fully Unconditional Model
-
-To analyze the perceived effort ratings, we are going to use linear mixed-effects models.
-To start, we are going to specify our fully unconditional model.
-In this model, we included random intercepts for subject (i.e.
-our listeners) and code (i.e.
-the different phrases).
-No fixed effects have been added to the model.
-We also calculated ICC from this model.
-
-Both the conditional and unconditional ICC is 0.099.
-
-```{r, warning = F, message = F}
-
-m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = perceived_effort_df)
-
-summary(m0)
-
-performance::icc(m0)
-
-```
-
-### Model 1
-
-Now we will add trial number (i.e.
-trial_c) as a random slope for both subject and code.
-A likelihood ratio test was also ran to test whether model 1 provided a better fit to the data than the unconditional model.
-
-```{r, warning = F, message = F}
-
-m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), 
-           data = perceived_effort_df)
-
-summary(m1)
-
-anova(m0, m1)
-
-```
-
-The likelihood ratio test indicated that model 1 provided a better fit to the data than the unconditional model (chi-sq(4) = 227.79, p <0.001). Therefore, we will retain model 1.
-
-### Model 2
-
-To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to model 1
-
-```{r, warning = F, message = F}
-
-m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), 
-              data = perceived_effort_df)
-
-summary(m2)
-
-anova(m1, m2)
-
-```
-
-As indicated from the model, adding trial order did not significantly improve the model fit to the data (chi-sq(1) = 0.412, p = 0.52).
-This indicates that trial order is not an important predictor for perceived listening effort ratings and that ratings for the speakers were fairly consistent across trials.
-
-### Model 3
-
-In model 3, we added the speaker fixed effect to test whether the mean difference in perceived listening effort rating is significantly difference for the ALS speaker compared to the control speaker.
-A likelihood ratio test was also conducted to test whether model 3 provides a better fit to the data compared to model 2.
-
-```{r, warning = F, message = F}
-
-m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), 
-           data = perceived_effort_df)
-
-summary(m3)
-
-anova(m2, m3)
-
-```
-
-Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1) = 469.74, p <0.001).
-The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.
-
-### Model 4
-
-To test whether there are any potential differences in perceived effort ratings across trials, we are testing whether adding an interaction between trial_c and speaker improves model fit.
-Again, a likelihood ratio test was conducted to see if model 4 provides a better fit to the data compared to model 3.
-
-```{r, warning = F, message = F}
-
-m4 <- glmmTMB(effort_rating ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) + (trial_c|code), 
-           data = perceived_effort_df)
-
-summary(m4)
-
-anova(m3, m4)
-
-```
-
-Based on the model output, adding the interaction term did not significantly improve model fit (chi-sq(1) = 0.14, p= 0.70).
-The fixed effect estimate for the interaction term is also not significant.
-This indicates that effort ratings across the trials for the ALS speaker were not significantly different from each other.
-
-### Final Model
-
-Based on our model building, we will now specify our final model.
-Since the trial fixed effect as well as the interaction between trial and speaker were not significant, nor did they improve model fit, they were removed from the final model.
-All other output remained the same.
-
 ```{r}
-
-per_effort_model_final <- glmmTMB(effort_rating ~ speaker + (trial_c|subject) + (trial_c|code), 
-                               data = perceived_effort_df)
-
-sjPlot::tab_model(per_effort_model_final, file = "Final Model Outputs/perceived_effort_model.html")
-```
-
-The final model indicates that the perceived effort ratings for the ALS speaker is about 2.35 points higher than they control speaker.
-This is consistent with the results shown in the descriptive statistics.
-
-```{r, include = F, warning = F, message = F}
-
-# Remove unneeded items from the environment
-
-rm(m0, m1, m2, m3, m4, per_effort_des, perceived_effort_df, per_effort_model_final)
-
-```
-
-# Pupil Dilation
-
-## Descriptives
-
-The ALS speaker has a slower speech rate compared to the control speaker.
-Therefore, we are going to calculate the average trial legnth to see on average, how different the lengths were for ALS trials and control speaker trials.
-We will need this information to be able to mark the average phrase offset time and average end point for the analysis region of interest for visualization purposes.
-
-```{r, warning = F, message = F}
 
@@
 ```
-
-Average trial length includes the phrase presentation plus the 3000 ms retention period.
-Average phrase legnth is the average length of the recording.
-Average Endpoint of the Analysis ROI is the 1000 ms after the average phrase endpoint (calculated this for data visualization purposes).
-On average, the phrase length (phrase presentation), was about 2461 ms longer for the ALS trials than the control speaker trials.
-
-### Visualization
-
-This is the smoothed pupil dilation tracking for the Control Speaker.
-The solid line indicates the start of the phrase.
-The dotted line indicates the average phrase offset.
-The grey box represents the region of interest for analysis.
-
-```{r, warning = F, message = F}
-
-control <- data %>%
-  dplyr::filter(rep_acc == "accurate") %>%
-  dplyr::filter(speaker == "Control") %>%
-  ggplot() +
-  aes(x = timebins,
-      y = pupil.binned,
-      color = speaker) +
-  geom_smooth() +
-  geom_vline(xintercept = 0, size = 0.35) +
-  geom_vline(xintercept = 1908, linetype = 2) +
-  annotate("rect", xmin = 500, xmax = 2908, ymin = -50, ymax = 150, alpha = 0.2) +
-  coord_cartesian(xlim = c(0,4908), ylim = c(-100, 200)) +
-  labs(title = "Control Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
-  theme_bw() +
-  scale_color_manual(values = my_pal[1]) +
-  theme(legend.position = "none", aspect.ratio = 1)
-
-control
-
-```
-
-This is the smoothed pupil dilation tracking for the ALS Speaker.
-The solid line indicates the start of the phrase.
-The dotted line indicates the average phrase offset.
-The grey box represents the region of interest for analysis.
-
-```{r}
-
-als <- data %>%
-  dplyr::filter(rep_acc == "accurate") %>%
-  dplyr::filter(speaker == "ALS") %>%
-  ggplot() +
-  aes(x = timebins,
-      y = pupil.binned,
-      color = speaker) +
-  geom_smooth() +
-  geom_vline(xintercept = 0, size = 0.35) +
-  geom_vline(xintercept = 4369, linetype = 2) +
-  annotate("rect", xmin = 500, xmax = 5369, ymin = -50, ymax = 150, alpha = 0.2) +
-  coord_cartesian(xlim = c(0,7369), ylim = c(-100, 200)) +
-  labs(title = "ALS Speaker", x = "Time (ms)", y = "Pupil Dilation (arbitrary units)") +
-  theme_bw() +
-  scale_color_manual(values = my_pal[2]) +
-  theme(legend.position = "none", aspect.ratio = 1)
-
-als
-
-```
-
-```{r, echo = F, eval = F}
-
-# This code block is a Work in Progress.
-Since the phrase legnth between the ALS speaker and the Control speaker are very different, this block is attempting to downsample the ALS trials to be a similar length of the control trials.
-The current code is able to downsample to a point where the number of rows for ALS trials and control trials are similar (in the new normed_pupil df).
-However, the timestamp labels were not updated yet.
-This analysis may be moved to an alternative script to further flesh this out.
-
-# This code block is not included in the ASHA 2023 output file.
-
-bin.length = 26.27
-
-data1 <- data %>%
-  filter(speaker == "ALS") %>%
-  group_by(code) %>%
-  mutate(bins = round(timebins/bin.length)*bin.length) %>%
-  group_by(subject, trial, bins, speaker, code) %>%
-  dplyr::summarize(pupil = mean(pupil.binned)) %>%
-  ungroup()
-
-control <- data %>%
-  filter(speaker == "Control") %>% 
-  select(subject, trial, timebins, speaker, code, pupil.binned) %>%
-  rename(pupil = pupil.binned,
-         bins = timebins)
-
-normed_pupil %>%
-  count(speaker, code) %>%
-  group_by(speaker) %>%
-  summarize(row_m = mean(n), sd = sd(n))
-
-data %>%
-  count(speaker, code) %>%
-  group_by(speaker) %>%
-  summarize(row_m = mean(n), sd = sd(n))
-
-```
-
-```{r, include = F, warning = F, message = F}
-
-pupil_plot <- ggarrange(control, als, ncol = 2)
-
-ggsave("Plots/Pupil_dilation.png", plot = pupil_plot, width = 10, height = 5, units = "in")
-
-rm(als, control)
-
-```
-
-## Pupil Dilation Analysis
-
-### Fully Unconditional Model
-
-First, we ran an unconditional growth model.
-In the intial model, I first specified a model with the orthogonal polynomials added as fixed effects and subject (i.e.
-listener) and code (i.e.
-phrase) as random intercepts.
-I also added the orthogonal polynomials as random slopes.
-I also used the AR(1) covariance structure to specifiy the random effect structure.
-However, I was not able to run the model and ran into convergence issues.
-
-Therefore, I decided to respecify the random effect structure by removing the AR(1) specification.
-This led to a model with singular fit.
-In response to this, I decided to remove code (i.e.phrase) from the random effect structure in order to simplify it.
-
-More details about how I respecified the model can be seen in the code block.
-
-The adjusted ICC for the unconditional growth model is .109. The unadjusted ICC for the conditional growth model is .107.
-
-```{r, warning = F, message = F}
-
-# Specifying the region of interest for analysis (500 ms to 1000 ms after relative offset of the trial)
-
-pupil_roi <- data %>%
-  # For this analysis, we are only analyzing pupil dilation for accurate responses
-  dplyr::filter(rep_acc == "accurate") %>%
-  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
-  dplyr::filter(timebins >= 500) %>%
-  dplyr::group_by(subject, trial) %>%
-  dplyr::filter(timebins <= max(timebins) - 2000) %>%
-  ungroup()
-
-# Adding linear, quadratic, and cubic orthogonal polynomials to the df
-
-pupil_roi <- gazer::code_poly(pupil_roi, predictor = "timebins", poly.order = 3, orthogonal = T, draw.poly = F)
-
-# Exporting region of interest df (this can be used to )
-
-rio::export(pupil_roi, "Cleaned Data/pupil_analysis_roi.csv")
-
-## I kept running into issues using glmmTMB, so I switched to lme4
-
-library(lme4)
-library(lmerTest)
-
-# specifying unconditional growth model for pupil data
-
-m0 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 +
-                (poly1 + poly2 + poly3|subject), 
-              data = pupil_roi)
-
-summary(m0)
-
-performance::icc(m0)
-
-```
-
-### Model 1
-
-In the next model, we are adding a fixed effect for speaker to test whether on average, pupil dilation is greater for the ALS speaker compared to the control speaker.
-A likelihood ratio test was also conducted to test whether model 1 provides a better fit to the data than the unconditional growth model.
-
-```{r, warning = F, message = F}
-
-m1 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
-                (poly1 + poly2 + poly3|subject), 
-              data = pupil_roi)
-
-summary(m1)
-
-anova(m0, m1)
-
-```
-
-As indicated in the model output, when listening to the ALS speaker, on average, listeners had greater pupil dilation compared to when they were listening to the control speaker.
-The likelihood ratio test also indicates model 1 is a better fit to the data compared to the unconditional model (chi-sq(1) = 353.67, p <0.001).
-
-### Model 2
-
-In model two, the interaction between poly1 and speaker was added to the model as a fixed effect to test whether linear growth was greater for the ALS speaker compared to the control speaker.
-A likelihood ratio test was also conducted to test if model 2 provides a better fit to the data compared to model 1.
-
-Note: The initial run of this model did not converge, so an optimizer control was added to the model. This optimizer will remain in each subsequent model.
-
-```{r, warning = F, message = F}
-
-# Model 2 initially failed to converge, so adding a control statement for bobyqa optimizer
-
-control_bobyqa <- lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e9))
-
-m2 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
-             poly1*speaker +
-             (poly1 + poly2 + poly3|subject), 
-              data = pupil_roi, control = control_bobyqa)
-
-summary(m2)
-
-anova(m1, m2)
-
-```
-
-Based on the model output, the interaction between poly1 and speaker is significant, indicating that, across listeners, linear growth in pupil dilation while listening to the ALS speaker was significantly greater than when listeners were listening to the control speaker.
-The likelihood ratio test also indicated that model 2 provided a better fit to the data than model 1 (chi-sq(1) = 649.22, p <0.001).
-
-### Model 3
-
-In model three, the interaction between the quadratic growth term and speaker was added to the model.
-This tests whether the quadratic shape of pupil dilation is significantly different between the two speakers.
-A likelihood ratio test was also conducted to test whether model 3 provides a better fit to the data compared to model 2.
-
-```{r, warning = F, message = F}
-
-m3 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
-             poly1*speaker + poly2*speaker +
-             (poly1 + poly2 + poly3|subject), 
-              data = pupil_roi, control = control_bobyqa)
-
-summary(m3)
-
-anova(m2, m3)
-
-```
-Based on the model output, the interaction between the quadratic growth term and speaker improved the model fit (ch-sq(1) = 147.72, p<0.001).
-This indicates, on average, the shape of the listeners' pupil dilation is different while listening to the ALS speaker vs the control speaker.
-
-### Model 4
-
-In model for, we added the interaction between the cubic growth term and speaker.
-As with the other models, a likelihood ratio test was conducted to test whether model 4 provides a better fit to the data compared to model 3.
-
-```{r, message = F, warning = F}
-
-m4 <- lmer(pupil.binned ~ poly1 + poly2 + poly3 + speaker +
-             poly1*speaker + poly2*speaker + poly3*speaker +
-             (poly1 + poly2 + poly3|subject), data = pupil_roi, 
-           control = control_bobyqa)
-
-summary(m4)
-
-anova(m3, m4)
-
-```
-Based on the model output, model 4 provides a better fit to the data compared to model 3 (chi-sq(1) = 9.25, P = 0.002).
-What is interesting is that the only fixed effects that are significant in this model are the main effect of the cubic growth term and the interaction between the cubic growth term and speaker, indicating the cubic pupil dilation growth is significantly different between these two speakers.
-
-```{r, message = F, warning = F}
-
-sjPlot::tab_model(m4, file = "Final Model Outputs/final_pupil_dilation_model.html")
-
-```
-
-```{r, include = F, message = F, warning = F}
-
-# Saving model 4 output
-
-sjPlot::tab_model(m1, m2, m3, m4, file = "Final Model Outputs/pupil_dilation_model.html")
-
-# removing unneeded items from the environment
-
-rm(m0, m1, m2, m3, m4, pupil_plot, control_bobyqa)
-
-```
-
-# Alternative Pupil Dilation Analysis
-
-Since the trials for the ALS speaker are much longer than the control speaker's trials, we are going to conduct an exploratory analysis for the pupil dilation measure.
-For each of the listener's trial, we are going to extract the peak pupil dilation during the analysis region and model it as a function of speaker.
-
-## Peak Pupil Dilation Descriptives
-
-```{r, warning = F, message = F}
-
-peak_pupil_df <- pupil_roi %>%
-  dplyr::group_by(subject, trial_c, speaker, code) %>%
-  dplyr::summarize(peak_pupil = max(pupil.binned)) %>%
-  ungroup()
-
-peak_pupil_des <- peak_pupil_df %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize(pupil = mean(peak_pupil), 
-                   sd = sd(peak_pupil, na.rm = T),
-                   se = sd/sqrt(n()))
-
-peak_pupil_des %>%
-  gt::gt() %>%
-  gt::fmt_number(columns = pupil:se,
-                 decimals = 2) %>%
-  gt::cols_label(speaker = "Speaker",
-                  pupil = "Peak Pupil Dilation",
-                  sd = "SD",
-                  se = "SE")
-
-```
-
-### Peak Pupil Dilation Distribution Plot
-
-```{r, warning = F, message = F}
-
-peak_pupil_df %>%
-  ggplot() +
-  aes(y = speaker,
-      x = peak_pupil,
-      color = speaker,
-      fill = speaker) +
-  geom_density_ridges(scale = 0.7, alpha = 0.5) +
-  labs(x = "Mean Peak Pupil Dilation (arbitrary units)", y = "Speaker") +
-  theme_bw() +
-  theme(legend.position = "none", aspect.ratio = 1) +
-  scale_fill_manual(values = my_pal) +
-  scale_color_manual(values = my_pal)
-
-ggsave("Plots/peak_pupil_dis.png", plot = last_plot(), height = 5, width = 5, units = "in")
-
-```
-### Peak Pupil Dilation Bar Plot
-
-```{r, warning = F, message = F}
-
-peak_pupil_des %>%
-  ggplot() +
-   aes(x = speaker,
-      y = pupil,
-      group = speaker,
-      color = speaker,
-      fill = speaker) +
-  geom_bar(stat = "identity", alpha = 0.6) +
-  geom_errorbar(aes(x = speaker, ymin = pupil - se, ymax = pupil + se), width = 0.4) +
-  labs(x = "Speaker", y = "Mean Peak Pupil Dilation (arbitrary units)") +
-  theme_bw() +
-  theme(legend.position = "none", aspect.ratio = 1) +
-  scale_color_manual(values = my_pal) +
-  scale_fill_manual(values = my_pal)
-
-ggsave("Plots/peak_pupil_bar.png", plot = last_plot(), height = 5, width = 5, units = "in")
-
-```
-## Analysis
-
-### Fully Unconditional Model
-
-We first fit a fully unconditional model to the data to calculate ICC. The adjusted and unadjusted ICC is 0.364.
-
-```{r, warning = F, message = F}
-
-# Note: I ran into issues using glmmTMB, so used lme4 instead
-
-m0 <- lmer(peak_pupil ~ 1 + (1|subject) + (1|code), data = peak_pupil_df)
-
-summary(m0)
-
-performance::icc(m0)
-
-```
-
-### Model 1
-
-We added trial as random slopes for both the subject and code random effects.
-The model initially failed to converge, so an optimizer was added to the model.
-Then a singular fit warning was displayed, so the random slope of trial_c was removed from the code random effect.
-
-```{r, warning = F, message = F}
-
-# Model failed to converge, so added an optimizer
-# Singular fit so removed random slope for code.
-
-m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject) + (1|code), 
-           data = peak_pupil_df, control = lmerControl(
-                           optimizer ='optimx', optCtrl=list(method='nlminb')))
-
-summary(m1)
-
-anova(m0, m1)
-
-```
-
-Model 1 provided a better fit to the data (chi-sq(2) = 43.44, p < 0.001).
-
-### Model 2
-
-Next, in order to control for trial order, trial was added to the model as a fixed effect.
-
-```{r, warning = F, message = F}
-
-m2 <- lmer(peak_pupil ~ trial_c + (trial_c|subject) + (1|code), 
-           data = peak_pupil_df, control = lmerControl(
-                           optimizer ='optimx', optCtrl=list(method='nlminb')))
-
-summary(m2)
-
-anova(m1, m2)
-
-```
-
-Adding trial to the model as a fixed effect significantly improved model fit (chi-sq(1) = p<0.001).
-Based on the fixed effect estimate for trial order, peak pupil dilation decreased approximately -1.23 units on average with each trial.
-
-### Model 3
-
-Next, we added the fixed effect of speaker to the model to test whether peak pupil dilation was, on average, higher for when listening to the ALS speaker compared to the control speaker.
-
-```{r, warning = F, message = F}
-
-m3 <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) + (1|code), 
-           data = peak_pupil_df, control = lmerControl(
-                           optimizer ='optimx', optCtrl=list(method='nlminb')))
-
-summary(m3)
-
-anova(m2, m3)
-
-```
-
-Adding speaker to the model improved the model fit to the data (ch-sq(1), p <0.001).
-Controlling for trial order, listeners' peak pupil dilation, on average, was about 85.22 units higher compared to when they were listening to the control speaker.
-This indicates listeners were exerting more effort while listening to the ALS speaker compared to the control speaker.
-
-### Model 4
-
-Finally, the interaction between trial and speaker was added to the model to test whether peak pupil dilation for the ALS speaker change at a different rate across trials.
-
-```{r, warning = F, message = F}
-
-# Ran into convergence problem, so removed the trial_c random slope for code
-
-m4 <- lmer(peak_pupil ~ trial_c + speaker + speaker*trial_c + (trial_c|subject) + (1|code), 
-              data = peak_pupil_df, control = lmerControl(
-                           optimizer ='optimx', optCtrl=list(method='nlminb')))
-
-summary(m4)
-
-anova(m3, m4)
-
-```
-
-Based on the model output, adding the speaker*trial interaction did not significantly improve model fit to the data (chi-sq(1) = 0.06, p = .80).
-This indicates that when listening to the speaker with ALS, the listeners' peak pupil dilation declined at a similar rate as when they were listening to the control speaker.
-
-#### Final Model
-
-Based on the model comparisons, model 3 provided the best fit to the data. Therefore, the final model output for peak pupil dilation is provided below.
-
-```{r, warning = F, message = F}
-
-final_model <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
-              data = peak_pupil_df, control = lmerControl(
-                           optimizer ='optimx', optCtrl=list(method='nlminb')))
-
-sjPlot::tab_model(final_model, file = "Final Model Outputs/peak_pupil_model.html")
-
-```
-
-```{r, include = F, warning = F, message = F}
-
-# Removing unneeded items from the environment
-
-rm(data, m0, m1, m2, m3, m4, peak_pupil_des, peak_pupil_df, pupil_roi, my_pal)
-
-```

### 2024-02-06 MSC 2024 Analysis/pupil_data_prep.R
loaded raw pupil tracking

@@ -1,0 +1,93 @@
+# Pupil Data Preprocessing
+
+# Author: Micah E. Hirsch, mhirsch@fsu.edu 
+
+# Date: 2/6/2024
+
+## Purpose: To load in raw data from Eyelink and prepare pupil data viz and analysis.
+## This particular data processing and analysis is being completed for a poster 
+## presentation at the Biannual Motor Speech Conference being held in San Diego, 
+## CA in February 2024.
+
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+## Need remotes to install gazer and saccades packages
+library(remotes) # install.packages("remotes")
+library(gazer) # remotes::install_github("dmirman/gazer")
+library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
+library(zoo) # install.packages("zoo")
+library(knitr) # install.packages("knitr")
+
+# Set the working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Load raw pupil files
+
+## Get file names from raw data folder
+file_list <- list.files(path = ".", pattern = ".txt")
+
+## Create empty list to temporarily store imported participant data
+data_list <- list()
+
+## Initiating loop to load in data for each participant
+for (file in file_list) {
+  
+  ## Import file
+  data <- rio::import(file) 
+  
+  ## For subject id's ending in _2, we have to discard their first 15 trials
+  ## because the experiment was restarted for them after the 15th trial
+  ## (due to technical errors). The first 15 trials were recorded in 
+  ## the initial output file for the participant.
+  if (any(grepl("_2", data$RECORDING_SESSION_LABEL))) {
+    
+    data <- data |>
+      # Creating pupil and blink variables. 
+      # Most participants had their right eye tracked, but some had their left eye tracked.
+      dplyr::mutate(pupil = ifelse(EYE_TRACKED == "Right", RIGHT_PUPIL_SIZE, LEFT_PUPIL_SIZE),
+                    blink = ifelse(EYE_TRACKED == "Right", RIGHT_IN_BLINK, LEFT_IN_BLINK)) %>%
+      # Bring variable names to lower case
+      dplyr::rename_all(., .funs = tolower) |>
+      # Renaming variables so it is consistent with gazer's requirements
+      dplyr::rename(subject = recording_session_label,
+                    trial = trial_index) |>
[…]

### 2024-02-06 MSC 2024 Analysis/pupil_data_prep.R
Trimmed pupil dilation df

@@ -91,3 +91,45 @@ pupil_data <- do.call(rbind, data_list)
 ## Removing unneeded items from the environment
 rm(data, data_list, file, file_list)
 
+# Filtering out rows before the trial start and after the response cue
+
+## Extracting trial start times
+trial_start <- pupil_data |>
+  dplyr::filter(sample_message == "TRIAL_START") |>
+  dplyr::select(subject, trial, start_time = timestamp)
+
+## Extracting phrase start times
+phrase_start <- pupil_data |>
+  dplyr::filter(sample_message == "PHRASE_START") |>
+  dplyr::select(subject, trial, phrase_start = timestamp)
+
+## Extracting phrase end times
+phrase_end <- pupil_data |>
+  dplyr::filter(sample_message == "PHRASE_END") |>
+  dplyr::select(subject, trial, phrase_end = timestamp)
+
+## Extracting trial end times (e.g. time response cue was presented)
+trial_end <- pupil_data |>
+  dplyr::filter(sample_message == "RESPONSE_CUE") |>
+  dplyr::select(subject, trial, end_time = timestamp)
+
+## Merging the dfs together
+pupil_data2 <- pupil_data |>
+  dplyr::left_join(trial_start, by = c("subject", "trial")) |>
+  dplyr::left_join(phrase_start, by =c("subject", "trial")) |>
+  dplyr::left_join(phrase_end, by = c("subject", "trial")) |>
+  dplyr::left_join(trial_end, by = c("subject", "trial"))
+
+
+## Filter out unneeded rows and trials
+trimmed_pupil_data <- pupil_data2 %>%
+  ## Filtering out rows before trial start and after trial end
+  dplyr::filter(timestamp >= start_time & timestamp <= end_time) %>%
+  ## Removing practice trials from df
+  dplyr::filter(practicetrial != 'Practice') %>%
+  dplyr::select(!practicetrial) %>%
+  ## Aligning data to onset of phrase presentation
+  dplyr::mutate(time_c = timestamp - phrase_start) %>%
+  ## Removing unneeded variables
+  dplyr::select(!c(timestamp, start_time:end_time)) %>%
+  dplyr::relocate(time_c, .after = pupil)

### 2024-02-06 MSC 2024 Analysis/pupil_data_prep.R
cleaned and exported pupil data

@@ -122,14 +122,100 @@ pupil_data2 <- pupil_data |>
 
 
 ## Filter out unneeded rows and trials
-trimmed_pupil_data <- pupil_data2 %>%
+trimmed_pupil_data <- pupil_data2 |>
   ## Filtering out rows before trial start and after trial end
-  dplyr::filter(timestamp >= start_time & timestamp <= end_time) %>%
+  dplyr::filter(timestamp >= start_time & timestamp <= end_time) |>
   ## Removing practice trials from df
-  dplyr::filter(practicetrial != 'Practice') %>%
-  dplyr::select(!practicetrial) %>%
+  dplyr::filter(practicetrial != 'Practice') |>
+  dplyr::select(!practicetrial) |>
   ## Aligning data to onset of phrase presentation
-  dplyr::mutate(time_c = timestamp - phrase_start) %>%
+  dplyr::mutate(time_c = timestamp - phrase_start) |>
   ## Removing unneeded variables
-  dplyr::select(!c(timestamp, start_time:end_time)) %>%
+  dplyr::select(!c(timestamp, start_time:end_time)) |>
   dplyr::relocate(time_c, .after = pupil)
+
+## Removing trials/participants with too much data loss
+trimmed_pupil_data <- gazer::count_missing_pupil(trimmed_pupil_data, pupil = "pupil", missingthresh = 0.2)
+
+## Remove unneeded items from environment 
+rm(phrase_end, phrase_start, pupil_data, pupil_data2, trial_end, trial_start)
+
+# Fill in missing data from blinks
+
+## Deblinking
+pupil_extend <- trimmed_pupil_data |>
+  dplyr::group_by(subject, trial) |>
+  dplyr::mutate(extendpupil = extend_blinks(pupil, 
+                                            fillback = 50, 
+                                            fillforward = 160, 
+                                            hz = 1000))
+
+## linear interpolation
+interp <- interpolate_pupil(pupil_extend,
+                            extendblinks = T, 
+                            type = "linear", 
+                            hz = 1000)
+
+## 10 Hz 5-point moving average filter
+smoothed <- interp |>
+  dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) |>
+  ## Selecting relevant variables
+  dplyr::select(c(subject, trial, sample_message, time_c, effort_rating, 
[…]

### 2024-02-06 MSC 2024 Analysis/pupil_data_prep.R
small script updates

@@ -219,3 +219,4 @@ setwd("~/Documents/Listening-Effort-in-Dysarthria/MSC 2024 Analysis/Cleaned Data
 ## Export
 rio::export(data.binned, "cleaned_pupil_data.csv")
 
+

### 2024-02-07 MSC 2024 Analysis/pupil_data_prep.R
downsampling als trials - wip

@@ -204,11 +204,11 @@ rm(baseline_pupil, interp, pupil_extend, smoothed, trimmed_pupil_data, trial_che
 
 bin.length <- 20
 
-data.binned <- mad_removal %>%
-  mutate(timebins = round(time/bin.length)*bin.length) %>%
+data.binned <- mad_removal |>
+  mutate(timebins = round(time/bin.length)*bin.length) |>
   dplyr::group_by(subject, trial, speaker, timebins, effort_rating,
-                  code, targetphrase, counterbalance) %>%
-  dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) %>%
+                  code, targetphrase, counterbalance) |>
+  dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) |>
   dplyr::ungroup()
 
 # Export Cleaned Data
@@ -219,4 +219,35 @@ setwd("~/Documents/Listening-Effort-in-Dysarthria/MSC 2024 Analysis/Cleaned Data
 ## Export
 rio::export(data.binned, "cleaned_pupil_data.csv")
 
+# Downsampling ALS speaker trials
+## Based on findings from the 2023 ASHA Convention data analysis,the trials from the ALS speaker are much longer than the control talker.
+## Therefore we are creating a separate df that downsamples the ALS speaker's trials.
 
+bin.length <- 26.23
+
+ALS_trials <- data.binned |>
+  dplyr::filter(speaker == "ALS") |>
+  dplyr::mutate(time_n = round(timebins/bin.length)*bin.length) |>
+  dplyr::group_by(subject, trial, speaker, time_n, effort_rating, code, targetphrase, counterbalance) |>
+  dplyr::summarize(normed_pupil = mean(pupil.binned)) |>
+  dplyr::ungroup()
+
+control_trials <- data.binned |>
+  dplyr::filter(speaker == "Control") |>
+  dplyr::rename(time_n = timebins,
+                normed_pupil = pupil.binned)
+
+normed_data <- rbind(ALS_trials, control_trials)
+
+# This is not working as intended yet
+normed_data <- normed_data |>
+  dplyr::mutate(time_norm = time_n * 1.31)
+
+normed_data |>
+  dplyr::filter(time_norm >= 0) %>%
+  dplyr::group_by(speaker, code) %>%
+  dplyr::summarize(length = max(time_norm) - min(time_norm)) %>%
+  dplyr::group_by(speaker) %>%
[…]

### 2024-02-07 MSC 2024 Analysis/pupil_data_prep.R
normalized als trial lengths

@@ -241,10 +241,11 @@ normed_data <- rbind(ALS_trials, control_trials)
 
 # This is not working as intended yet
 normed_data <- normed_data |>
-  dplyr::mutate(time_norm = time_n * 1.31)
+  dplyr::mutate(time_norm = case_when(speaker == "ALS" ~ time_n/1.5,
+                                      TRUE ~ time_n))
 
 normed_data |>
-  dplyr::filter(time_norm >= 0) %>%
+  dplyr::filter(time_norm>= 0) %>%
   dplyr::group_by(speaker, code) %>%
   dplyr::summarize(length = max(time_norm) - min(time_norm)) %>%
   dplyr::group_by(speaker) %>%

### 2024-02-07 MSC 2024 Analysis/pupil_data_prep.R
exported normalized pupil dilation df

@@ -252,3 +252,7 @@ normed_data |>
   dplyr::summarize(av_length = mean(length),
                    av_phrase = av_length - 3000,
                    av_end_roi = av_length - 2000)
+
+# Export Normalized dataset
+
+rio::export(normed_data, "cleaned_pupil_data_normalized.csv")

### 2024-02-08 MSC 2024 Analysis/transcription_prep.R
prepped phrase accuracy variable

@@ -1,0 +1,125 @@
+# Transcription Data Preprocessing
+
+# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
+
+# Date: 2/8/2024 
+
+## Purpose: To load in transcriptions of the listener responses,
+## determine phrase recognition accuracy, and merge with the pupil data
+## (both raw and normalized pupil dfs). This data processing and analysis is 
+## being completed for a poster presentation at the Motor Speech Conference 
+## San Diego, CA in February 2024.
+
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Remotes is needed to download autoscore
+library(remotes) # install.packages("remotes")
+library(autoscore) # remotes::install_github("autoscore/autoscore")
+library(SnowballC) # install.packages("SnowballC")
+
+# Set working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Loading in raw transcriptions
+
+transcriptions <- rio::import("transcriptions.csv", header = T)
+
+# Preparing the script for scoring
+
+transcriptions <- transcriptions |>
+  # Removing unneeded columns generated by Qualtrics
+  dplyr::select(!c(StartDate:RA, Time, Counter_List)) |>
+  # Removing first two rows
+  dplyr::filter(grepl("LE", ID)) |>
+  dplyr::filter(!grepl("Participant", ID)) |>
+  # LE01 has multiple submissions. Only keeping the completed submission.
+  dplyr::filter(`85` != "") |>
+  # converting df to long format
+  tidyr::pivot_longer(cols = "1":"85",
+                      names_to = "trial",
+                      values_to = "response") |>
+  dplyr::rename(id = ID) |>
+  # formatting trial and response variables
+  dplyr::mutate(trial = as.numeric(trial),
+                response = tolower(response),
+                response = trimws(response, "both")) |>
+  # Removing first 5 trials since these are practice
+  dplyr::filter(!trial %in% (1:5)) |>
+  # Some audio files were cut off but not marked during transcriptions
+  dplyr::mutate(response = case_when(id == "LE21" & trial %in% c(20, 34, 41, 50, 54, 55, 81) ~ "missing data",
+                                     id == "LE22" & trial %in% c(25, 26, 47, 57, 62, 64) ~ "missing data",
[…]

### 2024-02-08 MSC 2024 Analysis/transcription_prep.R
Updated transcription prep

@@ -1,0 +1,127 @@
+# Transcription Data Preprocessing
+
+# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
+
+# Date: 2/8/2024 
+
+## Purpose: To load in transcriptions of the listener responses,
+## determine phrase recognition accuracy, and merge with the pupil data
+## (both raw and normalized pupil dfs). This data processing and analysis is 
+## being completed for a poster presentation at the Motor Speech Conference 
+## San Diego, CA in February 2024.
+
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Remotes is needed to download autoscore
+library(remotes) # install.packages("remotes")
+library(autoscore) # remotes::install_github("autoscore/autoscore")
+library(SnowballC) # install.packages("SnowballC")
+
+# Set working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Loading in raw transcriptions
+
+transcriptions <- rio::import("transcriptions.csv", header = T)
+
+# Preparing the script for scoring
+
+transcriptions <- transcriptions |>
+  # Removing unneeded columns generated by Qualtrics
+  dplyr::select(!c(StartDate:RA, Time, Counter_List)) |>
+  # Removing first two rows
+  dplyr::filter(grepl("LE", ID)) |>
+  dplyr::filter(!grepl("Participant", ID)) |>
+  # LE01 has multiple submissions. Only keeping the completed submission.
+  dplyr::filter(`85` != "") |>
+  # converting df to long format
+  tidyr::pivot_longer(cols = "1":"85",
+                      names_to = "trial",
+                      values_to = "response") |>
+  dplyr::rename(id = ID) |>
+  # formatting trial and response variables
+  dplyr::mutate(trial = as.numeric(trial),
+                response = tolower(response),
+                response = trimws(response, "both")) |>
+  # Removing first 5 trials since these are practice
+  dplyr::filter(!trial %in% (1:5)) |>
+  # Some audio files were cut off but not marked during transcriptions
+  dplyr::mutate(response = case_when(id == "LE21" & trial %in% c(20, 34, 41, 50, 54, 55, 81) ~ "missing data",
+                                     id == "LE22" & trial %in% c(25, 26, 47, 57, 62, 64) ~ "missing data",
[…]

### 2024-02-08 MSC 2024 Analysis/msc_analysis.Rmd
started msc analysis script

--- 2023-12-01 ASHA 2023 Analysis/standardized_pupil_analysis.Rmd
@@
 ---
-title: "Standardized Pupil Analysis"
+title: "Listening Effort MSC Analysis"
 author: "Micah E. Hirsch"
-date: "2023-12-01"
+date: "2024-02-08"
 output: html_document
@@
 
-The preliminary pupillometry results (found in the data_analysis.Rmd file) indicicated that the shape of the pupil dilation tracking is different while listening to the speaker with ALS verses listening to the control speaker.
-However, it is not clear whether this finidng is influenced by the difference in the length of the audio files for the control speaker and the ALS speaker.
-Due to the speaker's slowed speech rate, the ALS speaker's audio files were longer than the control speaker's files.
-
-Therefore, in order to control for the effect of the difference in speech rate, we will normalize the trial-level data for the ALS speaker so that the trial lengths are similar to the control speaker's trials.
-
-This reanalysis is conducted using R version 4.3.1
-
-```{r, include = F, warning = F, message = F}
+```{r, echo = F, warning = F, message = F}
 
@@
 
-# Pupil Dilation Normalization
-
-Work in Progress
-
 ```{r}
 
-# Import pupil dilation data
-data <- rio::import("Cleaned Data/cleaned_data.csv") %>%
-  # removing trials with missing rep accuracy: 3 trials removed
-  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) %>%
-  dplyr::filter(!is.na(rep_acc)) %>%
-  # Releveling speaker variable
-  dplyr::mutate(speaker = as.factor(speaker),
-                speaker = fct_relevel(speaker, "Control", "ALS")) %>%
-  # Creating a trial variable that codes trial 6 (i.e. the first experimental trial) as 0.
-  dplyr::mutate(trial_c = trial - 6)
-
-# Pupil Normalization
-
-bin.length = 2.29
-
-ALS_trials <- data %>%
-  dplyr::filter(speaker == "ALS") %>%
-  dplyr::mutate(time_n = round(timebins/bin.length)*bin.length) %>%
-  dplyr::group_by(subject, trial_c, speaker, time_n, code) %>%
-  dplyr::summarize(normed_pupil = mean(pupil.binned)) %>%
-  dplyr::ungroup()
+plot(pressure)
 
 ```
-
-```{r}
-
-data %>%
-  # Selecting relevant parameters
-  dplyr::select(subject, trial, speaker, timebins, code) %>%
-  # Filtering out times before phase onset
-  dplyr::filter(timebins >= 0) %>%
-  dplyr::group_by(speaker, code) %>%
-  dplyr::summarize(length = max(timebins) - min(timebins)) %>%
-  dplyr::group_by(speaker) %>%
-  dplyr::summarize(av_length = mean(length),
-                   av_phrase = av_length - 3000,
-                   av_end_roi = av_length - 2000) %>%
-  gt::gt() %>%
-  gt::cols_label(speaker = "Speaker",
-                 av_length = "Average Trial Length (ms)",
-                 av_phrase = "Average Phrase Length (ms)",
-                 av_end_roi = "Average Endpoint for ROI (ms)")
-
-```

### 2024-02-08 MSC 2024 Analysis/listener_demo_prep.R
processing listener demo data

@@ -1,0 +1,76 @@
+# Listener Demographics Processing
+
+# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
+
+# Date: 2/8/2024
+
+## Purpose: To load and clean listener demographic information 
+## and export a cleaned version of the data. This preprocessing was completed
+## for a poster presentation at the Motor Speech Conference held in San Diego, CA
+## in February 2024.
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+
+# Load in listener demographic data from RedCap
+
+## Set working directory
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+demo <- rio::import("participant_demo_raw.csv")
+
+## NOTE: LE03 withdrew from the study. Their data was removed from 
+## RedCap prior to uploading the df here.
+
+# Cleaning demographic data
+
+demo1 <- demo |>
+  # Removing identifiable and unneeded variables from df
+  dplyr::select(!c(date_participated, dob, sex, fluent_eng, demographic_information_complete)) %>%
+  # Recoding variables so they are easily interpretable
+  dplyr::mutate(gender = case_when(gender == 1 ~ "Man",
+                                   gender == 2 ~ "Woman",
+                                   gender == 3 ~ "Nonbinary", 
+                                   gender == 4 ~ "Questioning",
+                                   gender == 5 ~ "Prefer not to answer",
+                                   TRUE ~ "Gender not listed"),
+                ethnicity = case_when(ethnicity == 1 ~ "Hispanic/Latino(a/e)",
+                                      ethnicity == 2 ~ "Not Hispanic/Latino(a/e)",
+                                      TRUE ~ "Prefer not to answer"),
+                race = case_when(race == 1 ~ "white/Caucasian",
+                                 race == 2 ~ "Black/African American",
+                                 race == 3 ~ "Asian/Asian American",
+                                 race == 4 ~ "Native Hawaiian or Other Pacific Islander",
+                                 race == 5 ~ "Native American or Alaska Native",
+                                 race == 6 ~ "Biracial or Multiracial",
+                                 race == 7 ~ "Prefer not to answer",
+                                 TRUE ~ "Race not listed"),
+                native_lang = ifelse(native_lang == 1, "American English", "Not American English"),
+                commdis_hx = ifelse(commdis_hx == 1, "no", "yes"),
+                fam_commdis = ifelse(fam_commdis == 1, "no", "yes"),
[…]

### 2024-02-08 MSC 2024 Analysis/msc_analysis.Rmd
Update msc_analysis.Rmd
listener demographics

@@
 
+This analysis was conducted using R version 4.3.1
+
 ```{r, echo = F, warning = F, message = F}
@@
 library(gt) # install.packages("gt")
+library(gtsummary) # install.packages("gtsummary")
 library(glmmTMB) # install.packages("glmmTMB")
@@
 
+# Listener Demographics
+
 ```{r}
 
-plot(pressure)
+listener_demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv") |>
+  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
+                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
+                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
+                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
+                                               "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
+                native_lang = factor(native_lang, c("American English", "Not American English")))
+
+demo_table <- listener_demo |>
+  dplyr::select(age, gender, ethnicity, race, native_lang) |>
+  tbl_summary(type = list(age ~ "continuous",
+                          gender ~ "categorical",
+                          ethnicity ~ "categorical",
+                          race ~ "categorical",
+                          native_lang ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd}, {min}-{max})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(age ~ "Age",
+                           gender ~ "Gender",
+                           ethnicity ~ "Ethnicity",
+                           race ~ "Race",
+                           native_lang ~ "Native Language")) |>
+  as_gt()
+
+demo_table
+
+demo_table |>
+  gt::gtsave("Tables/demo_tabe.html")
 

### 2024-02-12 MSC 2024 Analysis/transcription_prep.R
fixed transcription prep script

@@ -30,6 +30,9 @@ transcriptions <- rio::import("transcriptions.csv", header = T)
 # Preparing the script for scoring
 
 transcriptions <- transcriptions |>
+  # Fixing a data entry error 
+  dplyr::mutate(ID = case_when(ResponseId == "R_3l45R7NRdSEXrTr" ~ "LE24",
+                               TRUE ~ ID)) |>
   # Removing unneeded columns generated by Qualtrics
   dplyr::select(!c(StartDate:RA, Time, Counter_List)) |>
   # Removing first two rows

### 2024-02-12 MSC 2024 Analysis/msc_analysis.Rmd
small rmd edits
calculated mean speaker intelligibility

@@
 ```
+
+```{r}
+
+rm(demo_table, listener_demo)
+
+```
+
+# Speaker Intelligibility and Phrase Accuracy
+
+## Mean Speaker Intelligibility
+
+```{r}
+
+data <- rio::import("Cleaned Data/cleaned_data_normed.csv") |>
+  # removing trials with missing rep accuracy
+  dplyr::mutate(rep_acc = ifelse(rep_acc == "", NA, rep_acc)) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                trial_c = trial - 6,
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")))
+
+speaker_acc <- data |>
+  dplyr::select(subject, trial, speaker, correct_words, target_number, rep_acc) |>
+  dplyr::distinct() |>
+  dplyr::mutate(intel = (correct_words/target_number)*100) |>
+  dplyr::select(-c(subject, trial, correct_words, target_number)) |>
+  tbl_summary(by = speaker,
+              type = list(intel ~ "continuous",
+                          rep_acc ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(intel ~ "Intelligibility",
+                           rep_acc ~ "Repetition Accuracy")) |>
+  as_gt()
+
+speaker_acc
+
+speaker_acc |>
+  gt::gtsave("Tables/speaker_intel.html")
+
+```

### 2024-02-12 MSC 2024 Analysis/msc_analysis.Rmd
created data viz and perceived listening effort analysis

@@
 
-```{r}
+```{r, echo = F, warning = F, message = F}
 
@@
 ```
+
+```{r, echo = F, warning = F, message = F}
+
+rm(speaker_acc)
+
+```
+
+# Perceived Listening Effort
+
+## Descriptives
+
+```{r}
+
+perceived_effort_df <- data |>
+  dplyr::select(subject, trial, speaker, effort_rating, rep_acc, code, trial_c) |>
+  dplyr::distinct() |>
+  # recoding missing effort ratings (negative effort ratings indicate missing data)
+  dplyr::mutate(effort_rating = ifelse(effort_rating < 0, NA, effort_rating))
+
+perceived_effort_df |>
+  dplyr::select(speaker, effort_rating, rep_acc) |>
+  tbl_strata(
+    strata = rep_acc,
+    ~.x |>
+      tbl_summary(
+        by = speaker,
+        type = list(effort_rating ~ "continuous"),
+        statistic = list(all_continuous() ~ c("{mean} ({sd})")),
+        missing = "no",
+        digits = all_continuous() ~ 2,
+        label = list(effort_rating ~ "Perceived Listening Effort")
+      ))
+
+```
+
+## Visualization
+
+```{r}
+
+# Creating Custom Color Palette for Figures
+library(NatParksPalettes) # install.packages("NatParksPalettes")
+my_pal <- natparks.pals("Glacier", type = "discrete")
+
+# Getting Sum stats for perceived effort
+
+per_effort_des <- perceived_effort_df |>
+  dplyr::group_by(speaker, rep_acc) |>
+  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
+                   sd = sd(effort_rating, na.rm = T),
+                   se = sd/sqrt(n()))
+
+# Plotting the mean perceived effort ratings for the ALS and control speaker
+per_effort_des |>
+  ggplot() +
+   aes(x = rep_acc,
+       y = per_effort,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
+  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
+  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
+  theme_bw() +
+  theme(aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[3])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[3]))
+
+# Saving the ggplot
+
+ggsave("Plots/Perceived_effort.png", plot = last_plot(), height = 5, width = 5, units = "in")
+
+```
+
+## Perceived Listening Effort Models
+
+### Fully Unconditional Model
+
+```{r}
+
+m0 <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = perceived_effort_df)
+
+sjPlot::tab_model(m0, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+performance::icc(m0)
+
+```
+### Model 1
+
+```{r}
+
+m1 <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m1, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m0, m1)
+
+```
+
+### Model 2
+
+```{r}
+
+m2 <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), 
+              data = perceived_effort_df)
+
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m1, m2)
+
+```
+### Model 3
+
+```{r}
+
+m3 <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m2, m3)
+
+```
+### Model 4
+
+```{r}
+
+m4 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m3, m4)
+
+```
+### Model 5
+
+```{r}
+
+m5 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m5, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m4, m5)
+
+```
+### Model 6
+
+```{r}
+
+m6 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + speaker*trial_c + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m6, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m5, m6)
+
+```
+
+### Model 7
+
+```{r}
+
+m7 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                speaker*rep_acc + speaker*trial_c + rep_acc*trial_c + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m7, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Trial Order"),
+                  dv.labels = "Perceived Listening Effort")
+
+anova(m6, m7)
+
+```
+
+### Model 8
+
+```{r}
+
+m8 <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + 
+                speaker*rep_acc + speaker*trial_c + rep_acc*trial_c + speaker*rep_acc*trial_c + (trial_c|subject) + (trial_c|code), 
+           data = perceived_effort_df)
+
+sjPlot::tab_model(m8, pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Phrase Repetition Accuracy [Inaccurate]",
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Trial Order",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Trial Order", 
+                                      "Phrase Repetition Accuracy * Speaker * Trial Order"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/perceived_effort_model.html")
+
+anova(m7, m8)
+
+```
+
+```{r, echo = F, message = F, warning = F}
+
+rm(m0, m1, m2, m3, m4, m5, m6, m7, m8, per_effort_des)
+
+```
+
+# Pupil Dilation
+
+## Preliminary Visualizations
+
+Needing to spot any participants with odd trackings. Based on the visualizations below, LE13 and LE23 appear to have odd trackings, so they will be dropped from the current analyses.
+
+```{r, warning = F, message = F}
+
+data |>
+  ggplot() +
+  aes(x = time_norm,
+      y = normed_pupil) +
+  geom_smooth() +
+  coord_cartesian(xlim = c(0,4908), ylim = c(-100, 200)) +
+  facet_wrap("subject")
+
+```
+
+## Descriptives
+
+### Getting Time Landmarks
+
+```{r, message = F, warning = F}
+
+data <- data |>
+  dplyr::filter(subject != "LE13" & subject != "LE23")
+
+data |>
+  # Selecting relevant parameters
+  dplyr::select(subject, trial, speaker, time_norm, code) |>
+  # Filtering out times before phase onset
+  dplyr::filter(time_norm >= 0) |>
+  dplyr::group_by(speaker, code) |>
+  dplyr::summarize(length = max(time_norm) - min(time_norm)) |>
+  dplyr::ungroup() |>
+  dplyr::summarize(av_length = mean(length),
+                   av_phrase = av_length - 3000,
+                   av_end_roi = av_length - 2000) %>%
+  gt::gt() |>
+  gt::cols_label(av_length = "Average Trial Length (ms)",
+                 av_phrase = "Average Phrase Length (ms)",
+                 av_end_roi = "Average Endpoint for ROI (ms)")
+
+```
+
+## Visualizations
+
+```{r, warning = F, message = F}
+
+data |>
+  ggplot() +
+  aes(x = time_norm,
+      y = normed_pupil,
+      group = speaker,
+      color = speaker) +
+  geom_smooth() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -50, ymax = 200, alpha = 0.2) +
+  coord_cartesian(xlim = c(0,4910), ylim = c(-100, 200)) +
+  labs(x = "Normalized Time (ms)", y = "Pupil Dilation (arbitrary units)",
+       caption = "Grey Box: Analysis Region of Interest \nSolid Line: Phrase Onset \nDotted Line: Phrase Offset") +
+  facet_wrap("rep_acc") + 
+  theme_bw() +
+  scale_color_manual(values = c(my_pal[1], my_pal[3])) +
+  theme(aspect.ratio = 1,
+        plot.caption = element_text(hjust = 0))
+
+ggsave("Plots/Pupil_dilation.png", plot = last_plot(), width = 10, height = 5, units = "in")
+
+```

### 2024-02-12 MSC 2024 Analysis/msc_analysis.Rmd
pupil dilation analysis

@@
 ```
+
+## Pupil Dilation Models
+
+### Unconditional Growth Curve Model
+
+```{r}
+
+# Getting region of interest for analysis
+data_roi <- data |>
+  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
+  dplyr::filter(time_norm >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  # filtering out timebins 1000 ms after phrase offset
+  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
+  ungroup()
+
+# Adding linear, quadratic, and cubic orthogonal polynomials to the df
+data_roi <- gazer::code_poly(data_roi, predictor = "time_norm", poly.order = 3, orthogonal = T, draw.poly = F)
+
+# Unconditional Growth Model
+m0 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 +
+                (poly1 + poly2 + poly3|subject) + (poly1 + poly2 + poly3|code), 
+              data = data_roi)
+
+sjPlot::tab_model(m0, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term"),
+                  dv.labels = "Pupil Dilation")
+
+performance::icc(m0)
+
+```
+### Model 1
+
+```{r}
+
+# Running into convergence issues, so removed the random effects for phrase (i.e. code)
+m1 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m1, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m0, m1)
+
+```
+
+### Model 2
+
+```{r}
+
+m2 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m2, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m1, m2)
+
+```
+### Model 3
+
+```{r}
+
+m3 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m3, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m2, m3)
+
+```
+
+### Model 4
+
+```{r}
+
+m4 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m4, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m3, m4)
+
+```
+### Model 5
+
+```{r}
+
+m5 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m5, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m4, m5)
+
+```
+### Model 6
+
+```{r}
+
+m6 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m6, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m5, m6)
+
+```
+### Model 7
+
+```{r}
+
+m7 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 +
+                (poly1 + poly2 + poly3|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m7, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m6, m7)
+
+```
+
+### Model 8
+
+```{r}
+
+m8 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m8, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m7, m8)
+
+```
+
+### Model 9
+
+```{r}
+
+m9 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m9, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m8, m9)
+
+```
+### Model 10
+
+```{r}
+
+m10 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
+                speaker*rep_acc*poly1 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m10, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
+                                      "Speaker * Phrase Repetition Accuracy * Linear Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m9, m10)
+
+```
+### Model 11
+
+```{r}
+
+m11 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
+                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m11, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
+                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
+                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m10, m11)
+
+```
+### Model 12
+
+```{r}
+
+m12 <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
+                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 + speaker*rep_acc*poly3 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(m12, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
+                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
+                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term",
+                                      "Speaker * Phrase Repetition Accuracy * Cubic Term"),
+                  dv.labels = "Pupil Dilation")
+
+anova(m11, m12)
+
+```
+
+### Final Model
+
+```{r}
+
+final_model <- glmmTMB(normed_pupil ~ poly1 + poly2 + poly3 + speaker + rep_acc +
+                speaker*poly1 + speaker*poly2 + speaker*poly3 + speaker*rep_acc +
+                rep_acc*poly1 + rep_acc*poly2 + rep_acc*poly3 +
+                speaker*rep_acc*poly1 + speaker*rep_acc*poly2 +
+                (poly1 + poly2|subject), 
+              data = data_roi)
+
+sjPlot::tab_model(final_model, pred.labels = c("Intercept", "Linear Term", "Quadratic Term", "Cubic Term", "Speaker [ALS]",
+                                      "Phrase Repetition Accuracy [Inaccurate]", "Speaker [ALS] * Linear Term",
+                                      "Speaker [ALS] * Quadratic Term", "Speaker [ALS] * Cubic Term", 
+                                      "Speaker [ALS] * Phrase Repetition Accuracy [Inaccurate]", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Linear Term", 
+                                      "Phrase Repetition Accuracy [Inaccurate] * Quadratic Term",
+                                      "Phrase Repetition Accuracy [Inaccurate] * Cubic Term",
+                                      "Speaker * Phrase Repetition Accuracy * Linear Term",
+                                      "Speaker * Phrase Repetition Accuracy * Quadratic Term"),
+                  dv.labels = "Pupil Dilation", 
+                  file = "Tables/pupil_dilation_model.html")
+
+```

### 2024-02-14 MSC 2024 Analysis/msc_analysis.Rmd
cleaning up rmd file
knit rmd document
updated rmd report

@@
 author: "Micah E. Hirsch"
-date: "2024-02-08"
-output: html_document
+date: "2024-02-14"
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
 ---
@@
 # Purpose
+
+The purpose of this document is to conduct data analyses for the Listening Effort during Dysarthric Speech Perception study.
+These results will be presented at the 2024 Motor Speech Conference in February held in San Diego, CA.
 
@@
 # Listener Demographics
+
+The demographic information for our listeners are presented below.
+To date, we recruited 27 participants.
+However, 3 participants were removed from the analysis.
+One of these listeners decided to withdraw from the study, so their data was removed from the database.
+For the other two participants, pupillometry data was either not recorded or loss due to instrumentation error.
+Therefore, we have 24 total participants for this analysis.
 
@@
 
-## Mean Speaker Intelligibility
+During the experiment, participants listened to short phrases spoken by two speakers, one speaker with ALS and another speaker with no neurological impairment (control).
+The mean intelligibility for the speakers was calculated based on the listener responses.
+The proportion of accurate phrase repetitions (one of our predictors of interest) is also presented below for the speakers.
+Overall, the ALS speaker's dysarthria severity is determined to be moderate.
 
@@
 
+Descriptive statistics and figures showing perceived listening effort ratings are shown below.
+These ratings were collected after each trial.
+Listeners rated PLE on a 7-point Likert scale where higher numbers indicate more perceived effort.
+
 ## Descriptives
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
+To examine perceived listening effort, we are using linear mixed effects (LME) models.
+
 ### Fully Unconditional Model
 
-```{r}
+To start, we are fitting a fully unconditional model to obtain the intraclass correlation coefficient (ICC).
+The model only contains random intercepts for listener and phrase (code).
+The conditional and unconditional ICC is .149.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 1
 
-```{r}
+In model 1, we added trial order as a random slope for both listener and phrase random intercepts.
+A likelihood ratio test was also conducted to test whether model 1 provides a better fit to the data.
+The LRT indicates model 1 provides a better fit to the data.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+In model 2, we added trial order as a fixed effect to the model to test whether PLE ratings changed across trials in our study. Trial order did not significantly improve the model.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+Speaker was added as a fixed effect to model 3.
+The results indicate that including speaker as a fixed effect improved the model fit to the data.
+The estimate indicates that on average, listeners rated PLE the ALS speaker higher than the control speaker.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 4
 
-```{r}
+In model 4, phrase repetition accuracy was added as a fixed effect. This addition significantly improved the model fit.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 5
 
-```{r}
+An interaction term between speaker and phrase repetition accuracy was added to the model. This interaction did not significantly improve model fit.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 6
 
-```{r}
+The interaction between speaker and trial order was added to the model 6. This interaction did not significantly improve model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+In model 7, the interaction between phrase repetition accuracy and trial order was added as a fixed effect. This interaction did not significantly improve model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 8, a three-way interaction between speaker, phrase repetition accuracy, and trial order was added to the model. This 3-way interaction significantly improved model fit.
+
+Therefore, we will retain model 8 as our final model of perceived listening effort.
+The results indicate that on average, controlling for trial order and repetition accuracy, the ALS speaker was rated as more effortful to listen to than the control speaker.
+The significant interaction between speaker and phrase repetition accuracy indicates that the difference between PLE ratings when listeners accurately recognized the ALS speaker's phrase compared to when they didn't.
+The significant three way interaction suggests that on average, PLE ratings for the ALS speaker when the listener did not accurately recognize the phrase decreased slightly across trials.
+
+```{r, warning = F, message = F}
 
@@
 
-Needing to spot any participants with odd trackings. Based on the visualizations below, LE13 and LE23 appear to have odd trackings, so they will be dropped from the current analyses.
+Before starting the pupil dilation analyses, we need to spot any participants with odd pupil dilation tracking patterns and remove them from the analysis.
+Based on the visualizations below, LE13 and LE23 appear to have odd trackings, so they will be dropped from the current analyses.
 
@@
 ## Descriptives
+
+Trials from the ALS speaker were longer than trials from the control speaker.
+Therefore, in the data processing steps, the ALS speaker's trials were downsampled to a time scale similar to the control speakers (see the pupil_data_prep script for more info).
+For visualization purposes, we are extracting the major time landmarks below after phrase onset (i.e.
+not including the 3000 ms baseline period).
 
@@
 
+For the pupil dilation models, we are using a growth curve analysis to analyze the pupil dilation tracking.
+
 ### Unconditional Growth Curve Model
 
-```{r}
+First, an unconditional growth curve model was ran to get the ICC for the model.
+This initial model contains just the linear, quadratic, and cubic growth terms as fixed effects and random intercepts for listener and phrase.
+The growth terms were also added as random slopes.
+The adjusted ICC is .19 and unadjusted ICC is .186.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 1
 
-```{r}
+In model 1, we added speaker as a fixed effect to the model.
+At first, I ran into model convergence issues with having both listener and phrase random effects.
+Therefore, I decided to remove the phrase random intercepts and slopes from the model.
+Model comparisons were conducted using likelihood ratio tests.
+The LRT results indicate that speaker significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+Phrase repetition accuracy was added to the model as a fixed effect and significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 3
 
-```{r}
+For model three, the interaction between speaker and the linear growth term was added to the model. This interaction significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 4, the interaction between speaker and the quadratic growth term was added to the model. This interaction significantly improved the model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 5, the interaction between speaker and the cubic growth term was added. This interaction significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+The interaction between speaker and phrase repetition accuracy was added to model 6. This interaction significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 7, the interaction between phrase repetition accuracy and the linear growth term was added to the model. This interaction significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+In model 8, the interaction between phrase repetition accuracy and the quadratic growth term was added.
+At the first attempt at running this model, there were model convergence issues.
+I initially tried to run the model with increased iterations, and then changing the optimizer used.
+However, these changes did not resolve convergence issues.
+Therefore, the random effect structure was simplified by removing the cubic random slope.
+The addition of the interaction signficiantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 9, the interaction between phrase repetition accuracy and the cubic growth term was added. This interaction term did improve the model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 10, the three-way interaction between speaker, phrase repetition accuracy, and the linear growth term was added to the model. This interaction did not significantly improve model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+For model 11, the three-way interaction between speaker, repetition accuracy, and the quadratic growth term was added to the model. This interaction significantly improved model fit.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
 ### Model 12
 
-```{r}
+For model 12, the final three-way interaction between speaker, repetition accuracy, and the cubic growth term was added to the model. This interaction did not significantly improve model fit.
+
+```{r, warning = F, message = F}
 
@@
 
-```{r}
+Based on this model building approach, the model that provided the best fit to the data was model 11. Therefore, the final model is presented below.
+
+```{r, warning = F, message = F}
 
@@
 ```
+
+```{r, echo = F, warning = F, message = F}
+
+rm(m0, m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, m12)
+
+```

### 2024-02-14 MSC 2024 Analysis/transcription_prep.R

@@ -30,6 +30,9 @@ transcriptions <- rio::import("transcriptions.csv", header = T)
 # Preparing the script for scoring
 
 transcriptions <- transcriptions |>
+  # Fixing a data entry error 
+  dplyr::mutate(ID = case_when(ResponseId == "R_3l45R7NRdSEXrTr" ~ "LE24",
+                               TRUE ~ ID)) |>
   # Removing unneeded columns generated by Qualtrics
   dplyr::select(!c(StartDate:RA, Time, Counter_List)) |>
   # Removing first two rows
@@ -88,8 +91,10 @@ phrase_acc <- autoscore::autoscore(
   tense_rule = T,
   tense_add_rule = T,
   a_the_rule = T,
-  double_letter_rule = T) |>
-  dplyr::rename(., correct_words = autoscore)
+  double_letter_rule = T) 
+
+phrase_acc <- phrase_acc |>
+  dplyr::rename(correct_words = autoscore)
 
 # Creating dichotomous phrase repetition accuracy variable
 

### 2024-04-15 Manuscript Analysis/pupil_data_prep.R
fixed issue with extracting phrase start times

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, mhirsch@fsu.edu
 
-## Data: 3/25/2024
+## Data: 4/15/2024
 
 ## Purpose: To prepare the pupil dilation data for analysis.
 
@@ -92,22 +92,22 @@ rm(data, data_list, file, file_list)
 
 ## Extracting trial start times
 trial_start <- pupil_data |>
-  dplyr::filter(sample_message == "TRIAL_START") |>
+  dplyr::filter(grepl("TRIAL_START", sample_message)) |>
   dplyr::select(subject, trial, start_time = timestamp)
 
 ## Extracting phrase start times
 phrase_start <- pupil_data |>
-  dplyr::filter(sample_message == "PHRASE_START") |>
+  dplyr::filter(grepl("PHRASE_START", sample_message)) |>
   dplyr::select(subject, trial, phrase_start = timestamp)
 
 ## Extracting phrase end times
 phrase_end <- pupil_data |>
-  dplyr::filter(sample_message == "PHRASE_END") |>
+  dplyr::filter(grepl("PHRASE_END", sample_message)) |>
   dplyr::select(subject, trial, phrase_end = timestamp)
 
 ## Extracting trial end times (e.g. time response cue was presented)
 trial_end <- pupil_data |>
-  dplyr::filter(sample_message == "RESPONSE_CUE") |>
+  dplyr::filter(grepl("RESPONSE_CUE", sample_message)) |>
   dplyr::select(subject, trial, end_time = timestamp)
 
 ## Merging the dfs together 

### 2024-04-15 Manuscript Analysis/pupil_data_prep.R
Exported updated pupil data

@@ -139,3 +139,109 @@ rm(pupil_data)
 
 trimmed_pupil_data <- gazer::count_missing_pupil(trimmed_pupil_data, missingthresh = 0.2)
 
+# Fill in missing data from blinks
+
+## Deblinking
+pupil_extend <- trimmed_pupil_data |>
+  dplyr::group_by(subject, trial) |>
+  dplyr::mutate(extendpupil = extend_blinks(pupil, 
+                                            fillback = 50, 
+                                            fillforward = 160, 
+                                            hz = 1000))
+
+## Linear interpolation
+interp <- interpolate_pupil(pupil_extend,
+                            extendblinks = T, 
+                            type = "linear", 
+                            hz = 1000)
+
+## 10 Hz 5-point moving average filter
+smoothed <- interp |>
+  dplyr::mutate(smoothed_pupil = moving_average_pupil(interp, n = 5)) |>
+  ## Selecting relevant variables
+  dplyr::select(c(subject, trial, sample_message, time, effort_rating, 
+                  code, speaker, targetphrase, counterbalance,smoothed_pupil)) |>
+  dplyr::relocate(smoothed_pupil, .after = time)
+
+# Baseline Pupil Correction
+
+baseline_pupil <- baseline_correction_pupil(smoothed, pupil_colname = "smoothed_pupil",
+                                            baseline_window = c(-500, 0))   
+
+# Artifact Rejection
+
+## Looking for rapid changes in pupil dilation using median absolute deviation
+mad_removal <- baseline_pupil |>
+  dplyr::group_by(subject, trial) |>
+  dplyr::mutate(speed = speed_pupil(baselinecorrectedp, time)) |>
+  dplyr::mutate(MAD = calc_mad(speed, n=16)) |>
+  dplyr::filter(speed < MAD)
+
+## Proportion of rows removed (as of 4/15/2024: 1.45%)
+((nrow(baseline_pupil) - nrow(mad_removal)) / nrow(baseline_pupil)) * 100
+
+## Checking to see if whole trials were removed from any of the participants (No Trials Removed)
+trial_check <- mad_removal |>
+  select(subject, trial) |>
+  distinct() |>
+  group_by(subject) |>
+  summarize(n = n())
+
+## Removing unneeded items from the environment
+rm(baseline_pupil, interp, pupil_extend, smoothed, trimmed_pupil_data, trial_check)
+
[…]

### 2024-04-15 Manuscript Analysis/transcription_prep.R
transcription processing
some listeners have not been transcribed yet

@@ -1,0 +1,167 @@
+# Transcription Data Preprocessing
+
+# Author: Micah E. Hirsch, M.S., mhirsch@fsu.edu
+
+# Date: 4/15/2024 
+
+## Purpose: To load in transcriptions of the listener responses,
+## determine phrase recognition accuracy, and merge with the pupil data
+## (both raw and normalized pupil dfs). 
+
+# Loading Required Packages
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+# Remotes is needed to download autoscore
+library(remotes) # install.packages("remotes")
+library(autoscore) # remotes::install_github("autoscore/autoscore")
+library(SnowballC) # install.packages("SnowballC")
+
+# Set working directory to load data
+
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+# Loading in raw transcriptions
+
+transcriptions <- rio::import("transcriptions.csv", header = T)
+
+# Preparing the script for scoring
+
+transcriptions <- transcriptions |>
+  # Fixing a data entry error 
+  dplyr::mutate(ID = case_when(ResponseId == "R_3l45R7NRdSEXrTr" ~ "LE24",
+                               TRUE ~ ID)) |>
+  # Removing unneeded columns generated by Qualtrics
+  dplyr::select(!c(StartDate:UserLanguage, Counter_List)) |>
+  # Removing first two rows
+  dplyr::filter(grepl("LE", ID)) |>
+  dplyr::filter(!grepl("Participant", ID)) |>
+  # LE01 has multiple submissions. Only keeping the completed submission.
+  dplyr::filter(`85` != "") 
+  
+initial <- transcriptions |>
+  # selecting initial transcriptions only
+  dplyr::filter(Time == "Initial Transcription") |>
+  dplyr::select(!RA) |>
+  # converting df to long format
+  tidyr::pivot_longer(cols = "1":"85",
+                      names_to = "trial",
+                      values_to = "response") |>
+  dplyr::rename(id = ID) |>
+  # formatting trial and response variables
+  dplyr::mutate(trial = as.numeric(trial),
+                response = tolower(response),
+                response = trimws(response, "both")) |>
+  # Removing first 5 trials since these are practice
+  dplyr::filter(!trial %in% (1:5)) |>
+  # Some audio files were cut off but not marked during transcriptions
[…]

### 2024-04-15 Manuscript Analysis/main_analysis.Rmd
created analysis script

---
title: "Main Analysis"
author: "Micah E. Hirsch"
date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to analyze and report findings related to M. Hirsch's dissertation project.

```{r}

library(rio)
library(tidyverse)
library(mgcv)
library(itsadug)
library(irr)

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv")

```

# Transcription Reliability

```{r}

reliability <- phrase_acc |>
  dplyr::filter(subject != "LE38" & subject != "LE33") |>
  dplyr::filter(!is.na(correct_words_rel)) |>
  dplyr::select(rep_acc, rep_acc_rel)

# Computing Fleiss' Kappa

kappam.fleiss(reliability, detail = T)

```

# Pupil Dilation

```{r}

pupil_data <- rio::import("Cleaned Data/cleaned_pupil_data.csv")

pupil <- pupil_data |>
  left_join(phrase_acc, by = c("subject", "trial", "targetphrase")) |>
  dplyr::select(!c(rep_acc_rel, correct_words_rel, rel_response)) |>
  dplyr::mutate(speaker = ifelse(speaker == "Control", 0, 1),
                rep_acc = ifelse(rep_acc == "accurate", 0, 1),
                speaker = factor(speaker, levels = c(0, 1), labels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c(0, 1), labels = c("accurate", "inaccurate")),
                code = as.factor(code))

pupil1 <- pupil |>
  dplyr::filter(rep_acc != "")

```

### 2024-04-18 Manuscript Analysis/transcription_prep.R
Updated transcription accuracy script

@@ -83,8 +83,6 @@ setwd("~/Documents/Listening-Effort-in-Dysarthria/Manuscript Analysis/Cleaned Da
 
 pupil <- rio::import("cleaned_pupil_data.csv")
 
-pupil_normed <- rio::import("cleaned_pupil_data_normalized.csv")
-
 ## Getting phrase targets
 targets <- pupil |>
   dplyr::rename(id = subject,

### 2024-04-18 Manuscript Analysis/listener_demo_prep.R
Prepped listener demographic information

@@ -1,0 +1,78 @@
+# Participant Demographic Data Prep
+
+# Author: Micah Hirsch, mhirsch@fsu.edu
+
+# Date: 4/18/2024
+
+## Purpose: To clean exported listener demographic data from RedCap and to
+## prepare it for analysis.
+
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+
+# Load in listener demographic data from RedCap
+
+## Set working directory
+setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+
+demo <- rio::import("participant_demo_raw.csv")
+
+## NOTE: LE03 withdrew from the study. Their data was removed from 
+## RedCap prior to uploading the df here.
+
+# Cleaning demographic data
+
+demo1 <- demo |>
+  # Removing identifiable and unneeded variables from df
+  dplyr::select(!c(fluent_eng, demographic_information_complete)) %>%
+  # Re-coding variables so they are easily interpretable
+  ## NOTE: A coding error from previous file versions was fixed down below. 
+  ## This is the correct version.
+  dplyr::mutate(gender = case_when(gender == 1 ~ "Man",
+                                   gender == 2 ~ "Woman",
+                                   gender == 3 ~ "Nonbinary", 
+                                   gender == 4 ~ "Questioning",
+                                   gender == 5 ~ "Prefer not to answer",
+                                   TRUE ~ "Gender not listed"),
+                ethnicity = case_when(ethnicity == 1 ~ "Hispanic/Latino(a/e)",
+                                      ethnicity == 2 ~ "Not Hispanic/Latino(a/e)",
+                                      TRUE ~ "Prefer not to answer"),
+                race = case_when(race == 1 ~ "white/Caucasian",
+                                 race == 2 ~ "Black/African American",
+                                 race == 3 ~ "Asian/Asian American",
+                                 race == 4 ~ "Native Hawaiian or Other Pacific Islander",
+                                 race == 5 ~ "Native American or Alaska Native",
+                                 race == 6 ~ "Biracial or Multiracial",
+                                 race == 7 ~ "Prefer not to answer",
+                                 TRUE ~ "Race not listed"),
+                native_lang = ifelse(native_lang == 1, "American English", "Not American English"),
+                commdis_hx = ifelse(commdis_hx == 1, "no", "yes"),
+                fam_commdis = ifelse(fam_commdis == 1, "no", "yes"),
[…]

### 2024-05-13 Manuscript Analysis/pupil_data_prep.R
adding outlier flags - wip

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, mhirsch@fsu.edu
 
-## Data: 5/7/2024
+## Data: 5/13/2024
 
 ## Purpose: To prepare the pupil dilation data for analysis.
 
@@ -276,7 +276,78 @@ baseline_flags <- mad_removal |>
   dplyr::distinct() |>
   dplyr::group_by(subject) |>
   dplyr::mutate(speed = speed_pupil(baseline, trial),
-                MAD = calc_mad(speed, n=16))
+                MAD = calc_mad(speed, n=16)) |>
+  dplyr::ungroup()
+
+## Odd Pupil Slope Detection
+
+slope_df <- mad_removal |>
+  ### Limiting range to first 500 ms after stimulus onset
+  dplyr::filter(time >= 0 & time <= 500) |>
+  group_by(subject, trial) |>
+  ### Calculating change in pupil dilation (i.e. slope) in the first 500 ms
+  dplyr::mutate(pupil_slope = (last(baselinecorrectedp) - first(baselinecorrectedp))/ (last(time) - first(time))) |>
+  dplyr::ungroup()
+
+
+### Visually determining a cutoff point for steep downward slope by plotting the histogram
+### Based on this visualization, pupil slope change greater than -.55 will be the cutoff
+slope_df |>
+  dplyr::select(subject, trial, speaker, pupil_slope) |>
+  distinct() |>
+  ggplot() +
+  aes(x = pupil_slope,
+      fill = speaker,
+      color = speaker) +
+  geom_histogram() +
+  geom_vline(xintercept = -.55)
+
+### Calculating mean and sd of pupil slopes (M = .0194, sd = .271)
+slope_df |>
+  distinct() |>
+  dplyr::summarize(mean = mean(pupil_slope,),
+                   sd = sd(pupil_slope))
+
+### Flagging Trials with Steep Negative Pupil Dilation Slopes
+slope_df <- slope_df |>
+  dplyr::mutate(steep_slope = ifelse(pupil_slope <= -.55, TRUE, FALSE))
+
+## Creating Flag Variables and merging with original df
+
+### Baseline Deviation
+
+baseline_dev <- baseline_dev |>
+  dplyr::select(subject, base_min, base_max) 
+
+mad_removal <- mad_removal |>
[…]

### 2024-05-13 Manuscript Analysis/pupil_data_prep.R
filtered out outliers and exported dfs

@@ -349,13 +349,39 @@ slope_df <- slope_df |>
   dplyr::select(subject, trial, steep_slope) |>
   dplyr::distinct()
 
+mad_removal <- mad_removal |>
+  dplyr::left_join(slope_df, by = c("subject", "trial"))
+
+### Removing Extra Variables
+
+mad_removal <- mad_removal |>
+  dplyr::select(!c(base_min, base_max, peak_min, peak_max))
+
+### Identifying outlier trials that will be removed (21 trials)
+removed_df <- mad_removal |>
+  group_by(subject, trial) |>
+  dplyr::filter(rowSums(across(base_dev:steep_slope)) >= 2)
+
+removed_df <- removed_df |>
+  dplyr::select(subject, trial, speaker) |>
+  dplyr::distinct()
+
+### Filtering out those responses 
+
+filtered_df <- mad_removal |>
+  group_by(subject, trial) |>
+  dplyr::filter(rowSums(across(base_dev:steep_slope)) < 2)
+
+# Removing unneeded objects from the environment
+rm(baseline_dev, baseline_flags, mad_removal, peak_pupil_dev, removed_df, slope_df)
+
 # Downsampling
 
 bin.length <- 20
 
-data.binned <- mad_removal |>
+data.binned <- filtered_df |>
   mutate(timebins = round(time/bin.length)*bin.length) |>
-  dplyr::group_by(subject, trial, speaker, timebins, effort_rating,
+  dplyr::group_by(subject, trial, speaker, timebins,
                   code, targetphrase, counterbalance) |>
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) |>
   dplyr::ungroup()
@@ -380,7 +406,7 @@ bin.length <- 26.23
 ALS_trials <- data.binned |>
   dplyr::filter(speaker == "ALS") |>
   dplyr::mutate(time_n = round(timebins/bin.length)*bin.length) |>
-  dplyr::group_by(subject, trial, speaker, time_n, effort_rating, code, targetphrase, counterbalance) |>
+  dplyr::group_by(subject, trial, speaker, time_n, code, targetphrase, counterbalance) |>
   dplyr::summarize(normed_pupil = mean(pupil.binned)) |>
   dplyr::ungroup()
 

### 2024-05-13 Manuscript Analysis/transcription_prep.R
updated repetition acc script

@@ -81,10 +81,10 @@ reliability <- transcriptions |>
 ## Set working directory to location of cleaned pupil data
 setwd("~/Documents/Listening-Effort-in-Dysarthria/Manuscript Analysis/Cleaned Data")
 
-pupil <- rio::import("cleaned_pupil_data.csv")
+ple <- rio::import("cleaned_ple_data.csv")
 
 ## Getting phrase targets
-targets <- pupil |>
+targets <- ple |>
   dplyr::rename(id = subject,
                 target = targetphrase) |>
   dplyr::select(id, trial, target) |>

### 2024-05-13 Manuscript Analysis/main_analysis.Rmd
small changes to main analysis rmd
demo information summary

@@
 author: "Micah E. Hirsch"
-date: "2024-04-15"
-output: html_document
+date: "2024-05-13"
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
+
 ---
@@
 
-The purpose of this document is to analyze and report findings related to M. Hirsch's dissertation project.
+The purpose of this document is to analyze and report findings related to M. Hirsch's dissertation project. The data preparation and analysis was conducted using R Version 4.3.3.
 
-```{r}
+```{r, echo = F, warning = F, messsage = F}
+
+# Load in Needed Packages
 
@@
 library(irr)
+library(ggpubr) # installl.packages("ggpubr")
+library(gt) # install.packages("gt")
+library(gtsummary) # install.packages("gtsummary")
+library(glmmTMB) # install.packages("glmmTMB")
+library(performance) # install.packages("performance")
+library(sjPlot) # install.packages("sjPlot")
 
-phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv")
+```
+
+# Listener Demographics
+
+The table below shows the demographic information for the listeners in this study.
+A total of 39 listeners were recruited to participate in the study.
+However, 1 participant withdrew from the study (their data was already destroyed before exporting the listener demographic dataset) and 4 participants were dismissed due to difficulty during the calibration phase of the study.
+Therefore, there ended up being 34 total listeners in this study.
+The demographic table shows the demographic information for the 34 participants included in the study (complete) and the 4 who were not able to be tracked by the eye-tracking camera (incomplete).
+
+All listeners reported being a fluent speaker of English.
+However, two participants noted that their native language was different than American English.
+Their reported native languages were Turkish and Spanish.
+
+```{r, warning = F, message = F}
+
+demo <- rio::import("Cleaned Data/cleaned_listener_demo.csv") |>
+  dplyr::mutate(gender = factor(gender, levels = c("Man", "Woman", "Nonbinary", "Questioning", "Prefer not to answer")),
+                ethnicity = factor(ethnicity, levels = c("Hispanic/Latino(a/e)", "Not Hispanic/Latino(a/e)", "Prefer not to answer")),
+                race = factor(race, levels = c("white/Caucasian", "Black/African American", "Asian/Asian American",
+                                               "Native Hawaiian or Other Pacific Islander", "Native American or Alaska Native",
+                                               "Biracial or Multiracial", "Prefer not to answer", "Race not listed")),
+                native_lang = factor(native_lang, c("American English", "Not American English")))
+
+demo_table <- demo |>
+  dplyr::select(age, gender, ethnicity, race, native_lang, pupil_complete) |>
+  tbl_strata(
+    strata = pupil_complete,
+    ~.x |>
+      tbl_summary(type = list(age ~ "continuous",
+                          gender ~ "categorical",
+                          ethnicity ~ "categorical",
+                          race ~ "categorical",
+                          native_lang ~ "categorical"),
+                  statistic = list(all_continuous() ~ "{mean} ({sd}, {min}-{max})",
+                               all_categorical() ~ "{n} ({p}%)"),
+                  digits = list(everything() ~ c(2)),
+                  label = list(age ~ "Age",
+                           gender ~ "Gender",
+                           ethnicity ~ "Ethnicity",
+                           race ~ "Race",
+                           native_lang ~ "Native Language"))
+    ) |>
+  as_gt()
+
+demo_table
+
+demo_table |>
+  gt::gtsave("Tables/demographic_table.html")
 
@@
 
-pupil_data <- rio::import("Cleaned Data/cleaned_pupil_data.csv")
-
-pupil <- pupil_data |>
-  left_join(phrase_acc, by = c("subject", "trial", "targetphrase")) |>
-  dplyr::select(!c(rep_acc_rel, correct_words_rel, rel_response)) |>
-  dplyr::mutate(speaker = ifelse(speaker == "Control", 0, 1),
-                rep_acc = ifelse(rep_acc == "accurate", 0, 1),
-                speaker = factor(speaker, levels = c(0, 1), labels = c("Control", "ALS")),
-                rep_acc = factor(rep_acc, levels = c(0, 1), labels = c("accurate", "inaccurate")),
-                code = as.factor(code))
-
-pupil1 <- pupil |>
-  dplyr::filter(rep_acc != "")
-
 ```

### 2024-05-13 Manuscript Analysis/main_analysis.Rmd
transcription reliability

@@
 
-# Transcription Reliability
+```{r, echo = F, warning = F, message = F}
 
-```{r}
+# Removing unneeded items from the environment
+rm(demo, demo_table)
+
+```
+
+# Descriptives
+
+## Transcription Reliability
+
+Phrase repetition accuracy was determined by transcriptions of the listener's spoken response.
+These reponses were transcribed by 2 research assistants.
+To determine interrater reliability, each RA transcribed at least 20% of the other RA's listener reponses.
+Fleiss' kappa was then used to calculated interrater reliability for the resulting phrase repetition accuracy determination (accurate or inaccurate).
+Based on the result below, our interrater reliability for phrase repetition accuracy was strong.
+
+```{r, warning = F, message = F}
+
+phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv")
+
+phrase_acc <- phrase_acc |>
+  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
+                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
+                rep_acc_rel = ifelse(rel_response == "missing data", NA, 
+                                 ifelse(target_number == correct_words_rel, "accurate", "inaccurate")))
 
 reliability <- phrase_acc |>
-  dplyr::filter(subject != "LE38" & subject != "LE33") |>
-  dplyr::filter(!is.na(correct_words_rel)) |>
+  dplyr::filter(!is.na(rep_acc_rel)) |>
   dplyr::select(rep_acc, rep_acc_rel)
@@
 
+## Speaker Characteristics
+
 # Pupil Dilation

### 2024-05-13 Manuscript Analysis/transcription_prep.R

@@ -161,5 +161,7 @@ phrase_acc <- phrase_acc_initial |>
 
 # Export csv file of cleaned data
 
+
+
 rio::export(phrase_acc, "repetition_accuracy.csv")
 

### 2024-05-13 Manuscript Analysis/main_analysis.Rmd
perceived listening effort ratings

@@
 
+phrase_acc_df <- phrase_acc |>
+  dplyr::select(subject, trial, target_number, correct_words_initial, rep_acc)
+
 ```
@@
 ## Speaker Characteristics
+
+The speaker characteristics are listed below.
+Note: 41 trials were removed because they had missing listener responses (i.e.
+audio file cut out, etc.).
+Therefore, the summary statistics are based on 2679 trials total.
+
+```{r}
+
+ple_data <- rio::import("Cleaned Data/cleaned_ple_data.csv") |>
+  dplyr::left_join(phrase_acc_df, by = c("subject", "trial")) |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                trial_c = trial - 6,
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")))
+
+speaker_table <- ple_data |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::select(subject, trial, speaker, correct_words_initial, target_number, rep_acc) |>
+  dplyr::mutate(intel = (correct_words_initial/target_number)*100) |>
+  dplyr::select(-c(subject, trial, correct_words_initial, target_number)) |>
+  tbl_summary(by = speaker,
+              type = list(intel ~ "continuous",
+                          rep_acc ~ "categorical"),
+              statistic = list(all_continuous() ~ "{mean} ({sd})",
+                               all_categorical() ~ "{n} ({p}%)"),
+              digits = list(everything() ~ c(2)),
+              label = list(intel ~ "Intelligibility",
+                           rep_acc ~ "Repetition Accuracy")) |>
+  as_gt()
+
+speaker_table
+
+speaker_table |>
+  gt::gtsave("Tables/speaker_table.html")
+
+```
+
+# Perceived Listening Effort Ratings
+
+## Descriptives
+
+```{r, warning = F, message = F}
+
+ple_data |>
+  dplyr::select(speaker, effort_rating, rep_acc) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  tbl_strata(
+    strata = rep_acc,
+    ~.x |>
+      tbl_summary(
+        by = speaker,
+        type = list(effort_rating ~ "continuous"),
+        statistic = list(all_continuous() ~ c("{mean} ({sd})")),
+        missing = "no",
+        digits = all_continuous() ~ 2,
+        label = list(effort_rating ~ "Perceived Listening Effort")
+      ))
+
+```
+
+## PLE Models
+
+Linear mixed effects (LME) models were used to test the influence of speaker (Control vs ALS) and phrase repetition accuracy (accurate vs inaccurate) on perceived listening effort ratings.
+We used a model-building approach for this analysis.
+The steps are detailed below.
+
+### Fully Unconditional Model
+
+```{r}
+
+ple_data <- ple_data |>
+  dplyr::filter(!is.na(rep_acc))
+
+m0_ple <- glmmTMB(effort_rating ~ 1 + (1|subject) + (1|code), data = ple_data)
+sjPlot::tab_model(m0_ple, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+performance::icc(m0_ple)
+
+```
+### PLE Model 1
+
+In this step, we are adding trial order (trial_c) as a random slop for both the listener (subject) and phrase (code) random intercepts.
+The random intercept of trial order significantly improved model fit.
+
+```{r}
+
+m1_ple <- glmmTMB(effort_rating ~ 1 + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(m1_ple, pred.labels = "Intercept",
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m0_ple, m1_ple)
+
+```
+### PLE Model 2
+
+Trial order is the first random effect added to the model.
+Trial order is added in order to control for any possible order effects in the data.
+As noted in the results below, adding trial order did not significantly improve model fit to the data.
+However, I'll leave this fixed effect in the model to act as a covariate.
+
+```{r}
+
+m2_ple <- glmmTMB(effort_rating ~ trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(m2_ple, 
+                  pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m1_ple, m2_ple)
+
+```
+
+### PLE Model 3
+
+Adding the main fixed effect of speaker significantly improved the model fit.
+
+```{r}
+
+m3_ple <- glmmTMB(effort_rating ~ trial_c + speaker + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(m3_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m2_ple, m3_ple)
+
+```
+
+### PLE Model 4
+
+The main fixed effect of repetition accuracy significantly improved model fit.
+
+```{r}
+
+m4_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(m4_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m3_ple, m4_ple)
+
+```
+### PLE Model 5
+
+The interaction term between speaker and repetition accuracy did not significantly improve model fit.
+This indicates that the magnitude between accurate and inaccurate ratings was similar for both speakers.
+
+```{r}
+
+m5_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(m5_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m4_ple, m5_ple)
+
+```
+
+## Plot
+
+```{r}
+
+model_ple <- m5_ple |>
+  sjPlot::plot_model(type = "int")
+
+plot_data_ple <- model_ple[["data"]] |>
+  as.data.frame()
+
+```
 

### 2024-05-14 Manuscript Analysis/main_analysis.Rmd
perceived listening effort analysis

@@
 library(sjPlot) # install.packages("sjPlot")
+library(viridis) #install.packages("viridis")
 
@@
   dplyr::select(speaker, effort_rating, rep_acc) |>
+  # Filtering out trials with cut-off listener responses
   dplyr::filter(!is.na(rep_acc)) |>
+  # filtering out trials with missing effort ratings (5 trials missing, 4 additionally removed)
+  dplyr::filter(effort_rating > 0) |>
   tbl_strata(
@@
 
-```{r}
+```{r, warning = F, message = F}
 
 ple_data <- ple_data |>
-  dplyr::filter(!is.na(rep_acc))
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::filter(effort_rating > 0)
 
@@
                   dv.labels = "Perceived Listening Effort")
-
 performance::test_performance(m0_ple, m1_ple)
@@
 
+### Final PLE Model with Unstandardized Estimates
+
+```{r, warning = F, message = F}
+
+final_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(final_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/ple_mod_unstd.html")
+
+```
+
+### Final PLE Model with Standardized Estimates
+
+```{r, warning = F, message = F}
+
+final_ple_std <- glmmTMB(scale(effort_rating) ~ scale(trial_c) + speaker + rep_acc + 
+                    (trial_c|subject) + (trial_c|code), data = ple_data)
+sjPlot::tab_model(final_ple_std,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/ple_mod_std.html")
+
+```
+
+```{r, eval = F}
+
+sjPlot::tab_model(m1_ple, m2_ple, m3_ple, m4_ple, m5_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4", "Model 5"),
+                  file = "Tables/ple_models.html")
+
+```
+
 ## Plot
 
-```{r}
+```{r, warning = F, message = F}
 
-model_ple <- m5_ple |>
-  sjPlot::plot_model(type = "int")
+my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")
 
-plot_data_ple <- model_ple[["data"]] |>
-  as.data.frame()
+ple_data |>
+  dplyr::group_by(speaker, rep_acc) |>
+  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
+                   sd = sd(effort_rating, na.rm = T),
+                   se = sd/sqrt(n())) |>
+  ggplot() +
+   aes(x = rep_acc,
+       y = per_effort,
+       group = speaker,
+       color = speaker,
+       fill = speaker) +
+  geom_bar(stat = "identity", alpha = 0.6, position = position_dodge()) +
+  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
+  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
+  theme_bw() +
+  theme(aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/ple_plot.png", plot = last_plot())
 

### 2024-05-14 Manuscript Analysis/main_analysis.Rmd
started pupil dilation analysis - wip

@@
 
-```{r, echo = F, warning = F, messsage = F}
+```{r, echo = F, warning = F, message = F}
 
@@
 
+The summary statistics for perceived listening effort (PLE) ratings are shown in the table below.
+There were 5 trials with missing PLE ratings, so those were removed from the analysis.
+There is a total of 2675 trials represented in the data below.
+
 ```{r, warning = F, message = F}
@@
 
+Therefore, the best-fitting model for the data is model 4 with just the main effects of speaker and repetition accuracy.
+The results indicate that, on average, the ALS speaker was rated as more effortful by the listeners compared to the control speaker for accurately recognized phrases, controlling for trial order.
+Inaccurate repetitions, on average, were rated higher than accurately recognized phrases for the control speaker by the listeners.
+Since the interaction between speaker and repetition accuracy was not significant, this indicates the difference in PLE scores between accurately and inaccurately recognized phrases was similar between the two speakers.
+
 ```{r, warning = F, message = F}
@@
 ### Final PLE Model with Standardized Estimates
+
+This model shows the standardized estimates for the final PLE model.
 
@@
 
+```{r, echo = F}
+
+rm(final_ple, final_ple_std, m0_ple, m1_ple, m2_ple, m3_ple, m4_ple, m5_ple, reliability, speaker_table)
+
+```
+
 # Pupil Dilation
+
+## Total Number of Trials
+
+A total of 2654 trials are included in the analysis
 
@@
 
+pupil_df <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv") |>
+  # Merging phrase accuracy information with df
+  dplyr::left_join(phrase_acc_df, by = c("subject", "trial")) |>
+  # Filter out trials with missing phrase repetition data |>
+  dplyr::filter(!is.na(rep_acc))
+
+pupil_df |>
+  dplyr::select(subject, trial) |>
+  dplyr::distinct() |>
+  dplyr::summarize(n = n())
+
 ```
+
+## Pupil Dilation Models
+
+### Initial Model
+
+```{r}
+
+# Narrowing to the analysis region of interest
+
+pupil_roi <- pupil_df |>
+  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
+  dplyr::filter(time_norm >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  # filtering out timebins 1000 ms after phrase offset
+  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
+  ungroup() |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
+                phrase = as.factor(code),
+                listener = as.factor(subject))
+
+m0_pupil <- bam(normed_pupil ~ speaker +
+                  s(time_norm, by = speaker, k = 20) +
+                  s(time_norm, by = rep_acc, k = 20) +
+                  # Random Smooths for listener and phrase
+                  s(time_norm, listener, bs = 'fs', m =1) +
+                  s(time_norm, phrase, bs = 'fs', m = 1),
+                discrete = T,
+                data = pupil_roi)
+
+plot_smooth(m0_pupil,
+            view = "time_norm",
+            plot_all = c("speaker", "rep_acc"),
+            rug = F, se = 1, rm.ranef = T)
+
+```
+#### Model Check
+
+```{r}
+
+gam.check(m0_pupil)
+
+acf_resid(m0_pupil)
+
+```

### 2024-05-15 Manuscript Analysis/main_analysis.Rmd
draft of pupil dilation analysis

@@
 
+Initial model with smooths for speaker and phrase.
+
 ```{r}
@@
                 phrase = as.factor(code),
-                listener = as.factor(subject))
+                listener = as.factor(subject),
+                condition = case_when(speaker == "Control" & rep_acc == "accurate" ~ "Control Accurate",
+                                      speaker == "Control" & rep_acc == "inaccurate" ~ "Control Inaccurate",
+                                      speaker == "ALS" & rep_acc == "accurate" ~ "ALS Accurate",
+                                      TRUE ~ "ALS Inaccurate"),
+                condition = factor(condition, 
+                                   levels = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate")))
 
-m0_pupil <- bam(normed_pupil ~ speaker +
-                  s(time_norm, by = speaker, k = 20) +
-                  s(time_norm, by = rep_acc, k = 20) +
+m0_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 20) +
                   # Random Smooths for listener and phrase
@@
 
+summary(m0_pupil)
+
 plot_smooth(m0_pupil,
             view = "time_norm",
-            plot_all = c("speaker", "rep_acc"),
+            plot_all = "condition",
             rug = F, se = 1, rm.ranef = T)
@@
 ```
+
 #### Model Check
@@
 ```
+
+### Model 1
+
+Adding in AR(1) structure to address autocorrelation concerns. Also increased knots based on output from the initial model, and took away the random smooth for phrase due to model convergence issues.
+
+```{r}
+
+# Marking start of trials
+
+pupil_roi <- pupil_roi |>
+  dplyr::group_by(listener, speaker) |>
+  dplyr::mutate(num_points = n()) |>
+  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1))))
+
+m1_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener and phrase
+                  s(time_norm, listener, bs = 'fs', m =1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                data = pupil_roi)
+
+summary(m1_pupil)
+
+plot_smooth(m1_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+#### Model Check
+
+Autocorrelation improved. This model's residuals also look better, but there are still some heavy residual tails.
+
+```{r}
+
+gam.check(m1_pupil)
+
+acf_resid(m1_pupil)
+
+```
+
+### Model 2
+
+Adding scaled t distribution to fix model fit.
+
+```{r}
+
+m2_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener and phrase
+                  s(time_norm, listener, bs = 'fs', m =1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                family = "scat",
+                data = pupil_roi)
+
+summary(m2_pupil)
+
+m2_prediction <- plot_smooth(m2_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+#### Model Check
+
+Model residuals look better, but not great
+
+```{r}
+
+gam.check(m2_pupil)
+
+acf_resid(m2_pupil)
+
+```
+
+#### Difference Plots
+
+```{r}
+
+plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Accurate", "Control Accurate")),
+          rm.ranef = T)
+
+plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("Control Inaccurate", "Control Accurate")),
+          rm.ranef = T)
+
+plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")),
+          rm.ranef = T)
+
+```
+## Plot
+
+Draft pupil dilation plot
+
+I am trying to find a way to predict model-predicted values against the actual data.
+geom_smooth doesn't consider the random effects, so I am thinking I need to extract model data from the final model to show the goodness of fit
+
+It might be good to do this for every iteration of the model to show the process too...
+
+```{r}
+
+ind_pupil_df <- pupil_df |>
+  dplyr::group_by(speaker, subject, rep_acc, time_norm) |>
+  summarize(pupil = mean(normed_pupil, na.rm = T),
+            se = sd(normed_pupil, na.rm = T)/sqrt(n()))
+
+sum_pupil_df <- ind_pupil_df |>
+  dplyr::group_by(speaker, rep_acc, time_norm) |>
+  summarize(m_pupil = mean(pupil, na.rm = T),
+            se_pupil = sd(pupil, na.rm = T)/sqrt(n()))
+
+# I'm trying to find a way to plot predicted values against actual data.
+predicted_values <- m2_prediction[[1]][c("condition", "time_norm", "ul", "ll")]
+
+sum_pupil_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = m_pupil,
+      color = speaker,
+      fill = speaker) +
+  geom_smooth() +
+  geom_line() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  facet_wrap("rep_acc") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/pupil_plot.png", plot = last_plot())
+
+```

### 2024-05-15 Manuscript Analysis/main_analysis.Rmd
draft pupil plot

@@
   geom_smooth() +
-  geom_line() +
+  geom_line(size = 1) +
   geom_vline(xintercept = 0, size = 0.35) +

### 2024-05-16 Manuscript Analysis/adaptation_analysis.Rmd
ple adaptation analysis

---
title: 'Exploratory Analyses: Adaptation'
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains the code and results for an exploratory analysis done with the data collected from this study. Specifically, this document reports findings about change in listening effort (measured by perceived listening effort ratings and pupil dilation) across the study trials.

```{r}

library(rio)
library(tidyverse)
library(mgcv)
library(itsadug)
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(viridis) #install.packages("viridis")

```

# Perceived Listening Effort

## PLE Models

### Model 5

```{r}

# Loading in Phrase Repetition Accuracy Data
phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::select(subject, trial, targetphrase, initial_response, target_number, correct_words_initial, rep_acc) |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate"))) |>
  dplyr::select(!c(targetphrase, initial_response))

# Loading in PLE data and merging transcription accuracy with it
ple_data <- rio::import("Cleaned Data/cleaned_ple_data.csv") |>
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                trial_c = trial - 6,
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
  dplyr::filter(!is.na(rep_acc)) |>
  dplyr::filter(effort_rating > 0)

m5_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m5_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

```

### Model 6

```{r}

m6_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m6_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m6_ple)

```

### Model 7

```{r}

m7_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + rep_acc*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m7_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m7_ple)

```

### Model 8

```{r}

m8_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
                    speaker*trial_c + rep_acc*trial_c +  speaker*rep_acc*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
sjPlot::tab_model(m8_ple,
                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
                  dv.labels = "Perceived Listening Effort")

performance::test_performance(m5_ple, m8_ple)

```
## Plot

```{r}

my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")

ple_data |>
  dplyr::group_by(speaker, rep_acc, trial_c) |>
  dplyr::summarize(per_effort = mean(effort_rating, na.rm = T), 
                   sd = sd(effort_rating, na.rm = T),
                   se = sd/sqrt(n())) |>
  ggplot() +
   aes(x = trial_c,
       y = per_effort,
       group = speaker,
       color = speaker) +
  geom_line() +
  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0) +
  labs(x = "Trial", y = "Perceived Listening Effort Rating") +
  facet_wrap("rep_acc", ncol = 1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
  scale_fill_manual(values = c(my_pal[1], my_pal[5]))

ggsave("Figures/ple_trial_plot.png", plot = last_plot())

```

# Pupil Dilation

### 2024-05-16 Manuscript Analysis/adaptation_analysis.Rmd
first attempt at pupil adaptation analysis

@@
 # Pupil Dilation
+
+```{r}
+
+pupil_df <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv") |>
+  # Merging phrase accuracy information with df
+  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
+  # Filter out trials with missing phrase repetition data |>
+  dplyr::filter(!is.na(rep_acc))
+
+pupil_roi <- pupil_df |>
+  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
+  dplyr::filter(time_norm >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  # filtering out timebins 1000 ms after phrase offset
+  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
+  ungroup() |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
+                phrase = as.factor(code),
+                listener = as.factor(subject),
+                condition = case_when(speaker == "Control" & rep_acc == "accurate" ~ "Control Accurate",
+                                      speaker == "Control" & rep_acc == "inaccurate" ~ "Control Inaccurate",
+                                      speaker == "ALS" & rep_acc == "accurate" ~ "ALS Accurate",
+                                      TRUE ~ "ALS Inaccurate"),
+                condition = factor(condition, 
+                                   levels = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate")),
+                trial_c = trial - 6) |>
+  dplyr::group_by(listener, speaker) |>
+  dplyr::mutate(num_points = n()) |>
+  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
+  ungroup()
+
+adapt_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 20) +
+                  te(time_norm, trial_c, by = condition) +
+                  # Random Smooths for listener, phrase, trial
+                  s(time_norm, listener, bs = 'fs', m =1) +
+                  s(time_norm, phrase, bs = 'fs', m = 1),
+                discrete = T,
+                family = "scat",
+                AR.start = start_event,
+                rho = .92,
+                data = pupil_roi)
+
+summary(adapt_pupil)
+
+plot_smooth(adapt_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+```{r}
+
+gam.check(adapt_pupil)
+
+acf_resid(adapt_pupil)
+
+```

### 2024-05-16 Manuscript Analysis/main_analysis.Rmd
first attempt at pupil adaptation analysis

@@
   dplyr::mutate(num_points = n()) |>
-  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1))))
+  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
+  ungroup()
 

### 2024-05-16 Manuscript Analysis/adaptation_analysis.Rmd
Small changes

@@
 
-```{r}
+```{r, echo = F, warning = F, message = F}
 
-library(rio)
-library(tidyverse)
-library(mgcv)
-library(itsadug)
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+library(mgcv) # install.packages("mgcv")
+library(itsadug) # install.packages("itsadug")
 library(ggpubr) # installl.packages("ggpubr")
@@
 library(sjPlot) # install.packages("sjPlot")
-library(viridis) #install.packages("viridis")
 
@@
 ```
+
+```{r, echo = F, message = F, warning = F}
+
+sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker * Repetition Accuracy"),
+                  dv.labels = c("Model 5", "Model 6", "Model 7", "Model 8"),
+                 file = "Tables/ple_adaptation_models.html")
+
+```
+
 ## Plot
 
-```{r}
+```{r, warning = F, message = F}
 

### 2024-05-17 Manuscript Analysis/adaptation_analysis.Rmd
added rep acc analysis

@@
 
-# Perceived Listening Effort
-
-## PLE Models
-
-### Model 5
+# Proportion of Correct Responses per Trial
 
@@
   dplyr::filter(effort_rating > 0)
+
+my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")
+
+rep_acc_df <- ple_data |>
+  dplyr::select(subject, trial_c, speaker, rep_acc)
+
+rep_acc_df |>
+  dplyr::mutate(trial_c = trial_c + 1) |>
+  dplyr::group_by(speaker, trial_c, rep_acc) |>
+  dplyr::summarize(n = n()) |>
+  dplyr::ungroup() |>
+  tidyr::pivot_wider(names_from = "rep_acc",
+                     values_from = "n") |>
+  dplyr::mutate(accurate = ifelse(is.na(accurate), 0, accurate),
+                inaccurate = ifelse(is.na(inaccurate), 0, inaccurate),
+                prop_accurate = accurate/(accurate + inaccurate)) |>
+  ggplot() +
+  aes(x = trial_c,
+      y = prop_accurate,
+      group = speaker,
+      color = speaker) +
+  geom_line() +
+  labs(y = "Proportion of Accurate Responses",
+       x = "Trial Number") +
+  facet_wrap("speaker", ncol = 1) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/acc_response_plot.png", plot = last_plot())
+
+```
+
+```{r}
+
+rep_acc_df <- rep_acc_df |>
+  dplyr::mutate(accurate = ifelse(rep_acc == "accurate", 1, 0))
+
+rep_acc_mod <- glmmTMB(accurate ~ trial_c*speaker + (trial_c|subject), family = "binomial", data = rep_acc_df)
+
+summary(rep_acc_mod)
+
+sjPlot::tab_model(rep_acc_mod, transform = NULL, show.est = TRUE, show.se = TRUE,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                  "Trial Order * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/rep_acc_model.html")
+
+```
+
+# Perceived Listening Effort
+
+## PLE Models
+
+### Model 5
+
+```{r}
 
@@
 
-my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")
-
 ple_data |>
+  dplyr::mutate(trial_c = trial_c + 1) |>
   dplyr::group_by(speaker, rep_acc, trial_c) |>
@@
 ggsave("Figures/ple_trial_plot.png", plot = last_plot())
+
+```
+
+```{r}
+
+rm(m5_ple, m6_ple, m7_ple, m8_ple, rep_acc_df, rep_acc_mod)
 

### 2024-05-20 Manuscript Analysis/adaptation_analysis.Rmd
adjusted figures

--- 2024-05-16 Manuscript Analysis/adaptation_analysis.Rmd
@@
   dplyr::filter(effort_rating > 0)
+
+my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")
+
+rep_acc_df <- ple_data |>
+  dplyr::select(subject, trial_c, speaker, rep_acc)
+
+rep_acc_df |>
+  dplyr::mutate(trial_c = trial_c + 1) |>
+  dplyr::group_by(speaker, trial_c, rep_acc) |>
+  dplyr::summarize(n = n()) |>
+  dplyr::ungroup() |>
+  tidyr::pivot_wider(names_from = "rep_acc",
+                     values_from = "n") |>
+  dplyr::mutate(accurate = ifelse(is.na(accurate), 0, accurate),
+                inaccurate = ifelse(is.na(inaccurate), 0, inaccurate),
+                prop_accurate = accurate/(accurate + inaccurate)) |>
+  ggplot() +
+  aes(x = trial_c,
+      y = prop_accurate,
+      group = speaker,
+      color = speaker,
+      fill = speaker) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  labs(y = "Proportion of Accurate Responses",
+       x = "Trial Number") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/acc_response_plot.png", plot = last_plot())
+
+```
+
+```{r}
+
+rep_acc_df <- rep_acc_df |>
+  dplyr::mutate(accurate = ifelse(rep_acc == "accurate", 1, 0))
+
+rep_acc_mod <- glmmTMB(accurate ~ trial_c*speaker + (trial_c|subject), family = "binomial", data = rep_acc_df)
+
+summary(rep_acc_mod)
+
+sjPlot::tab_model(rep_acc_mod, transform = NULL, show.est = TRUE, show.se = TRUE,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", 
+                                  "Trial Order * Speaker [ALS]"),
+                  dv.labels = "Perceived Listening Effort",
+                  file = "Tables/rep_acc_model.html")
+
+```
+
+# Perceived Listening Effort
+
+## PLE Models
+
+### Model 5
+
+```{r}
 
@@
        group = speaker,
-       color = speaker) +
-  geom_line() +
-  geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0) +
+       color = speaker,
+       fill = speaker) +
+  geom_point() +
+  geom_smooth(method = "lm") +
   labs(x = "Trial", y = "Perceived Listening Effort Rating") +
-  facet_wrap("rep_acc", ncol = 1) +
+  facet_wrap("rep_acc") +
   theme_bw() +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
   scale_color_manual(values = c(my_pal[1], my_pal[5])) +
@@
   dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
-  # Filter out trials with missing phrase repetition data |>
+  # Filter out trials with missing phrase repetition data 
   dplyr::filter(!is.na(rep_acc))

### 2024-05-21 Manuscript Analysis/main_analysis.Rmd
updated main pupil analyses - wip

@@
 
-# Narrowing to the analysis region of interest
-
-pupil_roi <- pupil_df |>
-  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
-  dplyr::filter(time_norm >= 500) |>
-  dplyr::group_by(subject, trial) |>
-  # filtering out timebins 1000 ms after phrase offset
-  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
-  ungroup() |>
+pupil_df <- pupil_df |>
   dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
@@
                 discrete = T,
-                data = pupil_roi)
+                data = pupil_df)
 
@@
 
-Adding in AR(1) structure to address autocorrelation concerns. Also increased knots based on output from the initial model, and took away the random smooth for phrase due to model convergence issues.
+Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots too.
 
@@
 
-pupil_roi <- pupil_roi |>
+pupil_df <- pupil_df |>
   dplyr::group_by(listener, speaker) |>
@@
                   # Random Smooths for listener and phrase
-                  s(time_norm, listener, bs = 'fs', m =1),
+                  s(time_norm, listener, bs = 'fs', m =1) +
+                  s(time_norm, phrase, bs = 'fs', m =1),
                 discrete = T,
@@
                 rho = start_value_rho(m0_pupil),
-                data = pupil_roi)
+                data = pupil_df)
 
@@
 
-plot_smooth(m1_pupil,
+m1_prediction <- plot_smooth(m1_pupil,
             view = "time_norm",
@@
 
+#### Plot
+
+```{r}
+
+ind_pupil_df <- pupil_df |>
+  dplyr::group_by(speaker, subject, rep_acc, time_norm) |>
+  summarize(pupil = mean(normed_pupil, na.rm = T),
+            se = sd(normed_pupil, na.rm = T)/sqrt(n()))
+
+sum_pupil_df <- ind_pupil_df |>
+  dplyr::group_by(speaker, rep_acc, time_norm) |>
+  summarize(m_pupil = mean(pupil, na.rm = T),
+            se_pupil = sd(pupil, na.rm = T)/sqrt(n()))
+
+m1_predicted_values <- m1_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
+
+sum_pupil_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = m_pupil,
+      color = speaker,
+      fill = speaker) +
+  geom_line() +
+  geom_line(data = m1_predicted_values, aes(x = time_norm, y = fit, color = speaker, fill = speaker), linetype = 2) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  facet_wrap("rep_acc") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+```
+
 ### Model 2
 
-Adding scaled t distribution to fix model fit.
+Adding scaled t distribution to fix model fit. The model was originally running into an error with the random smooth for phrase. Therefore, it was removed in model 2.
 
@@
                 family = "scat",
-                data = pupil_roi)
+                data = pupil_df)
 
@@
 
-#### Difference Plots
+#### Plot
 
@@
 
-plot_diff(m2_pupil, view = "time_norm",
-          comp = list(condition = c("ALS Accurate", "Control Accurate")),
-          rm.ranef = T)
-
-plot_diff(m2_pupil, view = "time_norm",
-          comp = list(condition = c("Control Inaccurate", "Control Accurate")),
-          rm.ranef = T)
-
-plot_diff(m2_pupil, view = "time_norm",
-          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")),
-          rm.ranef = T)
-
-```
-## Plot
-
-Draft pupil dilation plot
-
-I am trying to find a way to predict model-predicted values against the actual data.
-geom_smooth doesn't consider the random effects, so I am thinking I need to extract model data from the final model to show the goodness of fit
-
-It might be good to do this for every iteration of the model to show the process too...
-
-```{r}
-
-ind_pupil_df <- pupil_df |>
-  dplyr::group_by(speaker, subject, rep_acc, time_norm) |>
-  summarize(pupil = mean(normed_pupil, na.rm = T),
-            se = sd(normed_pupil, na.rm = T)/sqrt(n()))
-
-sum_pupil_df <- ind_pupil_df |>
-  dplyr::group_by(speaker, rep_acc, time_norm) |>
-  summarize(m_pupil = mean(pupil, na.rm = T),
-            se_pupil = sd(pupil, na.rm = T)/sqrt(n()))
-
-# I'm trying to find a way to plot predicted values against actual data.
-predicted_values <- m2_prediction[[1]][c("condition", "time_norm", "ul", "ll")]
+m2_predicted_values <- m2_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
 
@@
       fill = speaker) +
-  geom_smooth() +
-  geom_line(size = 1) +
+  geom_line() +
+  geom_line(data = m2_predicted_values, aes(x = time_norm, y = fit, color = speaker, fill = speaker), linetype = 2) +
   geom_vline(xintercept = 0, size = 0.35) +
@@
 
-ggsave("Figures/pupil_plot.png", plot = last_plot())
+```
+
+### Model 3
+
+Adjusting random smooth to be for each listener by condition
+
+```{r}
+
+m3_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener and phrase
+                  s(time_norm, listener, by = condition, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                family = "scat",
+                data = pupil_df)
+
+summary(m3_pupil)
+
+m3_prediction <- plot_smooth(m3_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
 
 ```
+
+#### Model Check
+
+```{r}
+
+gam.check(m3_pupil)
+
+acf_resid(m3_pupil)
+
+```
+
+#### Plot
+
+```{r}
+
+m3_predicted_values <- m3_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
+
+sum_pupil_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = m_pupil,
+      color = rep_acc,
+      fill = rep_acc) +
+  geom_line() +
+  geom_line(data = m3_predicted_values, aes(x = time_norm, y = fit, color = rep_acc, fill = rep_acc), linetype = 2) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+```
+
+### Model Comparisons
+
+```{r}
+
+compareML(m2_pupil, m3_pupil, signif.stars=T)
+
+```

### 2024-05-22 Manuscript Analysis/main_analysis.Rmd
Completed main pupil analysis

@@
 
-A total of 2654 trials are included in the analysis
+A total of 2654 trials are included in the analysis.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
-Initial model with smooths for speaker and phrase.
+An initial GAMM was run with random smooths added for listener and phrases.
 
-```{r}
+```{r, warning = F, message = F}
 
@@
 
-summary(m0_pupil)
+#summary(m0_pupil)
 
@@
 
-summary(m1_pupil)
+# summary(m1_pupil)
 
 m1_prediction <- plot_smooth(m1_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T, plot = F)
+
+```
+
+#### Model Check
+
+Adding the AR(1) structure improved the autocorrelation from the initial. This model's residuals also look better, but there are still some heavy residual tails.
+
+```{r, warning = F, message = F}
+
+gam.check(m1_pupil)
+
+acf_resid(m1_pupil)
+
+```
+
+### Model 2
+
+In this model, the scaled t distribution was added to the model specifications to fix model fit to the data.
+The model was originally running into an error when the random smooth for phrase was included in the specification.
+Therefore, it was removed in this model.
+The random smooth for listener remained in this model
+
+```{r, warning = F, message = F}
+
+m2_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener
+                  s(time_norm, listener, bs = 'fs', m =1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                family = "scat",
+                data = pupil_df)
+
+# summary(m2_pupil)
+
+m2_prediction <- plot_smooth(m2_pupil,
             view = "time_norm",
@@
 
-Autocorrelation improved. This model's residuals also look better, but there are still some heavy residual tails.
+Model residuals look better in model 2. Adjusted R-squared = 0.173
 
-```{r}
+```{r, warning = F, message = F}
 
-gam.check(m1_pupil)
+gam.check(m2_pupil)
 
-acf_resid(m1_pupil)
+acf_resid(m2_pupil)
 
@@
 
-#### Plot
+## Final Pupil Dilation Model Plot
 
-```{r}
+### Overall Model Plot
 
-ind_pupil_df <- pupil_df |>
-  dplyr::group_by(speaker, subject, rep_acc, time_norm) |>
-  summarize(pupil = mean(normed_pupil, na.rm = T),
-            se = sd(normed_pupil, na.rm = T)/sqrt(n()))
+```{r, warning = F, message = F}
 
-sum_pupil_df <- ind_pupil_df |>
-  dplyr::group_by(speaker, rep_acc, time_norm) |>
-  summarize(m_pupil = mean(pupil, na.rm = T),
-            se_pupil = sd(pupil, na.rm = T)/sqrt(n()))
+m2_predicted_values <- m2_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
 
-m1_predicted_values <- m1_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
-  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
-                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
-
-sum_pupil_df |>
+m2_predicted_values |>
   ggplot() +
   aes(x = time_norm,
-      y = m_pupil,
+      y = fit,
       color = speaker,
       fill = speaker) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
   geom_line() +
-  geom_line(data = m1_predicted_values, aes(x = time_norm, y = fit, color = speaker, fill = speaker), linetype = 2) +
   geom_vline(xintercept = 0, size = 0.35) +
   geom_vline(xintercept = 1910, linetype = 2) +
-  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
   coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
@@
 
-```
-
-### Model 2
-
-Adding scaled t distribution to fix model fit. The model was originally running into an error with the random smooth for phrase. Therefore, it was removed in model 2.
-
-```{r}
-
-m2_pupil <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 25) +
-                  # Random Smooths for listener and phrase
-                  s(time_norm, listener, bs = 'fs', m =1),
-                discrete = T,
-                AR.start = start_event,
-                rho = start_value_rho(m0_pupil),
-                family = "scat",
-                data = pupil_df)
-
-summary(m2_pupil)
-
-m2_prediction <- plot_smooth(m2_pupil,
-            view = "time_norm",
-            plot_all = "condition",
-            rug = F, se = 1, rm.ranef = T)
+ggsave("Figures/pupil_plot.png", plot = last_plot())
 
@@
 
-#### Model Check
+### Curve Differences
 
-Model residuals look better, but not great
+#### Control vs ALS
 
-```{r}
+```{r, warning = F, message = F}
 
-gam.check(m2_pupil)
+add_limits_etc <- function(mod){
+  mod <- mod |>
+    dplyr::mutate(lower = est - CI,
+                  upper = est + CI,
+                  is_signif = ifelse(sign(upper) == sign(lower), "yes", "no"))
 
-acf_resid(m2_pupil)
+  return(mod)
+  }
+
+speaker_diff_df <- plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Accurate", "Control Accurate")), plot = F) |>
+  mutate(speaker_1 = "ALS", speaker_2 = "Control", rep_acc = "Accurate", Type = " ALS vs Control") |>
+  add_limits_etc()
+
+speaker_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(speaker_diff_df)) {
+  speaker_diff_df$signif_chunk[row_index] <- 
+    ifelse(speaker_diff_df$is_signif[row_index] != speaker_diff_df$is_signif[row_index - 1],
+           speaker_diff_df$signif_chunk[row_index - 1] + 1,
+           speaker_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[4])
+
+diff_fill <- c( my_pal[4])
+
+diff_plot_1 <- speaker_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "ALS vs Control for Accurately \nRecognized Phrases") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_1
 
@@
 
-#### Plot
+#### Control Accurate vs Inaccurate
 
-```{r}
+```{r, warning = F, message = F}
 
-m2_predicted_values <- m2_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
-  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
-                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
+control_diff_df <- plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("Control Inaccurate", "Control Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "Control", Type = "Control Inaccurate vs Accurate") |>
+  add_limits_etc()
 
-sum_pupil_df |>
+control_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(control_diff_df)) {
+  control_diff_df$signif_chunk[row_index] <- 
+    ifelse(control_diff_df$is_signif[row_index] != control_diff_df$is_signif[row_index - 1],
+           control_diff_df$signif_chunk[row_index - 1] + 1,
+           control_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[1])
+
+diff_fill <- c( my_pal[1])
+
+diff_plot_2 <- control_diff_df |>
   ggplot() +
   aes(x = time_norm,
-      y = m_pupil,
-      color = speaker,
-      fill = speaker) +
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
   geom_line() +
-  geom_line(data = m2_predicted_values, aes(x = time_norm, y = fit, color = speaker, fill = speaker), linetype = 2) +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
   geom_vline(xintercept = 0, size = 0.35) +
   geom_vline(xintercept = 1910, linetype = 2) +
-  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
-  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
-  facet_wrap("rep_acc") +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases For the Control Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
   theme_bw() +
+  geom_ribbon(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
   theme(aspect.ratio = 1,
-        legend.position = "bottom") +
-  labs(x = "Normalized Time (ms)",
-       y = "Pupil Dilation (Arbitrary Units)") +
-  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_2
 
@@
 
-### Model 3
+#### ALS Accurate vs Inaccurate
 
-Adjusting random smooth to be for each listener by condition
+```{r, warning = F, message = F}
 
-```{r}
+als_diff_df <- plot_diff(m2_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "ALS", Type = "ALS Inaccurate vs Accurate") |>
+  add_limits_etc()
 
-m3_pupil <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 25) +
-                  # Random Smooths for listener and phrase
-                  s(time_norm, listener, by = condition, bs = 'fs', m = 1),
-                discrete = T,
-                AR.start = start_event,
-                rho = start_value_rho(m0_pupil),
-                family = "scat",
-                data = pupil_df)
+als_diff_df$signif_chunk <- 1
 
-summary(m3_pupil)
+for (row_index in 2:nrow(als_diff_df)) {
+  als_diff_df$signif_chunk[row_index] <- 
+    ifelse(als_diff_df$is_signif[row_index] != als_diff_df$is_signif[row_index - 1],
+           als_diff_df$signif_chunk[row_index - 1] + 1,
+           als_diff_df$signif_chunk[row_index - 1])
 
-m3_prediction <- plot_smooth(m3_pupil,
-            view = "time_norm",
-            plot_all = "condition",
-            rug = F, se = 1, rm.ranef = T)
+}
+
+diff_color <- c(my_pal[5])
+
+diff_fill <- c( my_pal[5])
+
+diff_plot_3 <- als_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases for the ALS Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_3
 
@@
 
-#### Model Check
+```{r, eval = F}
 
-```{r}
+library(patchwork)
 
-gam.check(m3_pupil)
+diff_plot_1 + diff_plot_2 + diff_plot_3
 
-acf_resid(m3_pupil)
+ggsave("Figures/difference_plots.png", plot = last_plot(), width = 6, height = 2.5, units = "in", scale = 1.8)
 
 ```
-
-#### Plot
-
-```{r}
-
-m3_predicted_values <- m3_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
-  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
-                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"))
-
-sum_pupil_df |>
-  ggplot() +
-  aes(x = time_norm,
-      y = m_pupil,
-      color = rep_acc,
-      fill = rep_acc) +
-  geom_line() +
-  geom_line(data = m3_predicted_values, aes(x = time_norm, y = fit, color = rep_acc, fill = rep_acc), linetype = 2) +
-  geom_vline(xintercept = 0, size = 0.35) +
-  geom_vline(xintercept = 1910, linetype = 2) +
-  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.2) +
-  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
-  facet_wrap("speaker") +
-  theme_bw() +
-  theme(aspect.ratio = 1,
-        legend.position = "bottom") +
-  labs(x = "Normalized Time (ms)",
-       y = "Pupil Dilation (Arbitrary Units)") +
-  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
-
-```
-
-### Model Comparisons
-
-```{r}
-
-compareML(m2_pupil, m3_pupil, signif.stars=T)
-
-```

### 2024-05-23 Manuscript Analysis/adaptation_analysis.Rmd
small change to adaptation script
pupil analysis

@@
 date: "`r Sys.Date()`"
-output: html_document
+output: 
+  html_document:
+    toc: true
+    toc_float: true
+    code_folding: hide
 ---
@@
 
-```{r, echo = F, message = F, warning = F}
+```{r, eval = F, message = F, warning = F}
 
@@
 
+## Plot
+
 ```{r}
@@
 
-pupil_roi <- pupil_df |>
-  # filtering out timebins below 500 (i.e., keeping timebins above 500 ms)
-  dplyr::filter(time_norm >= 500) |>
-  dplyr::group_by(subject, trial) |>
-  # filtering out timebins 1000 ms after phrase offset
-  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
-  ungroup() |>
+pupil_df <- pupil_df |>
   dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
@@
 
+# Extracting max pupil dilation
+max_pupil <- pupil_df |>  
+  dplyr::filter(time_norm >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  # filtering out timebins 1000 ms after phrase offset
+  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
+  dplyr::ungroup() |>
+  # Getting Peak Pupil Dilation Value
+  dplyr::group_by(subject, trial, speaker, code, rep_acc) |>
+  dplyr::summarize(peak_pupil = max(normed_pupil)) |>
+  dplyr::ungroup() |>
+  dplyr::mutate(trial_c = trial - 6)
+
+max_pupil |>
+  dplyr::mutate(trial_c = trial_c + 1) |>
+  dplyr::group_by(speaker, rep_acc, trial_c) |>
+  dplyr::summarize(peak = mean(peak_pupil, na.rm = T), 
+                   se = sd(peak_pupil, na.rm = T)/sqrt(n())) |>
+  ggplot() +
+  aes(x = trial_c,
+      y = peak,
+      color = speaker,
+      fill = speaker) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  labs(x = "Trial", y = "Peak Pupil Dilation (Arbitrary Units)") +
+  facet_wrap("rep_acc") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+```
+
+## LME Model: Peak Pupil
+
+### Fully Unconditional Model
+
+Tried running with random intercept for phrase too, but ran into singular fit issues. So it was removed. ICC = .331
+
+```{r}
+
+m0_pupil <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = max_pupil)
+summary(m0_pupil)
+performance::icc(m0_pupil)
+
+```
+
+### Model 1
+
+Adding trial random slope
+
+```{r}
+
+m1_pupil <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = max_pupil)
+summary(m1_pupil)
+performance::test_performance(m0_pupil, m1_pupil)
+
+```
+
+### Model 2
+
+```{r}
+
+m2_pupil <- glmmTMB(peak_pupil ~ trial_c + (trial_c|subject), data = max_pupil)
+summary(m2_pupil)
+performance::test_performance(m1_pupil, m2_pupil)
+
+```
+
+### Model 3
+
+```{r}
+
+m3_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + (trial_c|subject), data = max_pupil)
+summary(m3_pupil)
+performance::test_performance(m2_pupil, m3_pupil)
+
+```
+
+### Model 4
+
+```{r}
+
+m4_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = max_pupil)
+summary(m4_pupil)
+performance::test_performance(m3_pupil, m4_pupil)
+
+```
+
+### Model 5
+
+```{r}
+
+m5_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
+summary(m5_pupil)
+performance::test_performance(m4_pupil, m5_pupil)
+
+```
+
+### Model 5b
+
+```{r}
+
+m5b_pupil <- glmmTMB(peak_pupil ~ speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
+summary(m5b_pupil)
+performance::test_performance(m5_pupil, m5b_pupil)
+
+```
+
+### Model 6
+
+```{r}
+
+m6_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+                      speaker*rep_acc + trial_c*speaker +(trial_c|subject), data = max_pupil)
+summary(m6_pupil)
+performance::test_performance(m4_pupil, m6_pupil)
+
+```
+
+### Model 7
+
+```{r}
+
+m7_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + (trial_c|subject), data = max_pupil)
+summary(m7_pupil)
+performance::test_performance(m6_pupil, m7_pupil)
+
+```
+
+### Model 8
+
+```{r}
+
+m8_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + 
+                      speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + 
+                      trial_c*speaker*rep_acc + (trial_c|subject), data = max_pupil)
+summary(m8_pupil)
+performance::test_performance(m7_pupil, m8_pupil)
+
+```
+
+### Final Model with Unstandardized Estimates
+
+```{r}
+
+sjPlot::tab_model(m7_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation",
+                  file = "Tables/peak_pupil_adapt_unstd.html")
+
+```
+
+### Final Model with Standardized Estimates
+
+```{r}
+
+m7_pupil_std <- glmmTMB(scale(peak_pupil) ~ scale(trial_c) + speaker + rep_acc + 
+                      speaker*rep_acc + scale(trial_c)*speaker + scale(trial_c)*rep_acc + (trial_c|subject), data = max_pupil)
+sjPlot::tab_model(m7_pupil_std,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation",
+                  file = "Tables/peak_pupil_adapt_std.html")
+
+```
+
+```{r}
+
+rm(m0_pupil, m1_pupil, m2_pupil, m3_pupil, m4_pupil, m5_pupil, m6_pupil, m7_pupil, m8_pupil, m5b_pupil, max_pupil)
+
+```
+
+## GAMM
+
+```{r, eval = F}
+
 adapt_pupil <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 20) +
-                  te(time_norm, trial_c, by = condition) +
-                  # Random Smooths for listener, phrase, trial
-                  s(time_norm, listener, bs = 'fs', m =1) +
-                  s(time_norm, phrase, bs = 'fs', m = 1),
+                  s(time_norm, by = condition, k = 25) +
+                  te(time_norm, trial_c, k = 25) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m =1),
                 discrete = T,
@@
                 rho = .92,
-                data = pupil_roi)
+                data = pupil_df)
 
@@
 
-```{r}
+```{r, eval = F}
 

### 2024-05-23 Manuscript Analysis/main_analysis.Rmd
pupil analysis

@@
 
-```{r, warning = F, message = F}
+```{r}
 
@@
 
-# summary(m1_pupil)
+summary.gam(m1_pupil)
 
@@
                   s(time_norm, by = condition, k = 25) +
-                  # Random Smooths for listener
-                  s(time_norm, listener, bs = 'fs', m =1),
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
                 discrete = T,
@@
 
-# summary(m2_pupil)
+summary.gam(m2_pupil)
 
@@
 
-## Final Pupil Dilation Model Plot
+### Model 3
+
+```{r}
+
+m3_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener
+                  s(time_norm, listener, by = condition, bs = 'fs', m =1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                family = "scat",
+                data = pupil_df)
+
+summary.gam(m3_pupil)
+
+m3_prediction <- plot_smooth(m3_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+#### Model Check
+
+```{r}
+
+gam.check(m3_pupil)
+
+acf_resid(m3_pupil)
+
+```
+
+## Model 2 Fit
 
@@
 
-ggsave("Figures/pupil_plot.png", plot = last_plot())
+# ggsave("Figures/pupil_plot.png", plot = last_plot())
 
@@
 
+## Model 3 Fit
+
+### Overall Model Plot
+
+```{r, warning = F, message = F}
+
+m3_predicted_values <- m3_prediction[[1]][c("condition", "time_norm", "fit", "ul", "ll")] |>
+  dplyr::mutate(rep_acc = ifelse(grepl("inaccurate", condition, ignore.case = T), "inaccurate", "accurate"),
+                speaker = ifelse(grepl("Control", condition, ignore.case = T), "Control", "ALS"),
+                speaker = factor(speaker, levels = c("Control", "ALS")))
+
+m3_predicted_values |>
+  ggplot() +
+  aes(x = time_norm,
+      y = fit,
+      color = speaker,
+      fill = speaker) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  facet_wrap("rep_acc") +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+
+# ggsave("Figures/pupil_plot.png", plot = last_plot())
+
+```
+
+### Curve Differences
+
+#### Control vs ALS
+
+```{r, warning = F, message = F}
+
+add_limits_etc <- function(mod){
+  mod <- mod |>
+    dplyr::mutate(lower = est - CI,
+                  upper = est + CI,
+                  is_signif = ifelse(sign(upper) == sign(lower), "yes", "no"))
+
+  return(mod)
+  }
+
+speaker_diff_df <- plot_diff(m3_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Accurate", "Control Accurate")), plot = F) |>
+  mutate(speaker_1 = "ALS", speaker_2 = "Control", rep_acc = "Accurate", Type = " ALS vs Control") |>
+  add_limits_etc()
+
+speaker_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(speaker_diff_df)) {
+  speaker_diff_df$signif_chunk[row_index] <- 
+    ifelse(speaker_diff_df$is_signif[row_index] != speaker_diff_df$is_signif[row_index - 1],
+           speaker_diff_df$signif_chunk[row_index - 1] + 1,
+           speaker_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[4])
+
+diff_fill <- c( my_pal[4])
+
+diff_plot_1 <- speaker_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "ALS vs Control for Accurately \nRecognized Phrases") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_1
+
+```
+
+#### Control Accurate vs Inaccurate
+
+```{r, warning = F, message = F}
+
+control_diff_df <- plot_diff(m3_pupil, view = "time_norm",
+          comp = list(condition = c("Control Inaccurate", "Control Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "Control", Type = "Control Inaccurate vs Accurate") |>
+  add_limits_etc()
+
+control_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(control_diff_df)) {
+  control_diff_df$signif_chunk[row_index] <- 
+    ifelse(control_diff_df$is_signif[row_index] != control_diff_df$is_signif[row_index - 1],
+           control_diff_df$signif_chunk[row_index - 1] + 1,
+           control_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[1])
+
+diff_fill <- c( my_pal[1])
+
+diff_plot_2 <- control_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases For the Control Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_2
+
+```
+
+#### ALS Accurate vs Inaccurate
+
+```{r, warning = F, message = F}
+
+als_diff_df <- plot_diff(m3_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "ALS", Type = "ALS Inaccurate vs Accurate") |>
+  add_limits_etc()
+
+als_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(als_diff_df)) {
+  als_diff_df$signif_chunk[row_index] <- 
+    ifelse(als_diff_df$is_signif[row_index] != als_diff_df$is_signif[row_index - 1],
+           als_diff_df$signif_chunk[row_index - 1] + 1,
+           als_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[5])
+
+diff_fill <- c( my_pal[5])
+
+diff_plot_3 <- als_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases for the ALS Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_3
+
+```
+
 ```{r, eval = F}

### 2024-05-27 Manuscript Analysis/adaptation_analysis.Rmd
added model comparison

@@
 
-summary(rep_acc_mod)
-
 sjPlot::tab_model(rep_acc_mod, transform = NULL, show.est = TRUE, show.se = TRUE,
@@
 
-sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple,
+sjPlot::tab_model(m5_ple, m6_ple, m7_ple, m8_ple, show.ci = F,
                   pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
@@
 m0_pupil <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = max_pupil)
-summary(m0_pupil)
+sjPlot::tab_model(m0_pupil)
 performance::icc(m0_pupil)
@@
 m1_pupil <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = max_pupil)
-summary(m1_pupil)
+sjPlot::tab_model(m1_pupil)
 performance::test_performance(m0_pupil, m1_pupil)
@@
 m2_pupil <- glmmTMB(peak_pupil ~ trial_c + (trial_c|subject), data = max_pupil)
-summary(m2_pupil)
+sjPlot::tab_model(m2_pupil)
 performance::test_performance(m1_pupil, m2_pupil)
@@
 m3_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + (trial_c|subject), data = max_pupil)
-summary(m3_pupil)
+sjPlot::tab_model(m3_pupil)
 performance::test_performance(m2_pupil, m3_pupil)
@@
 m4_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = max_pupil)
-summary(m4_pupil)
+sjPlot::tab_model(m4_pupil)
 performance::test_performance(m3_pupil, m4_pupil)
@@
 m5_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
-summary(m5_pupil)
+sjPlot::tab_model(m5_pupil)
 performance::test_performance(m4_pupil, m5_pupil)
@@
 m5b_pupil <- glmmTMB(peak_pupil ~ speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
-summary(m5b_pupil)
+sjPlot::tab_model(m5b_pupil)
 performance::test_performance(m5_pupil, m5b_pupil)
@@
                       speaker*rep_acc + trial_c*speaker +(trial_c|subject), data = max_pupil)
-summary(m6_pupil)
+sjPlot::tab_model(m6_pupil)
 performance::test_performance(m4_pupil, m6_pupil)
@@
                       speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + (trial_c|subject), data = max_pupil)
-summary(m7_pupil)
+sjPlot::tab_model(m7_pupil)
 performance::test_performance(m6_pupil, m7_pupil)
@@
                       trial_c*speaker*rep_acc + (trial_c|subject), data = max_pupil)
-summary(m8_pupil)
+sjPlot::tab_model(m8_pupil)
 performance::test_performance(m7_pupil, m8_pupil)
@@
 adapt_pupil <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 25) +
-                  te(time_norm, trial_c, k = 25) +
+                  s(time_norm, by = condition, k = 30) +
+                  te(time_norm, trial_c, k = 30) +
                   # Random Smooth for listener
@@
 
-summary(adapt_pupil)
+summary.gam(adapt_pupil)
 
@@
 
+```
+
+```{r}
+
+vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "Control Accurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "Control Inaccurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "ALS Accurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "ALS Inaccurate"), 
+        theta = -45, phi = 55, color = "topo")
 ```

### 2024-05-27 Manuscript Analysis/main_analysis.Rmd
added model comparison

--- 2024-05-22 Manuscript Analysis/main_analysis.Rmd
@@
 
-library(rio)
-library(tidyverse)
-library(mgcv)
-library(itsadug)
-library(irr)
-library(ggpubr) # installl.packages("ggpubr")
+library(rio) # install.packages("rio")
+library(tidyverse) # install.packages("tidyverse")
+library(mgcv) # install.packages("mgcv")
+library(itsadug) # install.packages("itsadug")
+library(irr) # install.packages("irr")
+library(ggpubr) # install.packages("ggpubr")
 library(gt) # install.packages("gt")
@@
 
-Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots too.
+Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots to 25.
 
-```{r, warning = F, message = F}
+```{r}
 
@@
             plot_all = "condition",
-            rug = F, se = 1, rm.ranef = T, plot = F)
+            rug = F, se = 1, rm.ranef = T)
 
@@
 Therefore, it was removed in this model.
-The random smooth for listener remained in this model
+The random smooth for listener remained in this model.
 
-```{r, warning = F, message = F}
+```{r}
 
@@
 
-## Final Pupil Dilation Model Plot
+### Model 3
+
+In model 3, we are testing whether adjusting the random smooth specification significantly improves model fit to the data.
+In model 2, there was just a random smooth for listener.
+This random smooth specification is updated here to be a random smooth per listener for each condition level (interactions between speaker and phrase repetition accuracy).
+
+```{r}
+
+m3_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooths for listener
+                  s(time_norm, listener, by = condition, bs = 'fs', m =1),
+                discrete = T,
+                AR.start = start_event,
+                rho = start_value_rho(m0_pupil),
+                family = "scat",
+                data = pupil_df)
+
+summary.gam(m3_pupil)
+
+m3_prediction <- plot_smooth(m3_pupil,
+            view = "time_norm",
+            plot_all = "condition",
+            rug = F, se = 1, rm.ranef = T)
+
+```
+
+#### Model Check
+
+The model residuals look similar. The adjusted R-squared value is .192.
+
+```{r}
+
+gam.check(m3_pupil)
+
+acf_resid(m3_pupil)
+
+```
+
+## Model Comparison
+
+To test whether model 3 significantly improved the fit to the data, a generalized likelihood ratio test was conducted.
+The results suggest that model 3 did not significantly improve model fit.
+Therefore, we will retain model 2.
+
+```{r}
+
+compareML(m2_pupil, m3_pupil, signif.stars = T, suggest.report = T)
+
+```
+
+## Final Model (Model 2)
 

### 2024-05-27 Manuscript Analysis/adaptation_analysis.Rmd
adaptation GAMM model

@@
       color = speaker,
-      fill = speaker) +
+      fill = speaker, 
+      linetype = speaker,
+      shape = speaker) +
   geom_point() +
@@
   scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_shape_manual(values = c(1, 16)) +
+  scale_linetype_manual(values = c("solid", "dashed"))
 
@@
        color = speaker,
-       fill = speaker) +
+       fill = speaker,
+       shape = speaker,
+       linetype = speaker) +
   geom_point() +
@@
       color = speaker,
-      fill = speaker) +
+      fill = speaker,
+      linetype = speaker,
+      shape = speaker) +
   geom_point() +
@@
   scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_shape_manual(values = c(1, 16)) +
+  scale_linetype_manual(values = c("solid", "dashed"))
+
+ggsave("Figures/peak_pupil_trial_plot.png", plot = last_plot())
 
@@
 m0_pupil <- glmmTMB(peak_pupil ~ 1 + (1|subject), data = max_pupil)
-sjPlot::tab_model(m0_pupil)
+sjPlot::tab_model(m0_pupil,
+                  pred.labels = c("Intercept"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::icc(m0_pupil)
@@
 m1_pupil <- glmmTMB(peak_pupil ~ 1 + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m1_pupil)
+sjPlot::tab_model(m1_pupil, 
+                  pred.labels = c("Intercept"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m0_pupil, m1_pupil)
@@
 m2_pupil <- glmmTMB(peak_pupil ~ trial_c + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m2_pupil)
+sjPlot::tab_model(m2_pupil, 
+                  pred.labels = c("Intercept", "Trial Order"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m1_pupil, m2_pupil)
@@
 m3_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m3_pupil)
+sjPlot::tab_model(m3_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m2_pupil, m3_pupil)
@@
 m4_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m4_pupil)
+sjPlot::tab_model(m4_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m3_pupil, m4_pupil)
@@
 m5_pupil <- glmmTMB(peak_pupil ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m5_pupil)
+sjPlot::tab_model(m5_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m4_pupil, m5_pupil)
@@
 m5b_pupil <- glmmTMB(peak_pupil ~ speaker + rep_acc + speaker*rep_acc + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m5b_pupil)
+sjPlot::tab_model(m5b_pupil,
+                  pred.labels = c("Intercept", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m5_pupil, m5b_pupil)
@@
                       speaker*rep_acc + trial_c*speaker +(trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m6_pupil)
+sjPlot::tab_model(m6_pupil, 
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m4_pupil, m6_pupil)
@@
                       speaker*rep_acc + trial_c*speaker + trial_c*rep_acc + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m7_pupil)
+sjPlot::tab_model(m7_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m6_pupil, m7_pupil)
@@
                       trial_c*speaker*rep_acc + (trial_c|subject), data = max_pupil)
-sjPlot::tab_model(m8_pupil)
+sjPlot::tab_model(m8_pupil,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]",
+                                  "Trial Order * Speaker * Repetition Accuracy"),
+                  dv.labels = "Peak Pupil Dilation")
 performance::test_performance(m7_pupil, m8_pupil)
@@
 
-```{r}
+```{r, eval = F, warning = F, message = F}
 
-rm(m0_pupil, m1_pupil, m2_pupil, m3_pupil, m4_pupil, m5_pupil, m6_pupil, m7_pupil, m8_pupil, m5b_pupil, max_pupil)
+sjPlot::tab_model(m1_pupil, m2_pupil, m3_pupil, m4_pupil, show.ci = F,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]"),
+                  dv.labels = c("Model 1", "Model 2", "Model 3", "Model 4"),
+                  file = "Tables/peak_pupil_adapt_models1.html")
+
+sjPlot::tab_model(m5_pupil, m6_pupil, m7_pupil, m8_pupil, show.ci = F,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]", 
+                                  "Trial Order * Speaker * Repetition Accuracy"),
+                  dv.labels = c("Model 5", "Model 6", "Model 7", "Model 8"),
+                  file = "Tables/peak_pupil_adapt_models2.html")
 
@@
 
-## GAMM
+```{r}
 
-```{r, eval = F}
+rm(m0_pupil, m1_pupil, m2_pupil, m3_pupil, m4_pupil, m5_pupil, m6_pupil, m7_pupil, m7_pupil_std, m8_pupil, m5b_pupil, max_pupil)
+
+```
+
+## GAMM Analysis
+
+```{r}
 
 adapt_pupil <- bam(normed_pupil ~ condition +
-                  s(time_norm, by = condition, k = 30) +
-                  te(time_norm, trial_c, k = 30) +
+                  s(time_norm, by = condition, k = 35) +
+                  te(time_norm, trial_c, k = 35) +
                   # Random Smooth for listener
@@
 ```
+
+### By Trial Contours
 
@@
 ```
+### Curve Differences
+
+#### Control vs ALS
+
+```{r, warning = F, message = F}
+
+add_limits_etc <- function(mod){
+  mod <- mod |>
+    dplyr::mutate(lower = est - CI,
+                  upper = est + CI,
+                  is_signif = ifelse(sign(upper) == sign(lower), "yes", "no"))
+
+  return(mod)
+  }
+
+speaker_diff_df <- plot_diff(adapt_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Accurate", "Control Accurate")), plot = F) |>
+  mutate(speaker_1 = "ALS", speaker_2 = "Control", rep_acc = "Accurate", Type = " ALS vs Control") |>
+  add_limits_etc()
+
+speaker_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(speaker_diff_df)) {
+  speaker_diff_df$signif_chunk[row_index] <- 
+    ifelse(speaker_diff_df$is_signif[row_index] != speaker_diff_df$is_signif[row_index - 1],
+           speaker_diff_df$signif_chunk[row_index - 1] + 1,
+           speaker_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[4])
+
+diff_fill <- c( my_pal[4])
+
+diff_plot_1 <- speaker_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "ALS vs Control for Accurately \nRecognized Phrases") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {speaker_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_1
+
+```
+
+#### Control Accurate vs Inaccurate
+
+```{r, warning = F, message = F}
+
+control_diff_df <- plot_diff(adapt_pupil, view = "time_norm",
+          comp = list(condition = c("Control Inaccurate", "Control Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "Control", Type = "Control Inaccurate vs Accurate") |>
+  add_limits_etc()
+
+control_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(control_diff_df)) {
+  control_diff_df$signif_chunk[row_index] <- 
+    ifelse(control_diff_df$is_signif[row_index] != control_diff_df$is_signif[row_index - 1],
+           control_diff_df$signif_chunk[row_index - 1] + 1,
+           control_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[1])
+
+diff_fill <- c( my_pal[1])
+
+diff_plot_2 <- control_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases For the Control Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {control_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_2
+
+```
+
+#### ALS Accurate vs Inaccurate
+
+```{r, warning = F, message = F}
+
+als_diff_df <- plot_diff(adapt_pupil, view = "time_norm",
+          comp = list(condition = c("ALS Inaccurate", "ALS Accurate")), plot = F) |>
+  mutate(rep_1 = "Inaccurate", rep_2 = "Accurate", speaker = "ALS", Type = "ALS Inaccurate vs Accurate") |>
+  add_limits_etc()
+
+als_diff_df$signif_chunk <- 1
+
+for (row_index in 2:nrow(als_diff_df)) {
+  als_diff_df$signif_chunk[row_index] <- 
+    ifelse(als_diff_df$is_signif[row_index] != als_diff_df$is_signif[row_index - 1],
+           als_diff_df$signif_chunk[row_index - 1] + 1,
+           als_diff_df$signif_chunk[row_index - 1])
+
+}
+
+diff_color <- c(my_pal[5])
+
+diff_fill <- c( my_pal[5])
+
+diff_plot_3 <- als_diff_df |>
+  ggplot() +
+  aes(x = time_norm,
+      y = est,
+      ymin = lower,
+      ymax = upper) +
+  geom_ribbon(fill = "white", alpha = 1) +
+  geom_ribbon(fill = "grey60", alpha = .5) +
+  geom_line() +
+  scale_color_manual(values = diff_color) +
+  scale_fill_manual(values = diff_fill) +
+  geom_hline(yintercept = 0) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = -20, ymax = 200, alpha = 0.15) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  labs(x = "Normed Time (ms)",
+       y = "Estimated Difference in Pupil Dilation",
+       title = "Inaccurately vs Accurately Recognized \nPhrases for the ALS Speaker") +
+  coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
+  theme_bw() +
+  geom_ribbon(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, fill = is_signif), alpha = .4, show.legend = F) +
+  geom_line(data = {als_diff_df |>
+                dplyr::filter(is_signif != "no")},
+              aes(group = signif_chunk, color = is_signif), alpha = .8, show.legend = F) +
+  theme(aspect.ratio = 1,
+        plot.title = element_text(hjust = 0.5))
+
+diff_plot_3
+
+```

### 2024-05-27 Manuscript Analysis/main_analysis.Rmd
adaptation GAMM model

@@
 
-sjPlot::tab_model(m1_ple, m2_ple, m3_ple, m4_ple, m5_ple,
+sjPlot::tab_model(m1_ple, m2_ple, m3_ple, m4_ple, m5_ple, show.ci = F,
                   pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
@@
 
-```{r}
+```{r, warning = F}
 

### 2024-05-27 Manuscript Analysis/adaptation_analysis.Rmd
updated adaptation analysis

@@
 
+### Main Model
+
+```{r}
+
+m2_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil_df)
+
+```
+
+### Adaptation Model
+
 ```{r}
@@
                 AR.start = start_event,
-                rho = .92,
+                rho = .9341941,
                 data = pupil_df)
@@
 
-### By Trial Contours
-
-```{r}
-
-vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
-        cond = list(condition = "Control Accurate"), 
-        theta = -45, phi = 55, color = "topo")
-
-vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
-        cond = list(condition = "Control Inaccurate"), 
-        theta = -45, phi = 55, color = "topo")
-
-vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
-        cond = list(condition = "ALS Accurate"), 
-        theta = -45, phi = 55, color = "topo")
-
-vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
-        cond = list(condition = "ALS Inaccurate"), 
-        theta = -45, phi = 55, color = "topo")
-```
+#### Model Check
 
@@
 ```
+
+#### Model Comparison
+
+```{r}
+
+compareML(m2_pupil, adapt_pupil, signif.stars = T, suggest.report = T)
+
+```
+#### By Trial Contours
+
+```{r}
+
+control_acc <- vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "Control Accurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+control_inacc <- vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "Control Inaccurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+ALS_acc <- vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "ALS Accurate"), 
+        theta = -45, phi = 55, color = "topo")
+
+ALS_inacc <- vis.gam(adapt_pupil, view = c("time_norm", "trial_c"), 
+        cond = list(condition = "ALS Inaccurate"), 
+        theta = -45, phi = 55, color = "topo")
+```
+
 ### Curve Differences

### 2024-05-28 Manuscript Analysis/adaptation_analysis.Rmd
Updated some figures

--- 2024-05-27 Manuscript Analysis/adaptation_analysis.Rmd
@@
 library(sjPlot) # install.packages("sjPlot")
+library(viridis) # install.packages("viridis")
 
@@
 
+### Logistic Mixed-Effects Model
+
 ```{r, warning = F, message = F}
@@
                                   "Trial Order * Speaker [ALS]"),
-                  dv.labels = "Perceived Listening Effort",
+                  dv.labels = "Accurately Recognized Phrases",
                   file = "Tables/rep_acc_model.html")
@@
 
-```{r}
+### Main Model
+
+```{r, warning = F, message = F}
+
+m2_pupil <- bam(normed_pupil ~ condition +
+                  s(time_norm, by = condition, k = 25) +
+                  # Random Smooth for listener
+                  s(time_norm, listener, bs = 'fs', m = 1),
+                discrete = T,
+                AR.start = start_event,
+                rho = .9341941,
+                family = "scat",
+                data = pupil_df)
+
+```
+
+### Adaptation Model
+
+```{r, warning = F, message = F}
 
@@
 
-### By Trial Contours
+#### Model Check
 
-```{r}
+```{r, eval = F}
+
+gam.check(adapt_pupil)
+
+acf_resid(adapt_pupil)
+
+```
+
+#### Model Comparison
+
+```{r, warning = F}
+
+compareML(m2_pupil, adapt_pupil, signif.stars = T, suggest.report = T)
+
+```
+#### By Trial Contours
+
+```{r, warning = F, message = F}
 
@@
 
-```{r, eval = F}
+```{r, warning = F, message = F}
 
-gam.check(adapt_pupil)
+df_pred <- expand_grid(
+  time_norm = seq(from=min(pupil_df$time_norm), 
+              to=max(pupil_df$time_norm), 
+              length.out = 100),
+  trial_c = seq(from=min(pupil_df$trial_c), 
+              to=max(pupil_df$trial_c), 
+              length.out = 100),
+  condition = c("Control Accurate", "Control Inaccurate", "ALS Accurate", "ALS Inaccurate"),
+  listener = c("LE01", "LE02", "LE04", "LE05", "LE06", "LE07", "LE08", "LE09", "LE10", "LE11",
+               "LE12", "LE13", "LE15", "LE16", "LE18", "LE19", "LE20", "LE21", "LE22", "LE23",
+               "LE24", "LE25", "LE26", "LE27", "LE28", "LE29", "LE30", "LE31", "LE32", "LE33", 
+               "LE34", "LE35", "LE36", "LE38")
+)
 
-acf_resid(adapt_pupil)
+df_pred <- predict(adapt_pupil, newdata = df_pred, 
+                      se.fit = TRUE) |> 
+  as_tibble() |>
+  cbind(df_pred)
+
+df_pred |>
+  dplyr::mutate(condition = factor(condition, levels = c("Control Accurate", "Control Inaccurate",
+                                                         "ALS Accurate", "ALS Inaccurate"))) |>
+  dplyr::filter(time_norm >= 0 & time_norm <= 5000) |>
+  ggplot() +
+  aes(x = time_norm,
+      y = trial_c,
+      group = condition) +
+  geom_tile(aes(fill = fit)) +
+  geom_contour(aes(z = fit), colour = "white") +
+  facet_wrap("condition") +
+  theme_classic() +
+  labs(fill = "Model Estimate",
+       x = "Normed Time (ms)",
+       y = "Trial Number") +
+  scale_fill_viridis(option = "D") +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom",
+        axis.line = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust=1))
+
+ggsave("Figures/pupil_heatmap.png", plot = last_plot())
 
 ```
+
 ### Curve Differences

### 2024-05-28 Manuscript Analysis/main_analysis.Rmd
Updated some figures

@@
       color = speaker,
-      fill = speaker) +
+      fill = speaker,
+      linetype = speaker) +
   geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
@@
   scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
+  scale_linetype_manual(values = c("solid", "dashed"))
 
@@
 #### Control vs ALS
+
+First the difference in pupil dilation between the control and ALS speakers for accurately recognized phrases is plotted below.
+The grey box indicates the analysis region of interest.
+The dotted line represents the average phrase offset time.
+The highlighted sections of the line (i.e., the green sections) indicate there was a significant difference in pupil dilation between the two speakers.
+Positive values indicate that pupil dilation while listening to the ALS speaker was greater than the control speaker for accurately recognized phrases.
 
@@
        y = "Estimated Difference in Pupil Dilation",
-       title = "ALS vs Control for Accurately \nRecognized Phrases") +
+       title = "ALS vs Neurotypical Speaker") +
   coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
@@
 #### Control Accurate vs Inaccurate
+
+The difference in pupil dilation for accurately and inaccurately recognized phrases for the control speaker are shown below.
+A significant difference is indicated by the purple highlight.
+Positive values mean that inaccurately recognized phrases had larger pupil dilation at that time point.
+The grey box indicates the analysis region of interest.
+The dotted line represents the average phrase offset time.
 
@@
        y = "Estimated Difference in Pupil Dilation",
-       title = "Inaccurately vs Accurately Recognized \nPhrases For the Control Speaker") +
+       title = "Neurotypical Speaker") +
   coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
@@
 #### ALS Accurate vs Inaccurate
+
+The difference in pupil dilation for accurately and inaccurately recognized phrases for the ALS speaker are shown below.
+A significant difference is indicated by the aqua highlight.
+Positive values mean that inaccurately recognized phrases had larger pupil dilation at that time point.
+The grey box indicates the analysis region of interest.
+The dotted line represents the average phrase offset time.
 
@@
        y = "Estimated Difference in Pupil Dilation",
-       title = "Inaccurately vs Accurately Recognized \nPhrases for the ALS Speaker") +
+       title = "ALS Speaker") +
   coord_cartesian(ylim = c(-10, 150), xlim = c(0, 4910)) +
@@
 
+ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1)
+
 library(patchwork)
 
-diff_plot_1 + diff_plot_2 + diff_plot_3
+diff_plot_2 + diff_plot_3
 
-ggsave("Figures/difference_plots.png", plot = last_plot(), width = 6, height = 2.5, units = "in", scale = 1.8)
+ggsave("Figures/difference_plots_accuracy.png", plot = last_plot(), width = 4, height = 2.5, units = "in", scale = 1.8)
 

### 2024-05-29 Manuscript Analysis/adaptation_analysis.Rmd
updated figures

@@
   labs(y = "Proportion of Accurate Responses",
-       x = "Trial Number") +
+       x = "Trial Number",
+       color = "Speaker", fill = "Speaker", 
+       linetype = "Speaker", shape = "Speaker") +
   theme_bw() +
@@
         aspect.ratio = 1) +
-  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_shape_manual(values = c(1, 16)) +
-  scale_linetype_manual(values = c("solid", "dashed"))
+  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_shape_manual(values = c(1, 16), labels = c("Neurotypical", "ALS")) +
+  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))
 
@@
   geom_smooth(method = "lm") +
-  labs(x = "Trial", y = "Perceived Listening Effort Rating") +
+  labs(x = "Trial", y = "Perceived Listening Effort Rating",
+       color = "Speaker", fill = "Speaker",
+       shape = "Speaker", linetype = "Speaker") +
   facet_wrap("rep_acc") +
@@
   geom_smooth(method = "lm") +
-  labs(x = "Trial", y = "Peak Pupil Dilation (Arbitrary Units)") +
+  labs(x = "Trial", y = "Peak Pupil Dilation (Arbitrary Units)",
+       color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +
   facet_wrap("rep_acc") +
@@
   dplyr::mutate(condition = factor(condition, levels = c("Control Accurate", "Control Inaccurate",
+                                                         "ALS Accurate", "ALS Inaccurate"),
+                                   labels = c("Neurotypical Accurate", "Neurotypical Inaccurate",
                                                          "ALS Accurate", "ALS Inaccurate"))) |>
@@
   theme_classic() +
-  labs(fill = "Model Estimate",
+  labs(fill = "Pupil Dilation Estimate",
        x = "Normed Time (ms)",
@@
 
-ggsave("Figures/pupil_heatmap.png", plot = last_plot())
+ggsave("Figures/pupil_heatmap.png", plot = last_plot(), scale = 1.8)
 

### 2024-05-29 Manuscript Analysis/main_analysis.Rmd
updated figures

@@
   geom_errorbar(aes(ymin = per_effort - se, ymax = per_effort + se), width = 0.4, position = position_dodge(.9)) +
-  labs(x = "Accuracy", y = "Perceived Listening Effort Rating") +
+  labs(x = "Accuracy", y = "Perceived Listening Effort Rating", color = "Speaker", fill = "Speaker") +
   theme_bw() +
   theme(aspect.ratio = 1) +
-  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]))
+  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels =  c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS"))
 
@@
   labs(x = "Normalized Time (ms)",
-       y = "Pupil Dilation (Arbitrary Units)") +
-  scale_color_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5])) +
-  scale_linetype_manual(values = c("solid", "dashed"))
+       y = "Pupil Dilation (Arbitrary Units)",
+       color = "Speaker",
+       fill = "Speaker",
+       linetype = "Speaker") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))
 
@@
 
-ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1)
+ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1, scale = 1.8)
 

### 2024-06-17 Manuscript Analysis/exploratory_figures.Rmd
Created pupil dilation and intelligibility heatmap

---
title: "Exploratory Figures"
author: "Micah E. Hirsch"
date: '`r Sys.Date()`'
output:  
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

The purpose of this document is to conduct and create a few exploratory analyses and figures. These analyses are unplanned and not part of the main study, but were mainly created to include in M. Hirsch's dissertation defense.

```{r}

library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
library(mgcv) # install.packages("mgcv")
library(itsadug) # install.packages("itsadug")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(gtsummary) # install.packages("gtsummary")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(viridis) # install.packages("viridis")

```

# Listening Effort and Intelligibility

## Pupil Dilation

```{r}

phrase_acc <- rio::import("Cleaned Data/repetition_accuracy.csv") |>
  dplyr::select(subject, trial, targetphrase, initial_response, target_number, correct_words_initial, rep_acc) |>
  dplyr::mutate(rep_acc = ifelse(initial_response == "missing data", NA, 
                                 ifelse(target_number == correct_words_initial, "accurate", "inaccurate")),
                intel = (correct_words_initial/target_number) * 100) |>
  dplyr::select(!c(targetphrase, initial_response))

pupil_df <- rio::import("Cleaned Data/cleaned_pupil_data_normalized.csv") |>
  # Merging phrase accuracy information with df
  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
  # Filter out trials with missing phrase repetition data 
  dplyr::filter(!is.na(rep_acc)) |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate")),
                phrase = as.factor(code),
                listener = as.factor(subject),
                trial_c = trial - 6) |>
  dplyr::group_by(listener, speaker) |>
  dplyr::mutate(num_points = n()) |>
  dplyr::mutate(start_event = c(TRUE, rep(FALSE, each = (num_points - 1)))) |>
  ungroup()

```

```{r}

intel_pupil <- bam(normed_pupil ~ speaker +
                  s(time_norm, by = speaker, k = 10) +
                  te(time_norm, intel, k = 10) +
                  # Random Smooth for listener
                  s(time_norm, listener, bs = 'fs', m =1),
                discrete = T,
                family = "scat",
                AR.start = start_event,
                rho = .9341941,
                data = pupil_df)

# summary.gam(intel_pupil)

gam.check(intel_pupil)

acf_resid(intel_pupil)

```

```{r}

df_pred <- expand_grid(
  time_norm = seq(from=min(pupil_df$time_norm), 
              to=max(pupil_df$time_norm), 
              length.out = 100),
  intel = seq(from=min(pupil_df$intel), 
              to=max(pupil_df$intel), 
              length.out = 100),
  speaker = c("Control", "ALS"),
  listener = c("LE01", "LE02", "LE04", "LE05", "LE06", "LE07", "LE08", "LE09", "LE10", "LE11",
               "LE12", "LE13", "LE15", "LE16", "LE18", "LE19", "LE20", "LE21", "LE22", "LE23",
               "LE24", "LE25", "LE26", "LE27", "LE28", "LE29", "LE30", "LE31", "LE32", "LE33", 
               "LE34", "LE35", "LE36", "LE38"))

df_pred <- predict(intel_pupil, newdata = df_pred, 
                      se.fit = TRUE) |> 
  as_tibble() |>
  cbind(df_pred)

df_pred |>
  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
                                   labels = c("Neurotypical", "ALS"))) |>
  dplyr::filter(time_norm >= 0 & time_norm <= 5000) |>
  ggplot() +
  aes(x = time_norm,
      y = intel,
      group = speaker) +
  geom_tile(aes(fill = fit)) +
  geom_contour(aes(z = fit), colour = "white") +
  facet_wrap("speaker") +
  theme_classic() +
  labs(fill = "Pupil Dilation Estimate",
       x = "Normed Time (ms)",
       y = "Intelligibility (% Words Correct)") +
  scale_fill_viridis(option = "D") +
  theme(aspect.ratio = 1,
        legend.position = "bottom",
        axis.line = element_blank(),
        axis.text.x = element_text(angle = 45, hjust=1))

ggsave("Figures/intel_heatmap.png", plot = last_plot())

```

## Perceived Listening Effort (PLE) Ratings

```{r}

```

# Correlations between Peak Pupil Dilaiton and PLE Ratings

```{r}

```

### 2024-06-18 Manuscript Analysis/main_analysis.Rmd
created additional pupil figures

@@
 ```
+```{r}
+
+m2_predicted_values |>
+  dplyr::filter(rep_acc == "accurate") |>
+  ggplot() +
+  aes(x = time_norm,
+      y = fit,
+      color = speaker,
+      fill = speaker,
+      linetype = speaker) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line() +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)",
+       title = "ALS vs Neurotypical Speaker",
+       color = "Speaker",
+       fill = "Speaker",
+       linetype = "Speaker") +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
+  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))
+
+ggsave("Figures/pupil_model_speakers.png", plot = last_plot())
+
+```
 
@@
 diff_plot_2
+
+```
+
+```{r}
+
+m2_predicted_values |>
+  dplyr::filter(speaker == "Control") |>
+  ggplot() +
+  aes(x = time_norm,
+      y = fit,
+      linetype = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, fill = my_pal[1], color = NA) +
+  geom_line(color = my_pal[1]) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)",
+       title = "Neurotypical Speaker",
+       linetype = "Repetition Accuracy") +
+  scale_linetype_manual(values = c("solid", "twodash"))
+
+ggsave("Figures/pupil_model_control.png", plot = last_plot())
 
@@
 
+```{r}
+
+m2_predicted_values |>
+  dplyr::filter(speaker == "ALS") |>
+  ggplot() +
+  aes(x = time_norm,
+      y = fit,
+      linetype = rep_acc) +
+  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, fill = my_pal[5], color = NA) +
+  geom_line(color = my_pal[5]) +
+  geom_vline(xintercept = 0, size = 0.35) +
+  geom_vline(xintercept = 1910, linetype = 2) +
+  annotate("rect", xmin = 500, xmax = 2910, ymin = 0, ymax = 200, alpha = 0.15) +
+  coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
+  theme_bw() +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Normalized Time (ms)",
+       y = "Pupil Dilation (Arbitrary Units)",
+       title = "ALS Speaker",
+       linetype = "Repetition Accuracy") +
+  scale_linetype_manual(values = c("solid", "twodash"))
+
+ggsave("Figures/pupil_model_ALS.png", plot = last_plot())
+
+```
+
 ```{r, eval = F}
 
-ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1, scale = 1.8)
+ggsave("Figures/difference_plots_speakers.png", plot = diff_plot_1, scale = .8)
+
+ggsave("Figures/difference_plots_control.png", plot = diff_plot_2, scale = .8)
+
+ggsave("Figures/difference_plots_ALS.png", plot = diff_plot_3, scale = .8)
 

### 2024-06-18 Manuscript Analysis/exploratory_figures.Rmd
intelligibility plots

@@
 
+## Peak Pupil Dilation
+
+### LME Model
+
+```{r}
+
+peak_pupil <- pupil_df |>  
+  dplyr::filter(time_norm >= 500) |>
+  dplyr::group_by(subject, trial) |>
+  # filtering out timebins 1000 ms after phrase offset
+  dplyr::filter(time_norm<= max(time_norm) - 2000) |>
+  dplyr::ungroup() |>
+  # Getting Peak Pupil Dilation Value
+  dplyr::group_by(subject, trial, speaker, code, intel) |>
+  dplyr::summarize(peak_pupil = max(normed_pupil)) |>
+  dplyr::ungroup() |>
+  dplyr::mutate(trial_c = trial - 6)
+
+peak_pupil_model <- glmmTMB(peak_pupil ~ intel*speaker + (1|subject), data = peak_pupil)
+#summary(peak_pupil_model)
+
+peak_pupil_model_2 <- glmmTMB(peak_pupil ~ poly(intel, degree = 2)*speaker + (1|subject), data = peak_pupil)
+#summary(peak_pupil_model_2)
+
+anova(peak_pupil_model, peak_pupil_model_2)
+
+sjPlot::tab_model(peak_pupil_model, peak_pupil_model_2,
+                  file = "Tables/peak_pupil_intel_model.html")
+
+```
+
+### Plot
+
+```{r}
+
+peak_pupil |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
+                                   labels = c("Neurotypical", "ALS")),
+                color = ifelse(speaker == "ALS", my_pal[5], my_pal[1])) |>
+  ggplot() +
+  aes(x = intel,
+      y = peak_pupil,
+      color = speaker) +
+  geom_line(method = "lm", stat = "smooth", aes(group = subject), alpha = .3) +
+  geom_line(stat = "smooth", method = "lm", size = 2) +
+  labs(x = "Intelligibility (% Words Correct)", 
+       y = "Peak Pupil Dilation (Arbitrary Units",
+       color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
+  coord_cartesian(ylim = c(0, 850)) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/pupil_intel.png", plot = last_plot())
+
+```
+
 ## Perceived Listening Effort (PLE) Ratings
 
+### LME Model
+
 ```{r}
+
+ple_data <- rio::import("Cleaned Data/cleaned_ple_data.csv") |>
+  dplyr::left_join(phrase_acc, by = c("subject", "trial")) |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS")),
+                trial_c = trial - 6,
+                rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
+  dplyr::filter(!is.na(rep_acc)) |>
+  dplyr::filter(effort_rating > 0)
+
+ple_model <- glmmTMB(effort_rating ~ intel*speaker +
+                       (trial_c|subject) + (trial_c|code), data = ple_data)
+summary(ple_model)
+
+```
+
+### Plot
+
+```{r}
+
+ple_data |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
+                                   labels = c("Neurotypical", "ALS")),
+                color = ifelse(speaker == "ALS", my_pal[5], my_pal[1])) |>
+  ggplot() +
+  aes(x = intel,
+      y = effort_rating,
+      color = speaker) +
+  geom_line(method = "lm", stat = "smooth", aes(group = subject), alpha = .3) +
+  geom_line(stat = "smooth", method = "lm", size = 2) +
+  labs(x = "Intelligibility (% Words Correct)", 
+       y = "Peak Pupil Dilation (Arbitrary Units",
+       color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +
+  facet_wrap("speaker") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
+  coord_cartesian(ylim = c(0, 7)) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/ple_intel.png", plot = last_plot())
 

### 2024-06-18 Manuscript Analysis/exploratory_figures.Rmd
LE corr plot

@@
   dplyr::filter(!is.na(rep_acc)) |>
-  dplyr::filter(effort_rating > 0)
+  dplyr::filter(effort_rating > 0) |>
+  dplyr::filter(effort_rating <= 7)
 
@@
 
+## Repeated Measures Correlation
+
+### Overall Correlation
+
 ```{r}
 
+peak_pupil <- peak_pupil |>
+  dplyr::select(subject, trial_c, speaker, peak_pupil, intel)
+
+ple_data <- ple_data |>
+  dplyr::select(subject, trial_c, speaker, effort_rating, intel)
+
+corr_df <- dplyr::left_join(peak_pupil, ple_data, by = c("subject", "trial_c", "speaker", "intel"))
+
+rmcorr::rmcorr(subject, peak_pupil, effort_rating, corr_df)
+
 ```
+
+#### Plot
+
+```{r}
+
+corr_df |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
+                                   labels = c("Neurotypical", "ALS")),
+                color = ifelse(speaker == "ALS", my_pal[5], my_pal[1])) |>
+  ggplot() +
+  aes(x = effort_rating,
+      y = peak_pupil) +
+  geom_line(method = "lm", stat = "smooth", aes(group = subject), alpha = .3) +
+  geom_line(stat = "smooth", method = "lm", size = 2) +
+  labs(x = "Perceived Listening Effort Rating", 
+       y = "Peak Pupil Dilation (Arbitrary Units)") +
+  coord_cartesian(xlim = c(1, 7)) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1)
+
+ggsave("Figures/effort_corr.png", plot = last_plot())
+
+```
+
+```{r}
+
+corr_df |>
+  dplyr::mutate(speaker = factor(speaker, levels = c("Control", "ALS"),
+                                   labels = c("Neurotypical", "ALS")),
+                color = ifelse(speaker == "ALS", my_pal[5], my_pal[1])) |>
+  ggplot() +
+  aes(x = effort_rating,
+      y = peak_pupil,
+      color = speaker) +
+  geom_line(method = "lm", stat = "smooth", aes(group = subject), alpha = .3) +
+  geom_line(stat = "smooth", method = "lm", size = 2) +
+  labs(x = "Perceived Listening Effort Rating", 
+       y = "Peak Pupil Dilation (Arbitrary Units)") +
+  facet_wrap("speaker") +
+  coord_cartesian(xlim = c(1, 7)) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        aspect.ratio = 1) +
+  scale_color_manual(values = c(my_pal[1], my_pal[5]))
+
+ggsave("Figures/effort_corr_speaker.png", plot = last_plot())
+
+```

### 2024-06-20 Manuscript Analysis/adaptation_analysis.Rmd
updated pupil dilation heatmap
updated figures and tables

@@
                 rep_acc = factor(rep_acc, levels = c("accurate", "inaccurate"))) |>
-  dplyr::filter(!is.na(rep_acc)) |>
-  dplyr::filter(effort_rating > 0)
+  dplyr::filter(!is.na(rep_acc))
 
@@
 
+ple_data <- ple_data |>
+  dplyr::filter(effort_rating > 0) |>
+  dplyr::filter(effort_rating <= 7)
+
 m5_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + (trial_c|subject) + (trial_c|code), data = ple_data)
@@
 
-```{r, message = F, warning = F}
+There was a convergence issue, so the random slope of trial order was removed from the phrase random intercept.
+Therefore, the significant model comparison below is likely due to the change in random effects.
+
+```{r}
 
 m8_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
-                    speaker*trial_c + rep_acc*trial_c +  speaker*rep_acc*trial_c + (trial_c|subject) + (trial_c|code), data = ple_data)
+                    speaker*trial_c + rep_acc*trial_c +  speaker*rep_acc*trial_c + (trial_c|subject) + (1|code), data = ple_data)
 sjPlot::tab_model(m8_ple,
@@
 performance::test_performance(m5_ple, m8_ple)
+
+```
+
+### Model 8b
+
+To determine whether the three-way interaction really does improve model fit, we reran model 7 without the random slope on phrase and compared this model to model 8 above (we are calling this model 8b).
+The addition of the three-way model did not significantly improve model fit.
+Therefore, the significant comparison between model 7 and model 8 is likely due to the differences in the random effects.
+
+Therefore, PLE ratings are not changing across trials for either speaker or accuracy level.
+
+```{r}
+
+m8b_ple <- glmmTMB(effort_rating ~ trial_c + speaker + rep_acc + speaker*rep_acc + 
+                    speaker*trial_c + rep_acc*trial_c + (trial_c|subject) + (1|code), data = ple_data)
+sjPlot::tab_model(m8b_ple,
+                  pred.labels = c("Intercept", "Trial Order", "Speaker [ALS]", "Repetition Accuracy [Inaccurate]",
+                                  "Speaker [ALS] * Repetition Accuracy [Inaccurate]", "Trial Order * Speaker [ALS]",
+                                  "Trial Order * Repetition Accuracy [Inaccurate]"),
+                  dv.labels = "Perceived Listening Effort")
+
+performance::test_performance(m8_ple, m8b_ple)
 
@@
 
-ggsave("Figures/pupil_heatmap.png", plot = last_plot(), scale = 1.8)
+ggsave("Figures/pupil_heatmap.png", plot = last_plot(), scale = .9)
 

### 2024-06-20 Manuscript Analysis/exploratory_figures.Rmd
updated figures and tables

@@
 
+my_pal <- c("#4E4187", "#3083DC", "#F8FFE5", "#7DDE92", "#2EBFA5")
+
 peak_pupil |>
@@
   labs(x = "Intelligibility (% Words Correct)", 
-       y = "Peak Pupil Dilation (Arbitrary Units",
+       y = "Peak Pupil Dilation (Arbitrary Units)",
        color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +
@@
   labs(x = "Intelligibility (% Words Correct)", 
-       y = "Peak Pupil Dilation (Arbitrary Units",
+       y = "Perceived Listening Effort Rating",
        color = "Speaker", fill = "Speaker", linetype = "Speaker", shape = "Speaker") +

### 2024-06-20 Manuscript Analysis/main_analysis.Rmd
updated figures and tables

@@
   dplyr::filter(effort_rating > 0) |>
+  dplyr::filter(effort_rating <= 7) |>
   tbl_strata(

### 2024-10-10 Manuscript Analysis/pupil_data_prep.R
fixed ALS downsampling issue

@@ -2,7 +2,7 @@
 
 # Author: Micah E. Hirsch, mhirsch@fsu.edu
 
-## Data: 5/13/2024
+## Date: 10/10/2024
 
 ## Purpose: To prepare the pupil dilation data for analysis.
 
@@ -19,7 +19,7 @@ library(knitr) # install.packages("knitr")
 
 # Set the working directory to load data
 
-setwd("~/Documents/Listening-Effort-in-Dysarthria/Raw Data")
+setwd("D:\\Listening Effort Study\\Raw Data\\Extracted Pupil and PLE Ratings")
 
 # Load raw pupil files
 
@@ -386,30 +386,32 @@ data.binned <- filtered_df |>
   dplyr::summarize(pupil.binned = mean(baselinecorrectedp)) |>
   dplyr::ungroup()
 
-# Export Cleaned Data
+normed_data |>
+  # Selecting relevant parameters
+  dplyr::select(subject, trial, speaker, time_n, code) |>
+  dplyr::group_by(speaker, trial) |>
+  dplyr::filter(time_n >= 0 & time_n <= max(time_n) - 3000) |>
+  dplyr::group_by(speaker) |>
+  summarize(count = n())
+  
 
-## Set working directory
-setwd("~/Documents/Listening-Effort-in-Dysarthria/Manuscript Analysis/Cleaned Data")
-
-## Export Pupil Dilation DF
-rio::export(data.binned, "cleaned_pupil_data.csv")
-
-## Export PLE Ratings
-rio::export(ple_data, "cleaned_ple_data.csv")
 
 # Downsampling ALS speaker trials
 ## Based on findings from the 2023 ASHA Convention data analysis,the trials from the ALS speaker are much longer than the control talker.
 ## Therefore we are creating a separate df that downsamples the ALS speaker's trials.
 
-bin.length <- 26.23
+bin.length <- 48.5
 
 ALS_trials <- data.binned |>
   dplyr::filter(speaker == "ALS") |>
-  dplyr::mutate(time_n = round(timebins/bin.length)*bin.length) |>
+  dplyr::group_by(subject, trial) |>
+  dplyr::mutate(time_n = case_when(timebins >= 0 & timebins <= max(timebins) - 3000 ~ round(timebins/bin.length)*bin.length,
+                TRUE ~ timebins)) |>
+  dplyr::ungroup() |>
   dplyr::group_by(subject, trial, speaker, time_n, code, targetphrase, counterbalance) |>
[…]

### 2024-10-10 Manuscript Analysis/main_analysis.Rmd
rerun some figures

--- 2024-05-29 Manuscript Analysis/main_analysis.Rmd
@@
       color = speaker,
-      fill = speaker,
-      linetype = speaker) +
-  geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
-  geom_line() +
+      linetype = rep_acc) +
+  #geom_ribbon(aes(ymin = ul, ymax = ll), alpha = .5, color = NA) +
+  geom_line(size = .8) +
   geom_vline(xintercept = 0, size = 0.35) +
@@
   coord_cartesian(ylim = c(-10, 250), xlim = c(0, 4910)) +
-  facet_wrap("rep_acc") +
   theme_bw() +
@@
        fill = "Speaker",
-       linetype = "Speaker") +
+       linetype = "Repetition Accuracy") +
   scale_color_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
-  scale_fill_manual(values = c(my_pal[1], my_pal[5]), labels = c("Neurotypical", "ALS")) +
-  scale_linetype_manual(values = c("solid", "dashed"), labels = c("Neurotypical", "ALS"))
+  scale_linetype_manual(values = c("solid", "dashed"))
 

### 2024-12-17 Manuscript Analysis/main_analysis.Rmd
updated figures

@@
 
-The purpose of this document is to analyze and report findings related to M. Hirsch's dissertation project. The data preparation and analysis was conducted using R Version 4.3.3.
+The purpose of this document is to analyze and report findings related to M. Hirsch's dissertation project. The results from this analysis were also used to prepare for publication.
+
+The data preparation and analysis was conducted using R Version 4.4.1.
 
@@
   theme_bw() +
-  theme(aspect.ratio = 1) +
+  theme(aspect.ratio = 1,
+        legend.position = "bottom") +
   scale_color_manual(values = c(my_pal[1], my_pal[5]), labels =  c("Neurotypical", "ALS")) +
@@
 
-Adding in AR(1) structure to address autocorrelation concerns. Bumped up knots to 25.
+Adding in AR(1) structure to address autocorrelation concerns.
 

### 2025-02-11 Manuscript Analysis/main_analysis.Rmd

@@
                   file = "Tables/ple_mod_unstd.html")
-
 ```
@@
 ```
-```{r}
+
+##### Estimated Peak Differences
+
+```{r, warning = F, message = F}
+
+max_1 <- speaker_diff_df |>
+  dplyr::filter(time_norm >= 950 & time_norm <=1200 ) |>
+  dplyr::slice_max(est,n=1)
+
+max_2 <- speaker_diff_df |>
+  dplyr::filter(time_norm >= 2500 & time_norm <=3500 ) |>
+  dplyr::slice_max(est,n=1)
+
+max_values <- rbind(max_1, max_2) |>
+  dplyr::select(time_norm, est, Type)
+
+max_values |>
+  dplyr::rename("Normed Time" = time_norm,
+                "Estimated Difference" = est,
+                "Contrast" = Type) |>
+  gt::gt()
+
+```
+
+```{r, warning = F, message = F}
 
@@
 
+##### Estimated Peak Differences
+
+```{r, warning = F, message = F}
+
+max_1 <- als_diff_df |>
+  dplyr::filter(time_norm >= 1300 & time_norm <=1700 ) |>
+  dplyr::slice_max(est,n=1)
+
+max_2 <- als_diff_df |>
+  dplyr::filter(time_norm >= 3500 & time_norm <=4000 ) |>
+  dplyr::slice_max(est,n=1)
+
+max_values <- rbind(max_1, max_2) |>
+  dplyr::select(time_norm, est, Type)
+
+max_values |>
+  dplyr::rename("Normed Time" = time_norm,
+                "Estimated Difference" = est,
+                "Contrast" = Type) |>
+  gt::gt()
+
+```
+
 ```{r, eval = F}
